### 2022-05-03 paper/paper.tex
Some updates on paper and on lean development

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

\usepackage{minted}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}
%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Towards a Sound and Complete Library for Modal Logics Reasoning in Coq}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ariel Agne da Silveira}
\author{Paulo Torrens}
\author{Karina Roggia}
\author{Rodrigo Ribeiro}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{da Silveira et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  The correct specification of requirements is crucial for software development.
  Modal logics are ...
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Modal logics, soundness, completeness, Coq proof assistant}

\maketitle

\section{Introduction}\label{sec:intro}

Our society depends heavily on computing technology: our cars, airplanes, medical devices
and banking applications all rely on software artifacts. In this scenario,
\emph{formal methods} play a crucial role to ensure that a program follows its
specification, thus avoiding failures which can cause castrophic outcomes. Interactive
proof assistants, like Coq~\cite{manual_coq} and Isabelle/HOL~\cite{manual_isabelle}, can
be used to verify software components and have been applied with success in
non-trivial projects like a certified compiler for C programming
language~\cite{Leroy09} and the kernel of seL4 operating system~\cite{Klein10}.

Usually, complex software systems are formed by various inter-communicating
components. A key aspect of the design of such programs is how to ensure that
a failure of a single component does not compromise the whole system.
The main difficulty is such formalization tasks is how to guarantee correct
error detection and that its corresponding fix will ensure that the software
sill continue to work under abnormal conditions. One way to express such
requirements is using \emph{modal logics}, which can specify communication
patterns and knowledge in software agents~\cite{Huth_Ryan}.

While modal logics allow the modelling of such requirements, proving properties
about them may be a hard task without proper tool support. There are some
dedicated provers for modal logics~\cite{Gleissner17,Mora11}, but we believe
that using mature tools, like a proof assistant, are a better option, for
verifying facts using modal logics. We justify this choice by: 1) the recent
successful histories on verification of software and
mathematics~\cite{Leroy09,Klein10,Gonthier13,Gonthier08}; 2) There is a growing
community of developers using such tools thanks to 
textbooks on the subject~\cite{Nipkow14,Bertot10,Chlipala13} which allows
newbies to join the game of formal verification and 3) The custom automation
facilities that current proof assistants support allows that some trivial facts can
be checked using decision procedures.

In this context, the present work aims to contribute by constructing a library
that formalize a family of modal logics and its related properties in Coq.
By using such a library, a developer could use all Coq's resources to prove
modal logics facts.

More specifically, we contribute:

\begin{itemize}
  \item We define a \emph{deep embedding} of modal logic syntax in Coq and
    define its Kripke semantics.
  \item We model the Hilbert deductive systems K, D, B, T, K4, K5, S4 and S5
    in Coq and prove its soundness with respect to Kripke semantics.
  \item We show a proof of the L\"ob's theorem using the Hilbert deductive system
    developed in our library.
\end{itemize}

The rest of this work is organized as follows: Section~\ref{sec:background} provides
brief introduction on modal logics and the Coq proof assistant. Details about
the implementation of our Coq library are presented in
Section~\ref{sec:implementation}. A case study in proving the L\"ob's theorem is
the subject of Section~\ref{sec:case}. Related work is discussed
on Section~\ref{sec:related} and Section~\ref{sec:conclusion} draws some conclusions and
provide some pointers for future works.

\section{Background}\label{sec:background}

\paragraph{An overview of Coq proof assistant.}

Coq is a proof assistant based on the calculus of inductive
constructions (CIC)~\cite{manual_cic}, a higher order typed
$\lambda$-calculus extended with inductive definitions.  Theorem
proving in Coq follows the ideas of the so-called
``BHK-cor\-res\-pon\-dence''\footnote{Abbreviation of Brouwer, Heyting,
  Kolmogorov, de Bruijn and Martin-L\"of Correspondence. This is also
  known as the Curry-Howard ``isomorphism''.}, where types represent
logical formulas, $\lambda$-terms represent proofs
\cite{Bertot10} and the task of checking if a piece of text is a
proof of a given formula corresponds to checking if the term that
represents the proof has the type corresponding to the given formula.

However, writing a proof term whose type is that of a logical formula
can be a hard task, even for very simple propositions.  In order to
make the writing of complex proofs easier, Coq provides
\emph{tactics}, which are commands that can be used to construct proof
terms in a more user friendly way.

As a tiny example, consider the task of proving the following simple
formula of propositional logic:
\[
(A \to B)\to (B\to C) \to A \to C
\]
In Coq, such theorem can be expressed as:
\begin{minted}{coq}
Section EXAMPLE.
   Variables A B C : Prop.
   Theorem example : (A -> B) -> (B -> C) -> A -> C.
   Proof.
       intros H H' HA. apply H'. apply H. assumption. 
   Qed.
End EXAMPLE.
\end{minted}
In the previous source code piece, we have defined a Coq section named
\mintinline{coq}{EXAMPLE}\footnote{In Coq, we can use sections to delimit the
  scope of local variables.} which declares variables \mintinline{coq}{A},
\mintinline{coq}{B} and \mintinline{coq}{C} as being propositions (i.e. with type
\mintinline{coq}{Prop}). Tactic \mintinline{coq}{intros} introduces variables
\mintinline{coq}{H}, \mintinline{coq}{H'} and \mintinline{coq}{HA} into the (typing) context,
respectively with types \mintinline{coq}{A -> B}, \mintinline{coq}{B -> C} and
\mintinline{coq}{A} and leaves goal \mintinline{coq}{C} to be proved. Tactic
\mintinline{coq}{apply}, used with a term \mintinline{coq}{t}, generates goal
\mintinline{coq}{P}
when there exists \mintinline{coq}{t: P -> Q} in the typing context and the
current goal is \mintinline{coq}{Q}. Thus, \mintinline{coq}{apply H'} changes the goal
from \mintinline{coq}{C} to \mintinline{coq}{B} and \mintinline{coq}{apply H}
changes the goal to \mintinline{coq}{A}. Tactic \mintinline{coq}{assumption}
traverses the typing context to find a hypothesis that matches with the goal.

We define next a proof of the previous propositional logical formula
that, in contrast to the previous proof, that was built using tactics
(\mintinline{coq}{intros}, \mintinline{coq}{apply} and \mintinline{coq}{assumption}), is coded
directly as a function:
\begin{minted}{coq}
Definition example :(A -> B) -> (B -> C) -> A -> C :=
 fun (H : A -> B)
     (H' : B -> C)
     (HA : A) => H' (H HA).
\end{minted}
However, even for very simple theorems, coding a definition directly
as a Coq term can be a hard task. Because of this, the use of tactics
has become the standard way of proving theorems in Coq. Furthermore,
the Coq proof assistant provides not only a great number of tactics
but also a domain specific language for scripted proof automation,
called $\mathcal{L}$tac. Details about $\mathcal{L}$tac and Coq can be found
in~\cite{Chlipala13,Bertot10,manual_coq}.

\paragraph{An overview of modal logic.} Modal logics are designed for reasoning
about truth across various --- abstract --- worlds. In such logics, a
proposition may be true in some world, but false in another one. The versions of
modal logics considered in this work extends traditional propositional logics
with two operators on propositions: $\Box$ (box) and $\Diamond$ (diamond).
The syntax of modal logic formulas are defined by the following context free
grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$, $\mathcal{P}$ is
the enumerable set of propositional variables and $A\in\mathcal{P}$ is arbitrary):
\[
  \begin{array}{lcl}
    \varphi & ::= & \bot\,|\,\top\,|\,A\,|\,\neg\varphi\,|\,\varphi\circ\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
  \end{array}  
\]
Intuitively, the semantics of the operator $\Box$ is similar to an universal
quantifier, since it express \emph{necessity}: $\Box\,\varphi$ is true at the
current world iff $\varphi$ is true at all worlds. Similarly, the semantics of operator
$\Diamond$ is that a formula $\Diamond\varphi$ is true at the current
world iff $\varphi$ is \emph{possible}, i.e. true in at least some world.

Formally, we can only interpret a formula with respect to a \emph{model}, which is
defined as follows.
A \emph{frame} $\mathcal{F} = \langle W, R \rangle$ consists of a pair formed by
a non-empty set of worlds, $W$, and a binary accessibility relation between
worlds, $R \subseteq W \times W$. A pair $w_1Rw_2$ denotes that world $w_2$ is
accessible from $w_1$. A model $\mathcal{M} = \langle \mathcal{F}, \mathcal{V} \rangle$
of a frame $\mathcal{F}$ and a total labeling function
$\mathcal{V} : \mathcal{P}\times W\to \{F,T\}$, which assigns
a truth value to a propositional variable at a given world, i.e. when $w \in \mathcal{V}(A)$
for some $A \in\mathcal{P}$, we say that $A$ holds in $w$. The satisfiability of a formula
in a given world is defined by inductive relation $\mathcal{M}\,;\,w\models \varphi$ as
follows (notation $\mathcal{M}\,;\,w\not\models \varphi$ denotes
that $\mathcal{M}\,;\,w\models \varphi$ does not hold):

\begin{enumerate}
  \item $\mathcal{M}\,;\,w\not\models \bot$ and $\mathcal{M}\,;\,w\models \top$;
  \item $\mathcal{M}\,;\,w\models A$ iff $\mathcal{V}(A,w) = T$;
  \item $\mathcal{M}\,;\,w\models \neg\varphi$ iff $\mathcal{M}\,;\,w\not\models\varphi$;
  \item $\mathcal{M}\,;\,w\models \varphi_1 \land \varphi_2$ iff
    $\mathcal{M}\,;\,w\models\varphi_1$ and $\mathcal{M}\,;\,w\models\varphi_2$;
  \item $\mathcal{M}\,;\,w\models \varphi_1 \lor \varphi_2$ iff
    $\mathcal{M}\,;\,w\models\varphi_1$ or $\mathcal{M}\,;\,w\models\varphi_2$;
  \item $\mathcal{M}\,;\,w\models \varphi_1 \to \varphi_2$ iff
    $\mathcal{M}\,;\,w\not\models\varphi_1$ or $\mathcal{M}\,;\,w\models\varphi_2$;
  \item $\mathcal{M}\,;\,w\models\Box\varphi$ iff $\forall w \in W. w R y \to \mathcal{M}\,;\,y \models\varphi$;
  \item $\mathcal{M}\,;\,w\models\Diamond\varphi$ iff $\exists w \in W. w R y \land \mathcal{M}\,;\,y \models\varphi$;
\end{enumerate}

We say that a formula $\varphi$ is satisfiable in a model $\mathcal{M}$,
denoted by $\mathcal{M}\models \varphi$, if $\forall w \in W. M\,;w\,\models \varphi$ holds.
Similarly, $\varphi$ is valid in a frame $\mathcal{F}$, written $\mathcal{F}\models\varphi$, if
it holds on all models $\mathcal{M} = \langle \mathcal{F},\mathcal{V}\rangle$, i.e.
for any $\mathcal{M}$, we have that $\mathcal{M}\models\varphi$. Finally, we say that a formula
is valid, written $\models \varphi$, if it is valid in any frame $\mathcal{F}$.

\section{Embedding Modal Logic in Coq}\label{sec:implementation}

In this section, we describe the implementation details of our modal logic
library in Coq. We start by describing the structure of formula syntax and the
representation of frames and models, needed to define the Kripke semantics.

\paragraph{Syntax of modal logics formulas}
In order to implement our modal logic library, we use a \emph{deep embedding} of
formulas, i.e. we represent the syntax as the following Coq inductive datatype.
\begin{minted}{coq}
Inductive modalFormula : Set :=
| Lit     : nat          -> modalFormula
| Neg     : modalFormula -> modalFormula
| Box     : modalFormula -> modalFormula
| Dia     : modalFormula -> modalFormula
| And     : modalFormula -> modalFormula -> modalFormula
| Or      : modalFormula -> modalFormula -> modalFormula
| Implies : modalFormula -> modalFormula -> modalFormula.
\end{minted}
Most constructors of the type have an immediate meaning, since they represent
connectives of propositional logics. The constructor \mintinline{coq}{Lit}
is used to construct propositional variables; constructor \mintinline{coq}{Box}
denotes the modal operator $\Box$ and \mintinline{coq}{Dia}, the operator
$\Diamond$.

Before defining the Kripke semantics, we need to represent the model used to
interpret a modal logic formula. We encode a frame by a record formed by a
a type \mintinline{coq}{W : Set} to represent the set of worlds and the
accessibility relation \mintinline{coq}{R : W -> W -> Prop}.
\begin{minted}{coq}
Record Frame : Type := {
  W : Set;
  R : W -> W -> Prop;
}.
\end{minted}
A model is denoted by another record formed by a frame and a labelling function,
which assign the truth value of a variable at a given world.
\begin{minted}{coq}
Record Model : Type := {
  F : Frame; 
  v : nat -> (W F) -> Prop; 
}.
\end{minted}
Note that the type of the labelling function \mintinline{coq}{v} ensures that it
refers to a world from the set of possible worlds in the frame \mintinline{coq}{F}.

\section{Case Study}\label{sec:case}

\section{Related work}\label{sec:related}

\section{Conclusion}\label{sec:conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}

### 2022-05-06 paper/paper.tex
More updates on text. Some litte progress on lean formalization. Updating gitignore

@@
 \usepackage{minted}
+\usepackage{bussproof}
 
@@
 newbies to join the game of formal verification and 3) The custom automation
-facilities that current proof assistants support allows that some trivial facts can
-be checked using decision procedures.
+facilities that current proof assistants support, allows that trivial facts can
+be proved using decision procedures, relieving the user from writing proof
+scripts for formulas that can be checked automatically.
 
@@
 The syntax of modal logic formulas are defined by the following context free
-grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$, $\mathcal{P}$ is
-the enumerable set of propositional variables and $A\in\mathcal{P}$ is arbitrary):
+grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$, $\mathcal{P} = \{A,B,C,...\}$ is
+the enumerable set of propositional variables):
 \[
@@
 
+\begin{figure}[H]
 \begin{enumerate}
@@
   \end{enumerate}
-
+  \caption{Kripke semantics}
+  \label{fig:kripke}
+\end{figure}
 We say that a formula $\varphi$ is satisfiable in a model $\mathcal{M}$,
@@
 it holds on all models $\mathcal{M} = \langle \mathcal{F},\mathcal{V}\rangle$, i.e.
-for any $\mathcal{M}$, we have that $\mathcal{M}\models\varphi$. Finally, we say that a formula
+for any $\mathcal{M}$, we have that $\mathcal{M}\models\varphi$. A formula
 is valid, written $\models \varphi$, if it is valid in any frame $\mathcal{F}$.
+
+Using the semantics, we can define the concept of \emph{entailment}: Given a set
+of formulas $\Gamma$ and a formula $\varphi$, we say that $\Gamma$ entails
+$\varphi$, written $\Gamma\models\varphi$, iff $\forall \varphi' \in\Gamma .
+\mathcal{M} \models \varphi' \to \mathcal{M} \models \varphi$, for some model $\mathcal{M}$.
+
+In this work, we consider Hilbert deductive systems for modal logics.
+Essentially, the considered system adds two axioms for the modalities
+to the set of axioms for classical propositional logic.
+The included axioms express that $\Box$ distributes over implication and
+$\Diamond$ over disjunction.
+\[
+  \begin{array}{ll}
+    \textbf{K}           & \Box(A \to B) \to (\Box A \to \Box B)\\
+    \textbf{Possibility} & \Diamond (A \lor B) \to (\Diamond A \lor \Diamond B)\\
+  \end{array}
+\]
+Apart from the axioms, we also need the traditional \emph{modus ponens} rule
+and a rule that allows to conclude formulas with necessity.
+
+\begin{figure}[H]
+\begin{tabular}{cc}
+    \AxiomC{$\Gamma\vdash\varphi_1\to\varphi_2$}
+    \AxiomC{$\Gamma\vdash\varphi_2$}
+    \RightLabel{MP}
+    \BinaryInfC{$\Gamma\vdash\varphi_2$}
+    \DisplayProof
+  &
+    \AxiomC{$\Gamma\vdash\varphi$}
+    \RightLabel{Nec}
+    \UnaryInfC{$\Gamma\vdash\Box \varphi$}
+    \DisplayProof
+\end{tabular}
+\end{figure}
+The set of axioms and the two previous rules forms a sound and complete Hilbert-style
+deductive system for modal logics~\cite{blackburn}.
 
@@
 
-\paragraph{Syntax of modal logics formulas}
-In order to implement our modal logic library, we use a \emph{deep embedding} of
-formulas, i.e. we represent the syntax as the following Coq inductive datatype.
+\paragraph{Syntax and semantics of modal logics}
+In order to implement our library, we use a \emph{deep embedding} of
+formulas, i.e. we represent formulas as the following Coq inductive datatype.
 \begin{minted}{coq}
-Inductive modalFormula : Set :=
-| Lit     : nat          -> modalFormula
-| Neg     : modalFormula -> modalFormula
-| Box     : modalFormula -> modalFormula
-| Dia     : modalFormula -> modalFormula
-| And     : modalFormula -> modalFormula -> modalFormula
-| Or      : modalFormula -> modalFormula -> modalFormula
-| Implies : modalFormula -> modalFormula -> modalFormula.
+Inductive form : Set :=
+| Lit     : nat          -> form
+| Neg     : form -> form
+| Box     : form -> form
+| Dia     : form -> form
+| And     : form -> form -> form
+| Or      : form -> form -> form
+| Implies : form -> form -> form.
 \end{minted}
@@
 a type \mintinline{coq}{W : Set} to represent the set of worlds and the
-accessibility relation \mintinline{coq}{R : W -> W -> Prop}.
+accessibility relation is denoted by a type \mintinline{coq}{R : W -> W -> Prop}.
 \begin{minted}{coq}
@@
 Note that the type of the labelling function \mintinline{coq}{v} ensures that it
-refers to a world from the set of possible worlds in the frame \mintinline{coq}{F}.
+refers to a world from the set of possible worlds in the frame
+\mintinline{coq}{F}. The encoding of the Kripke semantics is done by a Coq function
+that which implements the relation of Figure~\ref{fig:kripke}. We omit the code
+for propositional logics connectives for brevity. The interesting parts are for
+variables (constructor \mintinline{coq}{Lit}) which uses the labelling function
+\mintinline{coq}{v} from the model \mintinline{coq}{M}. The meaning of
+modalities $\Box$ and $\Diamond$ are represented by a quantifying over worlds
+and use the accessibility relation $R$. 
+\begin{minted}{coq}
+Fixpoint formula_valuation (M: Model)
+                           (w: W (F M))
+                           (phi: form): Prop :=
+  match phi with
+  | Lit     x            => v M x w 
+  | Box     psi
+     => forall w': W (F M), R (F M) w w' ->
+                            formula_valuation M w' psi
+  | Dia     psi
+     => exists w': W (F M), R (F M) w w' /\
+                   formula_valuation M w' psi
+  (** some code omitted* )
+  end.
+\end{minted}
+In order to ease the task of writing formulas and its semantics properties, we
+define Coq notations that mimics the standard way of writing them.
+
+\paragraph{Systems for modal logic}
+
+\paragraph{The Hilbert deductive system}
+Basically, a Hilbert-style deductive system allows the use of on four
+possibilities in a proof: 1) the use of a hypothesis; 2) use of an axiom;
+3) use of the modus ponens rule or 4) the use of the generalization (necessity)
+rule. The inductive type \mintinline{coq}{deduction} encodes these possibilities
+as its constructors. The \mintinline{coq}{deduction} type is parameterized by
+axioms that form a system for modal logics. In this way, our definition can be
+used in several systems of modal logic.
+\begin{minted}{coq}
+Inductive deduction (A: axiom -> Prop)
+                           : theory -> form -> Prop :=
+  (* Premise. *)
+  | Prem: forall (t: theory)
+                 (f: form)
+                 (i: nat),
+          (nth_error t i = Some f) -> deduction A t f
+  (* Axiom. *)
+  | Ax: forall (t: theory)
+               (a: axiom)
+               (f: form),
+        A a -> instantiate a = f -> deduction A t f
+  (* Modus Ponens. *)
+  | Mp: forall (t: theory)
+               (f g: form)
+               (d1: deduction A t (f -> g))
+               (d2: deduction A t f),
+        deduction A t g
+  (* Generalization. *)
+  | Nec: forall (t: theory)
+                (f: form)
+                (d1: deduction A t f),
+         deduction A t (Box f).
+\end{minted}
+The constructors \mintinline{coq}{Prem}, \mintinline{coq}{Mp} and \mintinline{coq}{Nec}
+represent rules for hypothesis, modus ponens and generalization, respectively.
+The \mintinline{coq}{Ax} constructor deserves a bit of explanation. Function
+\mintinline{coq}{instantiate} build a formula from a given axiom set for a given
+system represented by a theory, which parameterizes our deduction type.
+\begin{minted}{coq}
+Definition instantiate (a: axiom): modalFormula :=
+  match a with
+  | ax1    p   q
+     => [! p -> (q -> p) !]
+  | ax2    p   q   r
+     => [! (p -> (q -> r)) -> ((p -> q) -> (p -> r)) !]
+  | ax3    p   q
+     => [! (~ q -> ~ p) -> (p -> q) !]
+  | axK    p   q
+     => [! [] (p -> q) -> ([] p -> [] q) !]
+  | axPos  p   q
+     => [! <> (p \/ q) -> (<> p \/ <> q) !]
+  (** some code omitted *)
+  end.
+\end{minted}
+Notation \mintinline{coq}{[! p !]} is used to write modal logic formulas using a
+syntax closer to mathematical practice using Coq. This is necessary to avoid
+ambiguities, since some symbols are also used by Coq's syntax.
+Modalities $\Box$ and $\Diamond$ are denoted by \mintinline{coq}{[]}
+and \mintinline{coq}{<>}, respectively.
 
@@
 \section{Related work}\label{sec:related}
+
+\paragraph{Formalizations of modal logics on proof assistants}
+
+The formalization of epistemic modal logic in Coq was the subject of Wind's
+work~\cite{dewind}.
+The author made a Coq implementation of modal logic based on the S5 system,
+which consists of  the axioms $K$, $T$, $4$ and $5$. The objective was the
+modelling of two well known puzzles.
+Wind's formalization was separated into some components. The first consists in
+defining the syntax of formulas and the Kripke semantics. Next, the author uses
+their definitions to give some examples of modal operators, worlds and how they
+interact with each other. Using the built machinery, the author developed
+introduction and elimination rules for modal logic and use them to model the
+wize man's and muddy children puzzles.
 

### 2022-05-31 paper/paper.tex
Fixing an error on rules
Fixing some typos

@@
 \author{Ariel Agne da Silveira}
+\author{Miguel}
 \author{Paulo Torrens}
@@
 error detection and that its corresponding fix will ensure that the software
-sill continue to work under abnormal conditions. One way to express such
+still continue to work under abnormal conditions. One way to express such
 requirements is using \emph{modal logics}, which can specify communication
@@
 dedicated provers for modal logics~\cite{Gleissner17,Mora11}, but we believe
-that using mature tools, like a proof assistant, are a better option, for
+that using mature tools, like a proof assistant, are a better option for
 verifying facts using modal logics. We justify this choice by: 1) the recent
@@
 newbies to join the game of formal verification and 3) The custom automation
-facilities that current proof assistants support, allows that trivial facts can
+facilities supported by current proof assistants allows that trivial facts can
 be proved using decision procedures, relieving the user from writing proof
@@
     define its Kripke semantics.
-  \item We model the Hilbert deductive systems K, D, B, T, K4, K5, S4 and S5
-    in Coq and prove its soundness with respect to Kripke semantics.
+  \item We model the Hilbert deductive systems as a Coq inductive type that is
+    parameterized by different axiom systems like K, D, B, T, K4, K5, S4 and S5
+    and prove its soundness with respect to Kripke semantics. We also
+    discuss the current status of the formalization of the completeness result.
   \item We show a proof of the L\"ob's theorem using the Hilbert deductive system
@@
 changes the goal to \mintinline{coq}{A}. Tactic \mintinline{coq}{assumption}
-traverses the typing context to find a hypothesis that matches with the goal.
+traverses the typing context to find a hypothesis that matches the goal.
 
@@
 The syntax of modal logic formulas are defined by the following context free
-grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$, $\mathcal{P} = \{A,B,C,...\}$ is
-the enumerable set of propositional variables):
+grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$ and
+$\mathcal{P} = \{A,B,C,...\}$ is the enumerable set of propositional variables):
 \[
@@
 accessible from $w_1$. A model $\mathcal{M} = \langle \mathcal{F}, \mathcal{V} \rangle$
-of a frame $\mathcal{F}$ and a total labeling function
+is formed by a frame $\mathcal{F}$ and a total labeling function
 $\mathcal{V} : \mathcal{P}\times W\to \{F,T\}$, which assigns
@@
 In this work, we consider Hilbert deductive systems for modal logics.
-Essentially, the considered system adds two axioms for the modalities
+Essentially, the considered system adds axioms for the modalities
 to the set of axioms for classical propositional logic.
-The included axioms express that $\Box$ distributes over implication and
-$\Diamond$ over disjunction.
+The most basic axiom is \textbf{K} which express that $\Box$ distributes
+over implication. Axiom of possibility states that $\Diamond$ distributes
+over disjunction.
 \[
@@
     \AxiomC{$\Gamma\vdash\varphi_1\to\varphi_2$}
-    \AxiomC{$\Gamma\vdash\varphi_2$}
+    \AxiomC{$\Gamma\vdash\varphi_1$}
     \RightLabel{MP}
@@
 
+A central aspect in the study of modal logics is the so-called \emph{frame
+  correspondence}, which is a relationship between the
+accessibility relation and the truth of certain formulas in frames containing
+it. As an example, consider a frame $\mathcal{F} = \langle W,R
+\rangle$ in which the relation $R$ is reflexive: $\forall w \in W. wRw$, i.e.
+every world is accessible from itself. In such frame when $\Box A$ holds at
+a world $w$, $w$ itself is among the accessible worlds at which $A$ must be true.
+In this way, if the accessibility relation $R$ is reflexive, then whatever world
+$w$ and formula $\varphi$ we take, the formula $\Box \varphi \to \varphi$ will
+be true there. Example of properties that characterize certain formula schemas are
+presented in the table below.
+
+\begin{figure}[H]
+\begin{tabular}{|c|c|}
+  \hline
+  Property of $R$                          & Formula schema \\ \hline
+  reflexive: $\forall w.wRw$               & $\Box A \to A$ \\ 
+  serial: $\forall w\exists w'. wRw'$      & $\Box A \to \Diamond A$ \\
+  symmetric: $\forall w\,w'.wRw' \to w'Rw$ & $A \to \Box\Diamond A$ \\
+  \hline
+\end{tabular}
+\end{figure}
+There are other properties of interest for a relation $R$ like being
+transitive, functional, euclidean, and so on. More about the correspondence
+of frames and deductive systems properties can be found elsewhere~\cite{Chellas,blackburn}.
+
 \section{Embedding Modal Logic in Coq}\label{sec:implementation}
@@
 representation of frames and models, needed to define the Kripke semantics.
+Next, we define a modular Hilbert-style deductive system. The modularization is
+obtained by parameterizing it by a set of axioms used. This allow us to
+generalize results about derivations to several deductive systems, independently
+of the axioms used. We conclude by discussing the soundness, the current
+status of the completeness and the frame correspondence proofs.
 
@@
 modalities $\Box$ and $\Diamond$ are represented by a quantifying over worlds
-and use the accessibility relation $R$. 
+using the accessibility relation $R$. 
 \begin{minted}{coq}
@@
                    formula_valuation M w' psi
-  (** some code omitted* )
+  (** some code omitted *)
   end.
@@
 define Coq notations that mimics the standard way of writing them.
-
-\paragraph{Systems for modal logic}
 
@@
 axioms that form a system for modal logics. In this way, our definition can be
-used in several systems of modal logic.
+used by several systems of modal logic.
 \begin{minted}{coq}
@@
 represent rules for hypothesis, modus ponens and generalization, respectively.
-The \mintinline{coq}{Ax} constructor deserves a bit of explanation. Function
-\mintinline{coq}{instantiate} build a formula from a given axiom set for a given
-system represented by a theory, which parameterizes our deduction type.
+Before discuss the rule for axioms, we need to discuss how axioms are
+represented in our library. The inductive type \mintinline{coq}{axiom} specifies
+the different axioms supported by our library.
+\begin{minted}{coq}
+Inductive axiom : Set :=
+  | ax1   : modalFormula -> modalFormula -> axiom
+  | ax2   : modalFormula -> modalFormula ->
+                            modalFormula -> axiom
+  (** some code omitted *)
+  | axK   : modalFormula -> modalFormula -> axiom
+  | axPos : modalFormula -> modalFormula -> axiom
+  | axT   : modalFormula -> axiom
+  | axB   : modalFormula -> axiom
+  | axK4  : modalFormula -> axiom
+  | axD   : modalFormula -> axiom
+  | axK5  : modalFormula -> axiom.
+\end{minted}
+Each constructor of type \mintinline{coq}{axiom} represents an axiom and
+its type stores the subformulas which compose the axiom. The type has
+17 constructors: the first 10 consists encode axioms for propositional logics
+and the last 7 are for different formulas taken as axioms by systems of modal logics.
+Axioms are built using the function \mintinline{coq}{instantiate}, which
+receives an axiom and build its corresponding formula by pattern matching as
+follows:
 \begin{minted}{coq}
@@
 syntax closer to mathematical practice using Coq. This is necessary to avoid
-ambiguities, since some symbols are also used by Coq's syntax.
+parsing ambiguities, since some symbols are also used by Coq's syntax.
 Modalities $\Box$ and $\Diamond$ are denoted by \mintinline{coq}{[]}
 and \mintinline{coq}{<>}, respectively.
+
+The \mintinline{coq}{Ax} constructor uses the \mintinline{coq}{instantiate} to
+create a formula instance from using a given system, represented as a predicate
+\mintinline{coq}{A : axiom -> Prop}, which parameterizes our deduction type. An
+axiom \mintinline{coq}{a} is valid within a system \mintinline{coq}{A} whenever
+the type \mintinline{coq}{A a} is inhabited, i.e. it is not empty. Next, we
+detail the representation of different systems of modal logic in our library.
+
+\paragraph{Systems of modal logics} In order to use the deductive system to
+check the provability of formulas, we need to specify the axioms that are
+included in the system. Our approach is we represent to define predicates
+which denote each system. The simplest set of axioms is usually refereed as
+\textbf{K} and it includes the axiom $\Box(A \to B) \to (\Box A \to \Box B)$
+and all the axioms for propositional logic. We represent system \textbf{K}
+by the following inductive type:
+\begin{minted}{coq}
+Inductive K: axiom -> Prop :=
+  | K_ax1: forall p q, K (ax1 p q)
+  | K_ax2: forall p q r, K (ax2 p q r)
+  (** some code omitted *)
+  | K_axK: forall p q, K (axK p q)
+\end{minted}
+
+\paragraph{Properties}
 

### 2022-06-18 paper/paper.tex

@@
 \usepackage{bussproof}
+\usepackage{multirow}
 
@@
 to the set of axioms for classical propositional logic.
-The most basic axiom is \textbf{K} which express that $\Box$ distributes
-over implication. Axiom of possibility states that $\Diamond$ distributes
+As examples of modalities axioms we mention \textbf{K}, which express that $\Box$ distributes
+over implication, and possibility which states that $\Diamond$ distributes
 over disjunction.
@@
 \end{minted}
+Type \mintinline{coq}{K} is a predicate over axioms: it has one constructor for
+each of the valid axioms which form the system. Given a value
+of \mintinline{coq}{axiom}, a value of type \mintinline{coq}{K a} denotes a
+proof that a given axiom is part of the system \textbf{K}. This is the main
+trick used by the constructor \mintinline{coq}{Ax}, of type
+\mintinline{coq}{deduction}, to generically represent axioms for
+different systems of modal logics uniformly. In order to understand how it
+works, let us review its type (below, remeber that \mintinline{coq}{A : axiom -> Prop}
+denotes a system of axioms for modal logics):
+\begin{minted}{coq}
+Ax: forall (t: theory)
+           (a: axiom)
+           (f: form),
+           A a -> instantiate a = f -> deduction A t f
+\end{minted}
+Notice that the first parameter of constructor \mintinline{coq}{Ax} is a value
+of type \mintinline{coq}{A a}, which is only inhabited if the predicate
+\mintinline{coq}{A} holds for the axiom \mintinline{coq}{a}. Through this
+evidence, is possible to create the corresponding axiom formula instance by
+using function \mintinline{coq}{instantiate}. In this way, we are able to
+represent all systems of axioms for modal logics using a single constructor
+for its axioms.
 
@@
 Wind's formalization was separated into some components. The first consists in
-defining the syntax of formulas and the Kripke semantics. Next, the author uses
+defining the syntax of formulas and its Kripke semantics. Next, the author uses
 their definitions to give some examples of modal operators, worlds and how they
 interact with each other. Using the built machinery, the author developed
-introduction and elimination rules for modal logic and use them to model the
-wize man's and muddy children puzzles.
+introduction and elimination rules for modal logic operators and use them to
+model the wize man's and muddy children puzzles.
+
+A Coq formalization for modal systems $K$, $T$ and $K4$ was the subject of 
+Doczkal and Smolka's work. Author's interest was in the definition of dynamic
+and temporal logics. The development of the code establishes the construction
+and definitions of tactics, through lemmas propositional operators and relation
+properties were constructed. Each model was defined through structures, which
+consist of worlds, relationships and propositions~\cite{saarland}.
+
+Another proposal of formalizing modal logics was developed by Benzmüller and
+Paleo~\cite{Chris_Bruno}. Author's main objective is checking if current
+type theory-based proof assistants are suitable for modal logic reasoning.
+They defined a formalization of systems $K$ and $S5$ and built custom tactics
+which allows proofs using a natural deduction style.
+The authors argue that current proof assistants are a useful technology for
+reasoning about facts expressed using modal logics.
+
+\vspace{0.5cm} {
+  \begin{table}[H]
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+      \hline
+      \multirow{2}{*}{Authors} & \multicolumn{8}{c|}{Systems considered}         \\ \cline{2-9} 
+                               & K & D & B & T & K4 & K5 & S4 & S5 \\ \hline
+      Wind et. al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X \\ \hline
+      Doczkal et. al.~\cite{saarland}   & X &   &   & X & X  &    &    &    \\ \hline
+      Benzmüller et. al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X  \\ \hline
+      This work         & X & X & X & X & X  &  X &  X & X  \\ \hline
+    \end{tabular}
+    \centering
+    \caption{Comparision with different Coq modal logics developements.}
+    \label{tab:comparacao_trabalhos_rel}
+  \end{table}
+}
 

### 2022-06-19 paper/paper.tex
More writing

@@
   \item We model the Hilbert deductive systems as a Coq inductive type that is
-    parameterized by different axiom systems like K, D, B, T, K4, K5, S4 and S5
+    parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL
     and prove its soundness with respect to Kripke semantics. We also
@@
   \begin{array}{lcl}
-    \varphi & ::= & \bot\,|\,\top\,|\,A\,|\,\neg\varphi\,|\,\varphi\circ\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
+    \varphi & ::= & A\,|\,\neg\varphi\,|\,\varphi\circ\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
   \end{array}  
@@
 quantifier, since it express \emph{necessity}: $\Box\,\varphi$ is true at the
-current world iff $\varphi$ is true at all worlds. Similarly, the semantics of operator
+current world iff $\varphi$ is true at all worlds accessible from the current
+world. Similarly, the semantics of operator
 $\Diamond$ is that a formula $\Diamond\varphi$ is true at the current
-world iff $\varphi$ is \emph{possible}, i.e. true in at least some world.
+world iff $\varphi$ is \emph{possible}, i.e. true in at least some world
+accessible from the current world.
 
@@
     $\mathcal{M}\,;\,w\not\models\varphi_1$ or $\mathcal{M}\,;\,w\models\varphi_2$;
-  \item $\mathcal{M}\,;\,w\models\Box\varphi$ iff $\forall w \in W. w R y \to \mathcal{M}\,;\,y \models\varphi$;
-  \item $\mathcal{M}\,;\,w\models\Diamond\varphi$ iff $\exists w \in W. w R y \land \mathcal{M}\,;\,y \models\varphi$;
+  \item $\mathcal{M}\,;\,w\models\Box\varphi$ iff $\forall y \in W. w R y \to \mathcal{M}\,;\,y \models\varphi$;
+  \item $\mathcal{M}\,;\,w\models\Diamond\varphi$ iff $\exists y \in W. w R y \land \mathcal{M}\,;\,y \models\varphi$;
   \end{enumerate}
@@
   \begin{array}{ll}
-    \textbf{K}           & \Box(A \to B) \to (\Box A \to \Box B)\\
-    \textbf{Possibility} & \Diamond (A \lor B) \to (\Diamond A \lor \Diamond B)\\
+    \textbf{K} & \Box(\varphi_1 \to \varphi_2) \to (\Box \varphi_1 \to \Box \varphi_2)\\
+    \textbf{Possibility} & \Diamond (\varphi_1 \lor \varphi_2) \to
+                           (\Diamond \varphi_2 \lor \Diamond \varphi_2)\\
   \end{array}
@@
 Next, we define a modular Hilbert-style deductive system. The modularization is
-obtained by parameterizing it by a set of axioms used. This allow us to
+obtained by parameterizing it by a set of axioms used. This allows us to
 generalize results about derivations to several deductive systems, independently
@@
 check the provability of formulas, we need to specify the axioms that are
-included in the system. Our approach is we represent to define predicates
+included in the system. Our approach is to define predicates
 which denote each system. The simplest set of axioms is usually refereed as
@@
 
-\paragraph{Properties}
+\paragraph{Properties} A key property of the Hilbert deductive system formalized
+in our library is its soundness. If a formula $\varphi$ is derived from a
+context $\Gamma$, i.e. $\Gamma\vdash\varphi$, then $\varphi$ is entailed from
+the set $\Gamma$, i.e. $\Gamma\models\varphi$. The Coq statement of the soundness
+result is:
+
+\begin{minted}{coq}
+Theorem soundness:
+  forall G p, (K ; G |-- p) -> (G ||= p).
+\end{minted}
+
+The proof proceeds by induction on the derivation of $\Gamma \vdash\varphi$ and
+it has a case for each of the 11 axioms of the Hilbert system for propositional
+logics, one for the $K$ axiom and cases for using hypothesis and the rules modus
+ponens. In order to modularize the proof, we defined a lemma for each case.
+As example, the following is the statement for modus ponens case.
+\begin{minted}{coq}
+Lemma Modus_Ponens_soundness:
+  forall M w p q,
+  ((M ' w ||- p) /\
+  (M ' w ||- [! p -> q !])) ->
+  (M ' w ||- q).
+\end{minted}
+All these lemmas are direct consequences from the definitions of the Kripke
+semantics and entailment.
+
+While we have completely proved the soundness theorem, the current version of
+our library does not have a completeness theorem.
 
 \section{Case Study}\label{sec:case}
+
+In this section we describe a case study of the defined modal logic library.
+We proved the L\"ob theorem in the system K4 with modal fixed points.
+The importance of L\"ob's theorem is due to its use in provability
+logic~\cite{boolos_1994} to abstract details from encodings in G\"odel's
+incomleteness results.
 
@@
   \begin{table}[H]
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
       \hline
-      \multirow{2}{*}{Authors} & \multicolumn{8}{c|}{Systems considered}         \\ \cline{2-9} 
-                               & K & D & B & T & K4 & K5 & S4 & S5 \\ \hline
-      Wind et. al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X \\ \hline
-      Doczkal et. al.~\cite{saarland}   & X &   &   & X & X  &    &    &    \\ \hline
-      Benzmüller et. al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X  \\ \hline
-      This work         & X & X & X & X & X  &  X &  X & X  \\ \hline
+      \multirow{2}{*}{Authors} & \multicolumn{9}{c|}{Systems considered} \\ \cline{2-10} 
+                               & K & D & B & T & K4 & K5 & S4 & S5 & GL\\ \hline
+      Wind et. al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X & \\ \hline
+      Doczkal et. al.~\cite{saarland}   & X &   &   & X & X  &    &    &  &  \\ \hline
+      Benzmüller et. al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X &  \\ \hline
+      This work         & X & X & X & X & X  &  X &  X & X  & X \\ \hline
     \end{tabular}

### 2022-06-19 paper/paper.tex
Adding abstract

@@
 \begin{abstract}
-  The correct specification of requirements is crucial for software development.
-  Modal logics are ...
+  This work describes an encoding of modal logics using the Coq proof assistant.
+  Our formalization differs from previous attempts by providing a uniform
+  representation of several systems for modal logics, using Coq's rich type
+  structure. We illustrate the usefulness of our library in a formalization of
+  L\"ob's theorem, which closely follows a classical proof of this result.
 \end{abstract}

### 2022-06-20 paper/paper.tex
Finished section on Lobs theorem. Little changes on proof script.

--- 2022-06-19 paper/paper.tex
@@
 \begin{abstract}
-  The correct specification of requirements is crucial for software development.
-  Modal logics are ...
+  This work describes an encoding of modal logics using the Coq proof assistant.
+  Our formalization differs from previous attempts by providing a uniform
+  representation of several systems for modal logics using Coq's rich type
+  structure. We illustrate the usefulness of our library in a formalization of
+  L\"ob's theorem which closely follows a classical proof of this result.
 \end{abstract}
-
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
 
@@
 logic~\cite{boolos_1994} to abstract details from encodings in G\"odel's
-incomleteness results.
+incompleteness results.
+
+In order to prove L\"ob's theorem proof we need to encode modal
+fixed points. Let $\varphi(A)$ be a formula with only one propositional variable
+$A$. We define as a modal fixed point of $\varphi$ is a sentence $\Psi$ such that
+$\Psi \leftrightarrow \varphi(\Box \Psi)$.
+We represent the modal fixed points by definition \mintinline{coq}{fixed_point},
+which is parameterized by a system for modal logics, \mintinline{coq}{S}, and a
+theory, \mintinline{coq}{g}, and encodes the logical equivalence between a
+formula \mintinline{coq}{p} and its fixpoint \mintinline{coq}{f ([] p)}. 
+\begin{minted}{coq}
+Definition fixed_point S g
+   := forall f, exists p,
+          (S ; g |-- [! (p <-> f ([] p)) !]).
+\end{minted}
+
+Using the definition of fixed points, we can state L\"ob's theorem as follows:
+\begin{minted}{coq}
+Theorem Lob:
+  forall P, fixed_point K4 nil ->
+  (K4; nil |-- [! []P -> P !]) ->
+  (K4; nil |-- [! P !]).
+\end{minted}
+
+The proof starts by assuming an arbitrary formula $P$, the existence of modal fixed
+points and that $\Box P \to P$ is provable in K4. These deduction steps
+are performed by the following tactic:
+\begin{minted}{coq}
+intros P FP H1.
+\end{minted}
+
+Next, from the existence of modal fixed points for every formula,
+we know that there is a formula $\Psi$ such that $\Psi \leftrightarrow (\Box
+\Psi \to P)$. The following tactic performs this step of reasoning.
+\begin{minted}{coq}
+destruct FP with (fun A => [! A -> P !]) as (psi, H2).
+\end{minted}
+Now, we are able to deduce that $\Psi \to (\Box \Psi\to P)$ from the previous
+step result as follows:
+\begin{minted}{coq}
+assert (K4; nil |-- [! psi -> []psi -> P !]) as H3.
+apply modal_ax5 in H2; try repeat constructor ;
+   try eassumption.
+\end{minted}
+The \mintinline{coq}{assert} tactic allows the definition of
+auxiliary proof goals. In the previous code piece, we specified
+$\vdash \Psi \to (\Box \Psi \to P)$ as a new goal which
+generates a new proof hypothesis named \mintinline{coq}{H3}. The
+deduction of this fact is immediate from K4 axioms.
+
+From $\vdash \Psi \to (\Box \Psi \to P)$ and the necessitation
+rule, we can deduce $\vdash \Box (\Psi \to (\Box \Psi \to P))$
+using the following tactics:
+\begin{minted}{coq}
+assert (K4; nil |-- [! [](psi -> []psi -> P) !]) as H4.
+apply Nec; auto.
+\end{minted}
+
+Now, using the axiom \textbf{K} we can deduce $\vdash \Box \Psi \to
+\Box (\Box \Psi \to P)$. Again, we deduce this fact using the
+\mintinline{coq}{assert} tactic.
+\begin{minted}{coq}
+assert (K4; nil |-- [! []psi -> []([]psi -> P) !]) as H5.
+apply modal_axK in H4; try repeat constructor ;
+  try eassumption.
+\end{minted}
+
+The next proof step involves showing that
+$\vdash \Box (\Box \Psi \to P)\to (\Box \Box \Psi \to \Box P)$ holds
+by axiom \textbf{K}.
+\begin{minted}{coq}
+assert (K4; nil |-- [! []([]psi -> P) ->
+                       [][]psi -> []P !]) as H6.
+eapply Ax with (a := axK ?[X] ?[Y]);
+  try repeat constructor.
+\end{minted}
+
+Using $\vdash \Box (\Box \Psi \to P)\to (\Box \Box \Psi \to \Box P)$ and
+$\vdash \Box \Psi \to\Box (\Box \Psi \to P)$ we can deduce
+$\vdash \Box \Psi \to \Box \Box \Psi \to \Box P$ using composition of
+the implication.
+\begin{minted}{coq}
+assert (K4; nil |-- [! []psi -> [][]psi -> []P !]) as H7.
+eapply modal_compose; try repeat constructor ;
+   try eassumption.
+\end{minted}
+
+Now, we can show that $\vdash\Box\Psi\to\Box\Box\Psi$ using
+the axioms of K4. The next snippet of tactics perform this
+reasoning step.
+\begin{minted}{coq}
+assert (K4; nil |-- [! []psi -> [][]psi !]) as H8.
+apply modal_axK4; try constructor 2.
+\end{minted}
+
+From $\vdash \Box\Psi \to \Box P$ and $\vdash \Box P \to P$, we can
+conclude $\vdash \Box \Psi \to P$, by composition of implication as
+follows.
+\begin{minted}{coq}
+assert (K4; nil |-- [! []psi -> P !]) as H10.
+eapply modal_compose; try repeat constructor ;
+   try eassumption.
+\end{minted}
+
+Next, we can conclude $\vdash \Box (\Psi \to P) \to \Psi$ from
+$\Psi \leftrightarrow (\Box \Psi \to P)$ and the axioms of K4 using
+the following tactics:
+\begin{minted}{coq}
+assert (K4; nil |-- [! ([]psi -> P) -> psi !]) as H11.
+apply modal_ax6 in H2; try repeat constructor ;
+ try eassumption.
+\end{minted}
+
+From $\vdash (\Box\Psi \to P) \to \Psi$ and $\vdash \Box \Psi \to P$
+we can conclude $\vdash \Psi$ by modus ponens.
+\begin{minted}{coq}
+assert (K4; nil |-- psi) as H12.
+eapply Mp ; try eassumption.
+\end{minted}
+
+By necessitation rule, we can deduce $\vdash \Box \Psi$
+\begin{minted}{coq}
+assert (K4; nil |-- [! []psi !]) as H13.
+apply Nec ; try eassumption.
+\end{minted}
+
+Finally, the desired conclusion $\vdash P$ follows from
+using the rule modus ponens over $\vdash \Box \Psi \to P$
+and $\vdash \Box \Psi$, as presented by the next tactics
+snippet.
+\begin{minted}{coq}
+eapply Mp ; try eassumption.
+\end{minted}
+
+This concludes the formalization of the L\"ob's theorem using
+the developed library.
 
@@
 
+The next table summarizes the different systems of modal logics considered by
+previous works.
+
 \vspace{0.5cm} {

### 2022-06-20 paper/paper.tex

@@
   Our formalization differs from previous attempts by providing a uniform
-  representation of several systems for modal logics using Coq's rich type
+  representation of several systems for modal logic using Coq's rich type
   structure. We illustrate the usefulness of our library in a formalization of
@@
 
-The next table summarizes the different systems of modal logics considered by
-previous works.
-
+The next table summarizes the formalized modal logics systems by our library
+and related works.
 \vspace{0.5cm} {
@@
 }
+All developments considered systems K, T and K4 and the unique formalization that
+includes system GL is the library described in this work.
+
+\paragraph{Applications of modal logics in semantics} Another application of
+modal logics is its use for reasoning about programming languages semantics.
+Dynamic logic was introduced by Pratt in late 70's as modal logic for reasoning
+about programs~\cite{Pratt76,Pratt80}. The key idea on dynamic logic is to express the
+Hoare triples by a modality meaning ``after a program $p$ executes, formula
+$\varphi$ will be true''. The programming languages considered in Pratt's
+original work are composed by an assignment and regular language operators
+(union, concatenation and Kleene star)~\cite{Pratt76}.
+
+An application of modal logic IS5 as a type system for a distributed
+programming language was proposed in~\cite{Murphy04}. 
+In that work, modalities are used to represent types for mobile code
+and remote addresses. Murphy et. al defined the language by interpreting
+the possible worlds of the intutionistic modal logic IS5 to be nodes on a
+network and the modalities $\Box$ and $\Diamond$ to reflect mobility and
+locality, respectively. Authors develop an operational semantics and show
+that their formalization is type-safe.
+
+The application of a multimodal separation logic for the semantics of low-level
+was the subject of Dockins et. al. work~\cite{Dockins08}. In that work, authors
+combine several modal operators that do not need to satisfy the same frame
+conditions and seperation logics for reasoning about non-trivial subsets of
+low-level programming languages. Authors developed their formalization through a
+shallow embedding in Coq proof assistant and provide two case studies: one is
+the verification of a simple low-level virtual machine, called the list machine
+benchmark and the other is the subset of Sequential Cminor.
+
+A model for recursive, impredicatively quantified types with mutable references
+was studied by Appel et. al.~\cite{Appel07}. His interest in formalizing such
+advanced typing concepts was in the context of typed assembly languages.
+Authors stabilished a semantic type soundness proof by defining a Kripke-style 
+semantics of the G\"odel-L\"ob logic of provability. All their definitions were
+mechanized through a shallow embedding in Coq.
+
+The embedding of modal linear logic in Coq was the subject of
+Sadrzadeh's work~\cite{Soton03}. Like our work, Sadrzadeh's uses a
+deep-embedding approach and assumes that all modalities are reflexive
+and transitive. Using her embedding, Sadrzadeh shows some proofs of classical
+logic puzzles.
 

### 2022-06-20 paper/paper.tex
More updates on paper

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf]{acmart}
+\documentclass[sigconf,anonymous]{acmart}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Towards a Sound and Complete Library for Modal Logics Reasoning in Coq}
+\title{A Sound Deep Embedding of Arbitrary Normal Modal Logics in Coq}
 
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{Modal logics, soundness, completeness, Coq proof assistant}
+\keywords{Modal logics, soundness, Coq proof assistant}
 
@@
 
-Our society depends heavily on computing technology: our cars, airplanes, medical devices
-and banking applications all rely on software artifacts. In this scenario,
-\emph{formal methods} play a crucial role to ensure that a program follows its
-specification, thus avoiding failures which can cause castrophic outcomes. Interactive
-proof assistants, like Coq~\cite{manual_coq} and Isabelle/HOL~\cite{manual_isabelle}, can
-be used to verify software components and have been applied with success in
-non-trivial projects like a certified compiler for C programming
-language~\cite{Leroy09} and the kernel of seL4 operating system~\cite{Klein10}.
+The start of modal logics can be traced back to Lewis and his definition of 
+logical systems with strict implication instead of material implication \cite{Lewis1918}.
+Based in this work~\cite{Lewis1932}, independently of classical propositional logic,
+defined the systems $\mathsf{S1}$--$\mathsf{S5}$, where $\mathsf{S4}$ and $\mathsf{S5}$
+are well known normal modal systems. It was, in fact, G{\"o}del that had the idea 
+to build a modal logic system as an extension of the classical propositional calculus, 
+when constructing a system similar to $\mathsf{S4}$ in \cite{Godel1933}.
 
-Usually, complex software systems are formed by various inter-communicating
-components. A key aspect of the design of such programs is how to ensure that
-a failure of a single component does not compromise the whole system.
-The main difficulty is such formalization tasks is how to guarantee correct
-error detection and that its corresponding fix will ensure that the software
-still continue to work under abnormal conditions. One way to express such
-requirements is using \emph{modal logics}, which can specify communication
-patterns and knowledge in software agents~\cite{Huth_Ryan}.
+Despite that the original motivation of modal systems was to deal with strict implication, 
+for many years modal logic was seen as ``the logic of necessity and possibility''. 
+Only after the introduction of Kripke semantics they were seen also as ``simple yet
+expressive languages for talking about relational structures''~\cite{Blackburn-ModalLogics}. 
+A great multitude of modal systems, with various kinds of modal operators, each of them with 
+different meanings or properties, have then been constructed in order to provide effective
+formalisms for talking about time, space, knowledge, beliefs, actions, 
+obligations, among others.
 
-While modal logics allow the modelling of such requirements, proving properties
-about them may be a hard task without proper tool support. There are some
-dedicated provers for modal logics~\cite{Gleissner17,Mora11}, but we believe
-that using mature tools, like a proof assistant, are a better option for
+The interpretation of modal logic simply as ``the logic of necessity'' restricts 
+not only its applications but also the range of modal systems. In fact, modal logic 
+can also be seen as ``the study of \emph{modalities} -- logical operations that qualify 
+assertions about the truth of statements''~\cite{Goldblatt-MathofModality}. In this sense, 
+the necessitation rule (from $\varphi$ infers $\Box\varphi$) can be relaxed. This brings 
+a number of modal systems known as non-normal, whose semantics was defined only in the 
+1970's by Montague~\cite{Montague1970} and Scott~\cite{Scott_AdviceModal1970}.
+
+Such wide range of highly expressive logical languages is the ideal tool to 
+formalize the many complexities of modern advanced software systems, 
+such as those found in cars, airplanes, medical devices and banking applications.
+In combining said modal logics with interactive proof assistants,
+such as Coq~\cite{manual_coq} and Isabelle/HOL~\cite{manual_isabelle}, both of
+which have been applied with success in non-trivial projects like a certified compiler 
+for C programming language~\cite{Leroy09} and the kernel of seL4 operating
+system~\cite{Klein10}, it is possible to formalize and verify highly complex computer
+systems.
+
+Although there are some dedicated provers for modal logics~\cite{Gleissner17,Mora11},
+we believe that using mature tools, like a proof assistant, are a better option for
 verifying facts using modal logics. We justify this choice by: 1) the recent
@@
 
+% Our society depends heavily on computing technology: our cars, airplanes, medical devices
+% and banking applications all rely on software artifacts. In this scenario,
+% \emph{formal methods} play a crucial role to ensure that a program follows its
+% specification, thus avoiding failures which can cause castrophic outcomes. Interactive
+% proof assistants, like Coq~\cite{manual_coq} and Isabelle/HOL~\cite{manual_isabelle}, can
+% be used to verify software components and have been applied with success in
+% non-trivial projects like a certified compiler for C programming
+% language~\cite{Leroy09} and the kernel of seL4 operating system~\cite{Klein10}.
+
+% Usually, complex software systems are formed by various inter-communicating
+% components. A key aspect of the design of such programs is how to ensure that
+% a failure of a single component does not compromise the whole system.
+% The main difficulty is such formalization tasks is how to guarantee correct
+% error detection and that its corresponding fix will ensure that the software
+% still continue to work under abnormal conditions. One way to express such
+% requirements is using \emph{modal logics}, which can specify communication
+% patterns and knowledge in software agents~\cite{Huth_Ryan}.
+
+% While modal logics allow the modelling of such requirements, proving properties
+% about them may be a hard task without proper tool support. There are some
+% dedicated provers for modal logics~\cite{Gleissner17,Mora11}, but we believe
+% that using mature tools, like a proof assistant, is a better option for
+% verifying facts using modal logics. We justify this choice by: 1) the recent
+% successful histories on verification of software and
+% mathematics~\cite{Leroy09,Klein10,Gonthier13,Gonthier08}; 2) There is a growing
+% community of developers using such tools thanks to 
+% textbooks on the subject~\cite{Nipkow14,Bertot10,Chlipala13} which allows
+% newbies to join the game of formal verification and 3) The custom automation
+% facilities supported by current proof assistants allows that trivial facts can
+% be proved using decision procedures, relieving the user from writing proof
+% scripts for formulas that can be checked automatically.
+
+% In this context, the present work aims to contribute by constructing a library
+% that formalize a family of normal modal logics and its related properties in Coq
+% by parametrizing the proofs over a set of possible axioms. 
+% In using such a library, a developer could use all Coq's resources to prove
+% modal logics facts.
+
 More specifically, we contribute:
@@
     parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL
-    and prove its soundness with respect to Kripke semantics. We also
-    discuss the current status of the formalization of the completeness result.
+    and prove its soundness with respect to Kripke semantics.
+    We also formalize several frame correspondence results.
   \item We show a proof of the L\"ob's theorem using the Hilbert deductive system
@@
 Section~\ref{sec:implementation}. A case study in proving the L\"ob's theorem is
-the subject of Section~\ref{sec:case}. Related work is discussed
+the subject of Section~\ref{sec:case}. Related works are discussed
 on Section~\ref{sec:related} and Section~\ref{sec:conclusion} draws some conclusions and
@@
 \]
-Apart from the axioms, we also need the traditional \emph{modus ponens} rule
-and a rule that allows to conclude formulas with necessity.
+Apart from the axioms, we also need \emph{modus ponens} and necessitation rules.
 
@@
 \mintinline{coq}{F}. The encoding of the Kripke semantics is done by a Coq function
-that which implements the relation of Figure~\ref{fig:kripke}. We omit the code
+that implements the relation of Figure~\ref{fig:kripke}. We omit the code
 for propositional logics connectives for brevity. The interesting parts are for
@@
 Inductive axiom : Set :=
-  | ax1   : modalFormula -> modalFormula -> axiom
-  | ax2   : modalFormula -> modalFormula ->
-                            modalFormula -> axiom
+  | ax1   : form -> form -> axiom
+  | ax2   : form -> form -> form  -> axiom
   (** some code omitted *)
-  | axK   : modalFormula -> modalFormula -> axiom
-  | axPos : modalFormula -> modalFormula -> axiom
-  | axT   : modalFormula -> axiom
-  | axB   : modalFormula -> axiom
-  | axK4  : modalFormula -> axiom
-  | axD   : modalFormula -> axiom
-  | axK5  : modalFormula -> axiom.
+  | axK   : form -> form -> axiom
+  | axPos : form -> form -> axiom
+  | axT   : form -> axiom
+  | axB   : form -> axiom
+  | axK4  : form -> axiom
+  | axD   : form -> axiom
+  | axK5  : form -> axiom.
 \end{minted}
@@
 
-The \mintinline{coq}{Ax} constructor uses the \mintinline{coq}{instantiate} to
-create a formula instance from using a given system, represented as a predicate
+The \mintinline{coq}{Ax} constructor uses the \mintinline{coq}{instantiate}
+function to create a formula instance from using a given system,
+represented as a predicate
 \mintinline{coq}{A : axiom -> Prop}, which parameterizes our deduction type. An
@@
 
-While we have completely proved the soundness theorem, the current version of
-our library does not have a completeness theorem.
+\paragraph{Frame Correspondence} 
+As already mentioned, there are some modal formula schemas that correspond
+to properties of the accessibility relation of a modal system, if we take them
+as an axiom in the system.  While the proof that a certain property in a frame
+implies the correspondent axiom is usually  direct, the inverse often needs
+more effort. On the Coq proofs developed for this work,  the former case is
+mostly unfoldings, simplifications and applications of premises
+while the latter, appart from the necessity and burden of a proof by contraposition, 
+requires the construction of a witness valuation function. 
+Nevertheless, all Coq's proofs are fairly similar to the respective
+``pen and paper'' proofs that  can be found in the literature. 
+Our work proves the correspondence for the  following properties in
+accessibility relations: reflexivity, transitivity, symmetry,  euclidean,
+seriality, functionality, density, convergency, and converse well-foundness.
 
@@
 
-Using the definition of fixed points, we can state L\"ob's theorem as follows:
+\noindent Using the definition of fixed points, we can state L\"ob's theorem as follows:
+
 \begin{minted}{coq}
-Theorem Lob:
-  forall P, fixed_point K4 nil ->
+  Theorem Lob: forall P, fixed_point K4 nil ->
                (K4; nil |-- [! []P -> P !]) ->
@@
 
-The proof starts by assuming an arbitrary formula $P$, the existence of modal fixed
+\noindent The proof starts by assuming an arbitrary formula $P$, the existence of modal fixed
 points and that $\Box P \to P$ is provable in K4. These deduction steps
 are performed by the following tactic:
+
 \begin{minted}{coq}
@@
 
-Next, from the existence of modal fixed points for every formula,
+\noindent Next, from the existence of modal fixed points for every formula,
 we know that there is a formula $\Psi$ such that $\Psi \leftrightarrow (\Box
 \Psi \to P)$. The following tactic performs this step of reasoning.
+
 \begin{minted}{coq}
@@
 \end{minted}
-Now, we are able to deduce that $\Psi \to (\Box \Psi\to P)$ from the previous
+
+\noindent Now, we are able to deduce that $\Psi \to (\Box \Psi\to P)$ from the previous
 step result as follows:
+
 \begin{minted}{coq}
@@
 \end{minted}
-The \mintinline{coq}{assert} tactic allows the definition of
+
+\noindent The \mintinline{coq}{assert} tactic allows the definition of
 auxiliary proof goals. In the previous code piece, we specified
@@
 \begin{minted}{coq}
-assert (K4; nil |-- [! [](psi -> []psi -> P) !]) as H4.
+  assert (K4; nil |-- [! [](psi -> []psi -> P) !])
+    as H4.
   apply Nec; auto.
@@
 
-Now, using the axiom \textbf{K} we can deduce $\vdash \Box \Psi \to
+\noindent Now, using the axiom \textbf{K} we can deduce $\vdash \Box \Psi \to
 \Box (\Box \Psi \to P)$. Again, we deduce this fact using the
 \mintinline{coq}{assert} tactic.
+
 \begin{minted}{coq}
-assert (K4; nil |-- [! []psi -> []([]psi -> P) !]) as H5.
+  assert (K4; nil |-- [! []psi -> []([]psi -> P) !])
+    as H5.
   apply modal_axK in H4; try repeat constructor ;
@@
 
-The next proof step involves showing that
+\noindent The next proof step involves showing that
 $\vdash \Box (\Box \Psi \to P)\to (\Box \Box \Psi \to \Box P)$ holds
 by axiom \textbf{K}.
+
 \begin{minted}{coq}
@@
 
-Using $\vdash \Box (\Box \Psi \to P)\to (\Box \Box \Psi \to \Box P)$ and
+\noindent Using $\vdash \Box (\Box \Psi \to P)\to (\Box \Box \Psi \to \Box P)$ and
 $\vdash \Box \Psi \to\Box (\Box \Psi \to P)$ we can deduce
-$\vdash \Box \Psi \to \Box \Box \Psi \to \Box P$ using composition of
-the implication.
+$\vdash \Box \Psi \to \Box \Box \Psi \to \Box P$ using transitivity of
+implication.
+
 \begin{minted}{coq}
-assert (K4; nil |-- [! []psi -> [][]psi -> []P !]) as H7.
+  assert (K4; nil |-- [! []psi -> [][]psi -> []P !])
+    as H7.
   eapply modal_compose; try repeat constructor ;
@@
 
-Now, we can show that $\vdash\Box\Psi\to\Box\Box\Psi$ using
+\noindent Now, we can show that $\vdash\Box\Psi\to\Box\Box\Psi$ using
 the axioms of K4. The next snippet of tactics perform this
 reasoning step.
+
 \begin{minted}{coq}
@@
 
-From $\vdash \Box\Psi \to \Box P$ and $\vdash \Box P \to P$, we can
+\noindent From $\vdash \Box\Psi \to \Box P$ and $\vdash \Box P \to P$, we can
 conclude $\vdash \Box \Psi \to P$, by composition of implication as
 follows.
+
 \begin{minted}{coq}
@@
 
-Next, we can conclude $\vdash \Box (\Psi \to P) \to \Psi$ from
+\noindent Next, we can conclude $\vdash \Box (\Psi \to P) \to \Psi$ from
 $\Psi \leftrightarrow (\Box \Psi \to P)$ and the axioms of K4 using
 the following tactics:
+
 \begin{minted}{coq}
@@
 
-From $\vdash (\Box\Psi \to P) \to \Psi$ and $\vdash \Box \Psi \to P$
+\noindent From $\vdash (\Box\Psi \to P) \to \Psi$ and $\vdash \Box \Psi \to P$
 we can conclude $\vdash \Psi$ by modus ponens.
+
 \begin{minted}{coq}
@@
 
-By necessitation rule, we can deduce $\vdash \Box \Psi$
+\noindent By necessitation rule, we can deduce $\vdash \Box \Psi$
+
 \begin{minted}{coq}
@@
 
-Finally, the desired conclusion $\vdash P$ follows from
+\noindent Finally, the desired conclusion $\vdash P$ follows from
 using the rule modus ponens over $\vdash \Box \Psi \to P$
@@
 snippet.
+
 \begin{minted}{coq}
@@
 
-This concludes the formalization of the L\"ob's theorem using
+\noindent This concludes the formalization of the L\"ob's theorem using
 the developed library.
@@
 
+The development of a completeness proof for modal logic S5 in Lean proof
+assistant was the subject of the Bentzen's work~\cite{Bentzen2021}. The author
+follows the approach due to Henkin for S5 completeness and it depends on
+the \texttt{mathlib}, the Lean's library of mathematical results~\cite{mathlib20}.
+
+Maggesi et. al. developed a formalization of the completeness proof for GL using
+the HOL light theorem prover~\cite{Maggesi21}. The authors argue that using the
+standard Henkin strategy does not work in GL without modifications. The core idea of
+Maggesi et. al. strategy was building maximal consistent lists in terms of
+components of the given formula non-proven in the calculus.
+
 The next table summarizes the formalized modal logics systems by our library
@@
       Benzmüller et. al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X &  \\ \hline
+      Bentzen~\cite{Bentzen2021} & X & & & X & & X & X & X & \\ \hline
+      Maggesi et. al.~\cite{Maggesi21} & X & & & & & & & & X \\ \hline
       This work         & X & X & X & X & X  &  X &  X & X  & X \\ \hline
@@
 }
-All developments considered systems K, T and K4 and the unique formalization that
-includes system GL is the library described in this work.
+
+As we can see, as far we know, our work is the unique formalization that covers systems
+K, D, B, T, K4, K5, S4, S5 and GL. Also, since our deduction is
+parameterized by system of axioms considered, it can be used to represent any
+system of normal modal logics, unlike any other formalizations found in the literature.
 
@@
 advanced typing concepts was in the context of typed assembly languages.
-Authors stabilished a semantic type soundness proof by defining a Kripke-style 
+The authors stabilished a semantic type soundness proof by defining a Kripke-style 
 semantics of the G\"odel-L\"ob logic of provability. All their definitions were
@@
 
-The embedding of modal linear logic in Coq was the subject of
-Sadrzadeh's work~\cite{Soton03}. Like our work, Sadrzadeh's uses a
-deep-embedding approach and assumes that all modalities are reflexive
-and transitive. Using her embedding, Sadrzadeh shows some proofs of classical
-logic puzzles.
+% The embedding of modal linear logic in Coq was the subject of
+% Sadrzadeh's work~\cite{Soton03}. Like our work, Sadrzadeh's uses a
+% deep-embedding approach and assumes that all modalities are reflexive
+% and transitive. Using her embedding, Sadrzadeh shows some proofs of classical
+% logic puzzles.
 
 \section{Conclusion}\label{sec:conclusion}
+
+In this work, we presented a sound deep embedding of arbitrary
+normal modal logics in Coq. Our approach is able to represent uniformly
+deduction over different modal logic systems by using different axiom sets
+as predicate parameter in the deduction inductive type. Using this generic
+representation of deduction we are able to prove the soundness theorem and
+apply it on proving L\"ob's theorem.
+
+As future work we intend to prove the completeness result for the Hilbert
+deduction system, include support for multimodalities (fusion of modal logics)
+and non-normal modal logic systems.
 

### 2022-06-20 paper/paper.tex
Fixing a typo

@@
 deduction over different modal logic systems by using different axiom sets
-as predicate parameter in the deduction inductive type. Using this generic
+as a predicate parameter in the deduction inductive type. Using this generic
 representation of deduction we are able to prove the soundness theorem and

### 2022-06-21 paper/paper.tex
Small textual fixes

@@
 Although there are some dedicated provers for modal logics~\cite{Gleissner17,Mora11},
-we believe that using mature tools, like a proof assistant, are a better option for
+we believe that using mature tools, like a proof assistant, is a better option for
 verifying facts using modal logics. We justify this choice by: 1) the recent
@@
 generalize results about derivations to several deductive systems, independently
-of the axioms used. We conclude by discussing the soundness, the current
-status of the completeness and the frame correspondence proofs.
+of the axioms used. We conclude by discussing the soundness and the frame
+correspondence proofs.
 
@@
 \paragraph{The Hilbert deductive system}
-Basically, a Hilbert-style deductive system allows the use of on four
+Basically, a Hilbert-style deductive system allows using one out of four
 possibilities in a proof: 1) the use of a hypothesis; 2) use of an axiom;
@@
 represent rules for hypothesis, modus ponens and generalization, respectively.
-Before discuss the rule for axioms, we need to discuss how axioms are
+Before discuss the \mintinline{coq}{Ax}, we need to discuss how axioms are
 represented in our library. The inductive type \mintinline{coq}{axiom} specifies
@@
 The \mintinline{coq}{Ax} constructor uses the \mintinline{coq}{instantiate}
-function to create a formula instance from using a given system,
+function to create a formula instance from a given system,
 represented as a predicate
-\mintinline{coq}{A : axiom -> Prop}, which parameterizes our deduction type. An
+\mintinline{coq}{A : axiom -> Prop} which parameterizes our deduction type. An
 axiom \mintinline{coq}{a} is valid within a system \mintinline{coq}{A} whenever
 the type \mintinline{coq}{A a} is inhabited, i.e. it is not empty. Next, we
-detail the representation of different systems of modal logic in our library.
+describe how to represent different systems of modal logic.
 
@@
 different systems of modal logics uniformly. In order to understand how it
-works, let us review its type (below, remeber that \mintinline{coq}{A : axiom -> Prop}
+works, let us review its type (below, remember that \mintinline{coq}{A : axiom -> Prop}
 denotes a system of axioms for modal logics):
@@
 \end{minted}
-Notice that the first parameter of constructor \mintinline{coq}{Ax} is a value
+Note that the first parameter of constructor \mintinline{coq}{Ax} is a value
 of type \mintinline{coq}{A a}, which is only inhabited if the predicate

### 2022-06-21 paper/paper.tex
Adding fixes suggested by Ariel

@@
 that formalize a family of modal logics and its related properties in Coq.
-By using such a library, a developer could use all Coq's resources to prove
+By using such a library, a developer could use all of Coq's resources to prove
 modal logics facts.
@@
     We also formalize several frame correspondence results.
-  \item We show a proof of the L\"ob's theorem using the Hilbert deductive system
+  \item We show a proof of L\"ob's theorem using the Hilbert deductive system
     developed in our library.
@@
 the implementation of our Coq library are presented in
-Section~\ref{sec:implementation}. A case study in proving the L\"ob's theorem is
+Section~\ref{sec:implementation}. A case study in proving L\"ob's theorem is
 the subject of Section~\ref{sec:case}. Related works are discussed
@@
 Coq is a proof assistant based on the calculus of inductive
-constructions (CIC)~\cite{manual_cic}, a higher order typed
+constructions (CIC)~\cite{manual_cic}, a higher-order typed
 $\lambda$-calculus extended with inductive definitions.  Theorem
@@
 \emph{tactics}, which are commands that can be used to construct proof
-terms in a more user friendly way.
+terms in a more user-friendly way.
 
-As a tiny example, consider the task of proving the following simple
-formula of propositional logic:
+As a tiny example, consider the task of proving the following 
+formula of the propositional logic:
 \[
@@
 the Coq proof assistant provides not only a great number of tactics
-but also a domain specific language for scripted proof automation,
+but also a domain-specific language for scripted proof automation,
 called $\mathcal{L}$tac. Details about $\mathcal{L}$tac and Coq can be found
@@
 $w$ and formula $\varphi$ we take, the formula $\Box \varphi \to \varphi$ will
-be true there. Example of properties that characterize certain formula schemas are
+be true there. Examples of properties that characterize certain formula schemas are
 presented in the table below.
@@
 \paragraph{Syntax and semantics of modal logics}
-In order to implement our library, we use a \emph{deep embedding} of
+To implement our library, we use a \emph{deep embedding} of
 formulas, i.e. we represent formulas as the following Coq inductive datatype.
@@
 \end{minted}
-A model is denoted by another record formed by a frame and a labelling function,
+A model is denoted by another record formed by a frame and a labeling function,
 which assign the truth value of a variable at a given world.
@@
 \end{minted}
-Note that the type of the labelling function \mintinline{coq}{v} ensures that it
+Note that the type of the labeling function \mintinline{coq}{v} ensures that it
 refers to a world from the set of possible worlds in the frame
@@
 for propositional logics connectives for brevity. The interesting parts are for
-variables (constructor \mintinline{coq}{Lit}) which uses the labelling function
+variables (constructor \mintinline{coq}{Lit}) which uses the labeling function
 \mintinline{coq}{v} from the model \mintinline{coq}{M}. The meaning of
@@
 \paragraph{The Hilbert deductive system}
-Basically, a Hilbert-style deductive system allows using one out of four
-possibilities in a proof: 1) the use of a hypothesis; 2) use of an axiom;
-3) use of the modus ponens rule or 4) the use of the generalization (necessity)
+Basically, a Hilbert-style deductive system allows the use of four
+possibilities in a proof: 1) the use of a hypothesis; 2) the use of an axiom;
+3) the use of the modus ponens rule or 4) the use of the generalization (necessity)
 rule. The inductive type \mintinline{coq}{deduction} encodes these possibilities
@@
 represent rules for hypothesis, modus ponens and generalization, respectively.
-Before discuss the \mintinline{coq}{Ax}, we need to discuss how axioms are
+Before discussing the \mintinline{coq}{Ax}, we need to discuss how axioms are
 represented in our library. The inductive type \mintinline{coq}{axiom} specifies
@@
 Axioms are built using the function \mintinline{coq}{instantiate}, which
-receives an axiom and build its corresponding formula by pattern matching as
+receives an axiom and builds its corresponding formula by pattern matching as
 follows:
@@
 mostly unfoldings, simplifications and applications of premises
-while the latter, appart from the necessity and burden of a proof by contraposition, 
+while the latter, apart from the necessity and burden of a proof by contraposition, 
 requires the construction of a witness valuation function. 
@@
 In this section we describe a case study of the defined modal logic library.
-We proved the L\"ob theorem in the system K4 with modal fixed points.
+We proved L\"ob theorem in the system K4 with modal fixed points.
 The importance of L\"ob's theorem is due to its use in provability
@@
 
-\noindent This concludes the formalization of the L\"ob's theorem using
+\noindent This concludes the formalization of L\"ob's theorem using
 the developed library.
@@
 which consists of  the axioms $K$, $T$, $4$ and $5$. The objective was the
-modelling of two well known puzzles.
+modeling of two well known puzzles.
 Wind's formalization was separated into some components. The first consists in
@@
 As future work we intend to prove the completeness result for the Hilbert
-deduction system, include support for multimodalities (fusion of modal logics)
+deduction system, including support for multimodalities (fusion of modal logics)
 and non-normal modal logic systems.

### 2022-06-21 paper/paper.tex
Non pushed changes to paper

@@
 provide some pointers for future works.
+
+All source code in this article has been formalized in Coq version 8.15.0 and it
+is available at~\cite{modal-coq-lib}.
 
@@
 representation of deduction we are able to prove the soundness theorem and
-apply it on proving L\"ob's theorem.
+apply it on proving L\"ob's theorem. All source code in this article has
+been formalized in Coq version 8.15.0 and it is available at~\cite{modal-coq-lib}.
 
 As future work we intend to prove the completeness result for the Hilbert
-deduction system, including support for multimodalities (fusion of modal logics)
-and non-normal modal logic systems.
+deduction system and continue the development of our library by including
+support for multimodalities (fusion of modal logics) and non-normal modal
+logic systems.
 

### 2022-08-30 paper/paper.tex
Minor fixes (typos, reviewers' suggestions, etc)

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf,anonymous]{acmart}
+\documentclass[sigconf]{acmart}
 
@@
 \usepackage{minted}
-\usepackage{bussproof}
+\usepackage{bussproofs}
 \usepackage{multirow}
@@
 %% used to denote shared contribution to the research.
+
 \author{Ariel Agne da Silveira}
-\author{Miguel}
+\affiliation{\institution{Federal University of Ouro Preto}\country{}}
+\email{ariel.silveira@aluno.ufop.edu.br}
+
+\author{Miguel Nunes}
+\affiliation{\institution{Santa Catarina State University}\country{}}
+\email{miguel.alfredo.nunes@gmail.com}
+
+\author{Karina Roggia}
+\affiliation{\institution{Santa Catarina State University}\country{}}
+\email{karina.roggia@udesc.br}
+
 \author{Paulo Torrens}
-\author{Karina Roggia}
+\affiliation{\institution{Santa Catarina State University}\country{}}
+\email{paulotorrens@gnu.org}
+
 \author{Rodrigo Ribeiro}
+\affiliation{\institution{Federal University of Ouro Preto}\country{}}
+\email{rodrigo.ribeiro@ufop.edu.br}
 
@@
 for many years modal logic was seen as ``the logic of necessity and possibility''. 
-Only after the introduction of Kripke semantics they were seen also as ``simple yet
+Only after the introduction of Kripke semantics they were also seen as ``simple yet
 expressive languages for talking about relational structures''~\cite{Blackburn-ModalLogics}. 
@@
 assertions about the truth of statements''~\cite{Goldblatt-MathofModality}. In this sense, 
-the necessitation rule (from $\varphi$ infers $\Box\varphi$) can be relaxed. This brings 
-a number of modal systems known as non-normal, whose semantics was defined only in the 
+the necessitation rule (from $\varphi$ infer $\Box\varphi$) may be relaxed. This brings 
+a number of modal systems known as non-normal, whose semantics were defined only in the 
 1970's by Montague~\cite{Montague1970} and Scott~\cite{Scott_AdviceModal1970}.
@@
 successful histories on verification of software and
-mathematics~\cite{Leroy09,Klein10,Gonthier13,Gonthier08}; 2) There is a growing
+mathematics~\cite{Leroy09,Klein10,Gonthier13,Gonthier08}; 2) there is a growing
 community of developers using such tools thanks to 
 textbooks on the subject~\cite{Nipkow14,Bertot10,Chlipala13} which allows
-newbies to join the game of formal verification and 3) The custom automation
-facilities supported by current proof assistants allows that trivial facts can
-be proved using decision procedures, relieving the user from writing proof
+newcomers to join the game of formal verification; and 3) the custom automation
+facilities supported by current proof assistants allows for trivial facts to
+be proven using decision procedures, relieving the user from writing proof
 scripts for formulas that can be checked automatically.
@@
 In this context, the present work aims to contribute by constructing a library
-that formalize a family of modal logics and its related properties in Coq.
-By using such a library, a developer could use all of Coq's resources to prove
+that formalizes a family of normal modal logics and their related properties in Coq.
+By using such a library, a developer may use all of Coq's resources to prove
 modal logics facts.
@@
   \item We model the Hilbert deductive systems as a Coq inductive type that is
-    parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL
+    parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL,
     and prove its soundness with respect to Kripke semantics.
@@
   Kolmogorov, de Bruijn and Martin-L\"of Correspondence. This is also
-  known as the Curry-Howard ``isomorphism''.}, where types represent
+  known as the Curry-Howard isomorphism.}, where types represent
 logical formulas, $\lambda$-terms represent proofs
@@
 Section EXAMPLE.
-   Variables A B C : Prop.
-   Theorem example : (A -> B) -> (B -> C) -> A -> C.
+   Variables A B C: Prop.
+   Theorem example: (A -> B) -> (B -> C) -> A -> C.
    Proof.
@@
 Definition example :(A -> B) -> (B -> C) -> A -> C :=
- fun (H : A -> B)
-     (H' : B -> C)
-     (HA : A) => H' (H HA).
+ fun (H: A -> B)
+     (H': B -> C)
+     (HA: A) => H' (H HA).
 \end{minted}
@@
 \begin{minted}{coq}
-Inductive form : Set :=
+Inductive form: Set :=
 | Lit    : nat          -> form
@@
 | Or     : form -> form -> form
-| Implies : form -> form -> form.
+| Implies: form -> form -> form.
 \end{minted}
@@
 Before defining the Kripke semantics, we need to represent the model used to
-interpret a modal logic formula. We encode a frame by a record formed by a
-a type \mintinline{coq}{W : Set} to represent the set of worlds and the
-accessibility relation is denoted by a type \mintinline{coq}{R : W -> W -> Prop}.
+interpret a modal logic formula. We encode a frame by a record formed by
+a type \mintinline{coq}{W: Set} to represent the set of worlds and an
+accessibility relation denoted by a type \mintinline{coq}{R: W -> W -> Prop}.
 \begin{minted}{coq}
-Record Frame : Type := {
-  W : Set;
-  R : W -> W -> Prop;
+Record Frame: Type := {
+  W: Set;
+  R: W -> W -> Prop;
 }.
 \end{minted}
-A model is denoted by another record formed by a frame and a labeling function,
+A model is then denoted by another record, formed by a frame and a labeling function,
 which assign the truth value of a variable at a given world.
 \begin{minted}{coq}
-Record Model : Type := {
-  F : Frame; 
-  v : nat -> (W F) -> Prop; 
+Record Model: Type := {
+  F: Frame; 
+  v: nat -> (W F) -> Prop; 
 }.
@@
 refers to a world from the set of possible worlds in the frame
-\mintinline{coq}{F}. The encoding of the Kripke semantics is done by a Coq function
+\mintinline{coq}{F} (we recall that \mintinline{coq}{W F} is Coq's syntax for
+accessing field \mintinline{coq}{W} from the record \mintinline{coq}{F}). The encoding of the Kripke semantics is done by a Coq function
 that implements the relation of Figure~\ref{fig:kripke}. We omit the code
@@
 \begin{minted}{coq}
-Fixpoint formula_valuation (M: Model)
-                           (w: W (F M))
+Fixpoint formula_valuation (M: Model) (w: W (F M))
                                  (phi: form): Prop :=
   match phi with
-  | Lit     x            => v M x w 
+  | Lit     x
+     => v M x w (* should be inhabited, i.e., true *)
   | Box     psi
@@
                             formula_valuation M w' psi
-  (** some code omitted *)
+  (* some code omitted *)
   end.
@@
 \begin{minted}{coq}
-Inductive axiom : Set :=
+Inductive axiom: Set :=
   | ax1  : form -> form -> axiom
   | ax2  : form -> form -> form  -> axiom
-  (** some code omitted *)
+  (* some code omitted *)
   | axK  : form -> form -> axiom
-  | axPos : form -> form -> axiom
+  | axPos: form -> form -> axiom
   | axT  : form -> axiom
@@
      => [! <> (p \/ q) -> (<> p \/ <> q) !]
-  (** some code omitted *)
+  (* some code omitted *)
   end.
@@
 represented as a predicate
-\mintinline{coq}{A : axiom -> Prop} which parameterizes our deduction type. An
+\mintinline{coq}{A: axiom -> Prop} which parameterizes our deduction type. An
 axiom \mintinline{coq}{a} is valid within a system \mintinline{coq}{A} whenever
@@
   | K_ax2: forall p q r, K (ax2 p q r)
-  (** some code omitted *)
+  (* some code omitted *)
   | K_axK: forall p q, K (axK p q)
@@
 different systems of modal logics uniformly. In order to understand how it
-works, let us review its type (below, remember that \mintinline{coq}{A : axiom -> Prop}
+works, let us review its type (below, remember that \mintinline{coq}{A: axiom -> Prop}
 denotes a system of axioms for modal logics):
@@
 A Coq formalization for modal systems $K$, $T$ and $K4$ was the subject of 
-Doczkal and Smolka's work. Author's interest was in the definition of dynamic
+Doczkal and Smolka's work. The authors' interest was in the definition of dynamic
 and temporal logics. The development of the code establishes the construction
@@
 
-The next table summarizes the formalized modal logics systems by our library
-and related works.
+The following table summarizes the formalized modal logics systems by our library
+and related works:
 \vspace{0.5cm} {
@@
 
-As we can see, as far we know, our work is the unique formalization that covers systems
+As we can see, as far we know, our work is the only formalization that covers systems
 K, D, B, T, K4, K5, S4, S5 and GL. Also, since our deduction is
@@
 $\varphi$ will be true''. The programming languages considered in Pratt's
-original work are composed by an assignment and regular language operators
+original work are composed by assignment and regular language operators
 (union, concatenation and Kleene star)~\cite{Pratt76}.
@@
 shallow embedding in Coq proof assistant and provide two case studies: one is
-the verification of a simple low-level virtual machine, called the list machine
-benchmark and the other is the subset of Sequential Cminor.
+the verification of a simple low-level virtual machine, and the other is the
+subset of Sequential Cminor.
 
 A model for recursive, impredicatively quantified types with mutable references
-was studied by Appel et. al.~\cite{Appel07}. His interest in formalizing such
+was studied by Appel et. al.~\cite{Appel07}. Their interest in formalizing such
 advanced typing concepts was in the context of typed assembly languages.
@@
 mechanized through a shallow embedding in Coq.
+
+Finally, given that a deep embedding for the modal logic was used, it's
+natural to wonder how that might interact with the program extraction feature. The
+reader might refer to~\cite{Kavvos16c}, which presents a survey on the computational
+interpretation of several modal logic systems. Whether or not the current implementation
+is suitable for extraction is left for future work.
 
@@
 In this work, we presented a sound deep embedding of arbitrary
-normal modal logics in Coq. Our approach is able to represent uniformly
-deduction over different modal logic systems by using different axiom sets
+normal modal logics in Coq. Our approach is able to represent deduction
+uniformly over different modal logic systems by using different axiom sets
 as a predicate parameter in the deduction inductive type. Using this generic
 representation of deduction we are able to prove the soundness theorem and
-apply it on proving L\"ob's theorem. All source code in this article has
-been formalized in Coq version 8.15.0 and it is available at~\cite{modal-coq-lib}.
+apply it on proving L\"ob's theorem by assuming fixed points. % All source code in this article has
+% been formalized in Coq version 8.15.0 and it is available at~\cite{modal-coq-lib}.
 
@@
 support for multimodalities (fusion of modal logics) and non-normal modal
-logic systems.
+logic systems, along with improving usability through proof automation.
 

### 2022-08-31 paper/paper.tex
Adding acknolegments
Merge and fix acknowledgements

@@
 refers to a world from the set of possible worlds in the frame
-\mintinline{coq}{F} (we recall that \mintinline{coq}{W F} is Coq's syntax for
-accessing field \mintinline{coq}{W} from the record \mintinline{coq}{F}). The encoding of the Kripke semantics is done by a Coq function
+\mintinline{coq}{F}\footnote{We recall that \mintinline{coq}{W F} is Coq's syntax for
+accessing field \mintinline{coq}{W} from the record \mintinline{coq}{F}.}. The encoding of the Kripke semantics is done by a Coq function
 that implements the relation of Figure~\ref{fig:kripke}. We omit the code
@@
 
+\begin{acks}
+  The present work was supported by Fundação de Amparo à Pesquisa de Minas
+  Gerais (FAPEMIG) under grant APQ-01683-21 and Fundação de Amparo à Pesquisa de
+  Santa Catarina (FAPESC) under grant 2021TR961.
+\end{acks}
+
 \bibliographystyle{ACM-Reference-Format}

### 2022-09-01 paper/paper.tex
Minor fixes

@@
 denotes the modal operator $\Box$ and \mintinline{coq}{Dia}, the operator
-$\Diamond$.
+$\Diamond$. Furthermore, the type \mintinline{coq}{theory} is defined as a list
+of formulas.
+
+\begin{minted}{coq}
+Definition theory := list form.
+\end{minted}
 
@@
 as its constructors. The \mintinline{coq}{deduction} type is parameterized by
-axioms that form a system for modal logics. In this way, our definition can be
-used by several systems of modal logic.
+the set of axioms that form a system for some modal logic. In this way, our definition can be
+used by several systems alike.
 \begin{minted}{coq}
@@
 \begin{minted}{coq}
-Definition instantiate (a: axiom): modalFormula :=
+Definition instantiate (a: axiom): form :=
   match a with

### 2022-09-02 paper/paper.tex
Add rights management code

@@
 %% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{XXXXXXX.XXXXXXX}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Conference acronym 'XX]{Make sure to enter the correct
-  conference title from your rights confirmation emai}{June 03--05,
-  2018}{Woodstock, NY}
+\copyrightyear{2022}
+\acmYear{2022}
+\setcopyright{acmcopyright}\acmConference[SBLP 2022]{XXVI Brazilian Symposium
+on Programming Languages}{October 6--7, 2022}{Virtual Event, Brazil}
+\acmBooktitle{XXVI Brazilian Symposium on Programming Languages (SBLP 2022),
+October 6--7, 2022, Virtual Event, Brazil}
 \acmPrice{15.00}
-\acmISBN{978-1-4503-XXXX-X/18/06}
+\acmDOI{10.1145/3561320.3561329}
+\acmISBN{978-1-4503-9744-5/22/10}
 

### 2022-09-04 paper/paper.tex
Rename s/form/formula/ on the paper

@@
 \begin{minted}{coq}
-Inductive form: Set :=
-| Lit    : nat          -> form
-| Neg    : form -> form
-| Box    : form -> form
-| Dia    : form -> form
-| And    : form -> form -> form
-| Or     : form -> form -> form
-| Implies: form -> form -> form.
+Inductive formula: Set :=
+| Lit    : nat -> formula
+| Neg    : formula -> formula
+| Box    : formula -> formula
+| Dia    : formula -> formula
+| And    : formula -> formula -> formula
+| Or     : formula -> formula -> formula
+| Implies: formula -> formula -> formula.
 \end{minted}
@@
 \begin{minted}{coq}
-Definition theory := list form.
+Definition theory := list formula.
 \end{minted}
@@
 Fixpoint formula_valuation (M: Model) (w: W (F M))
-                                 (phi: form): Prop :=
+                           (phi: formula): Prop :=
   match phi with
@@
 \begin{minted}{coq}
-Inductive deduction (A: axiom -> Prop)
-                           : theory -> form -> Prop :=
+Inductive deduction (A: axiom -> Prop):
+                          theory -> formula -> Prop :=
   (* Premise. *)
   | Prem: forall (t: theory)
-                 (f: form)
+                 (f: formula)
                  (i: nat),
@@
                (a: axiom)
-               (f: form),
+               (f: formula),
         A a -> instantiate a = f -> deduction A t f
@@
   | Mp: forall (t: theory)
-               (f g: form)
+               (f g: formula)
                (d1: deduction A t (f -> g))
@@
   | Nec: forall (t: theory)
-                (f: form)
+                (f: formula)
                 (d1: deduction A t f),
@@
 Inductive axiom: Set :=
-  | ax1  : form -> form -> axiom
-  | ax2  : form -> form -> form  -> axiom
+  | ax1  : formula -> formula -> axiom
+  | ax2  : formula -> formula -> formula  -> axiom
   (* some code omitted *)
-  | axK  : form -> form -> axiom
-  | axPos: form -> form -> axiom
-  | axT  : form -> axiom
-  | axB  : form -> axiom
-  | axK4 : form -> axiom
-  | axD  : form -> axiom
-  | axK5 : form -> axiom.
+  | axK  : formula -> formula -> axiom
+  | axPos: formula -> formula -> axiom
+  | axT  : formula -> axiom
+  | axB  : formula -> axiom
+  | axK4 : formula -> axiom
+  | axD  : formula -> axiom
+  | axK5 : formula -> axiom.
 \end{minted}
@@
 \begin{minted}{coq}
-Definition instantiate (a: axiom): form :=
+Definition instantiate (a: axiom): formula :=
   match a with
@@
            (a: axiom)
-           (f: form),
+           (f: formula),
            A a -> instantiate a = f -> deduction A t f
@@
 low-level programming languages. Authors developed their formalization through a
-shallow embedding in Coq proof assistant and provide two case studies: one is
-the verification of a simple low-level virtual machine, and the other is the
-subset of Sequential Cminor.
+shallow embedding in Coq proof assistant and provide two case studies:
+the verification of a low-level virtual machine and a C-like language.
 

### 2022-09-04 paper/paper.tex
Fix Löb's theorem section

@@
 logical systems with strict implication instead of material implication \cite{Lewis1918}.
-Based in this work~\cite{Lewis1932}, independently of classical propositional logic,
-defined the systems $\mathsf{S1}$--$\mathsf{S5}$, where $\mathsf{S4}$ and $\mathsf{S5}$
+Based in this work, independently of classical propositional logic,
+defined the systems $\mathsf{S1}$--$\mathsf{S5}$~\cite{Lewis1932}, where $\mathsf{S4}$ and $\mathsf{S5}$
 are well known normal modal systems. It was, in fact, G{\"o}del that had the idea 
@@
 \mintinline{coq}{v} from the model \mintinline{coq}{M}. The meaning of
-modalities $\Box$ and $\Diamond$ are represented by a quantifying over worlds
+modalities $\Box$ and $\Diamond$ are represented by quantifying over worlds
 using the accessibility relation $R$. 
@@
 its type stores the subformulas which compose the axiom. The type has
-17 constructors: the first 10 consists encode axioms for propositional logics
-and the last 7 are for different formulas taken as axioms by systems of modal logics.
-Axioms are built using the function \mintinline{coq}{instantiate}, which
-receives an axiom and builds its corresponding formula by pattern matching as
+17 constructors: the first 10 encode axioms for propositional logics
+and, the last 7 are for different formulas which are taken as axioms by common
+systems of modal logic. Axioms are built using the function \mintinline{coq}{instantiate},
+which receives an axiom and builds a corresponding formula by pattern matching as
 follows:
@@
 
-\paragraph{Properties} A key property of the Hilbert deductive system formalized
-in our library is its soundness. If a formula $\varphi$ is derived from a
-context $\Gamma$, i.e. $\Gamma\vdash\varphi$, then $\varphi$ is entailed from
-the set $\Gamma$, i.e. $\Gamma\models\varphi$. The Coq statement of the soundness
-result is:
+\paragraph{Properties} A key property of the Hilbert deductive system as formalized
+in our library is its soundness (through system K). If a formula $\varphi$ is derived from a
+context $\Gamma$, i.e. $\Gamma\vdash\varphi$, then $\varphi$ is entailed within the deep embedding
+from the set of hypotheses $\Gamma$, i.e. $\Gamma\models\varphi$. The Coq statement for the soundness
+result is given as follow:
 
@@
 Theorem soundness:
-  forall G p, (K ; G |-- p) -> (G ||= p).
+  forall G p, (K; G |-- p) -> (G ||= p).
 \end{minted}
@@
   forall M w p q,
-  ((M ' w ||- p) /\
-  (M ' w ||- [! p -> q !])) ->
+  (M ' w ||- p) /\ (M ' w ||- [! p -> q !]) ->
   (M ' w ||- q).
@@
 All these lemmas are direct consequences from the definitions of the Kripke
-semantics and entailment.
+semantics and entailment. We note that system K is used as it doesn't impose
+any restrictions on the frame.
 
@@
 Nevertheless, all Coq's proofs are fairly similar to the respective
-``pen and paper'' proofs that  can be found in the literature. 
-Our work proves the correspondence for the  following properties in
-accessibility relations: reflexivity, transitivity, symmetry,  euclidean,
-seriality, functionality, density, convergency, and converse well-foundness.
+``pen and paper'' proofs found in the literature. 
+Our work proves correspondence for the following properties: reflexivity,
+transitivity, symmetry,  euclidean, seriality, functionality, density, convergency,
+and converse well-foundness.
 
@@
 
-In this section we describe a case study of the defined modal logic library.
-We proved L\"ob theorem in the system K4 with modal fixed points.
+In this section we describe a case study of the defined modal logic library:
+a proof L\"ob theorem in an extension of system K4 with modal fixed points.
 The importance of L\"ob's theorem is due to its use in provability
-logic~\cite{boolos_1994} to abstract details from encodings in G\"odel's
+logic~\cite{boolos_1994}, used to abstract details from encodings in G\"odel's
 incompleteness results.
 
-In order to prove L\"ob's theorem proof we need to encode modal
-fixed points. Let $\varphi(A)$ be a formula with only one propositional variable
-$A$. We define as a modal fixed point of $\varphi$ is a sentence $\Psi$ such that
+In order to state L\"ob's theorem we need a way to encode modal
+fixed points. Formally, let $\varphi(X)$ be a formula with only one propositional variable
+$X$. The modal fixed point of $\varphi$ is then a formula $\Psi$ such that such that
 $\Psi \leftrightarrow \varphi(\Box \Psi)$.
-We represent the modal fixed points by definition \mintinline{coq}{fixed_point},
-which is parameterized by a system for modal logics, \mintinline{coq}{S}, and a
-theory, \mintinline{coq}{g}, and encodes the logical equivalence between a
-formula \mintinline{coq}{p} and its fixpoint \mintinline{coq}{f ([] p)}. 
-\begin{minted}{coq}
-  Definition fixed_point S g
-      := forall f, exists p,
-             (S ; g |-- [! (p <-> f ([] p)) !]).
-\end{minted}
-
-\noindent Using the definition of fixed points, we can state L\"ob's theorem as follows:
-
-\begin{minted}{coq}
-  Theorem Lob: forall P, fixed_point K4 nil ->
-               (K4; nil |-- [! []P -> P !]) ->
-               (K4; nil |-- [! P !]).
+We represent the existence of modal fixed points for all one-variable formulas by a definition \mintinline{coq}{fixed_point},
+which is parameterized by a system of modal logic, \mintinline{coq}{S}, and a
+theory, \mintinline{coq}{g}, and encodes the logical equivalence demonstrating
+that \mintinline{coq}{p} is a modal fixed point for \mintinline{coq}{f}.
+\begin{minted}{coq}
+  Definition fixed_point S g: Prop :=
+    forall f, exists p,
+    (S; g |-- [! (p <-> f ([]p)) !]).
+\end{minted}
+
+\noindent We can now write down L\"ob's theorem as follows:
+
+\begin{minted}{coq}
+  Theorem Lob:
+    forall A, subset K4 A /\ fixed_point A nil ->
+    forall P, (A; nil |-- [! []P -> P !]) ->
+              (A; nil |-- [! P !]).
 \end{minted}
@@
 \noindent The proof starts by assuming an arbitrary formula $P$, the existence of modal fixed
-points and that $\Box P \to P$ is provable in K4. These deduction steps
+points and that $\Box P \to P$ is provable with axioms A. These deduction steps
 are performed by the following tactic:
@@
 \begin{minted}{coq}
-  intros P FP H1.
+  intros A [I FP] P H1.
 \end{minted}
@@
 \begin{minted}{coq}
-  destruct FP with (fun A => [! A -> P !]) as (psi, H2).
+  destruct FP with (fun X => [! X -> P !]) as (psi, H2).
 \end{minted}
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! psi -> []psi -> P !]) as H3.
-  apply modal_ax5 in H2; try repeat constructor ;
-     try eassumption.
+  assert (A; nil |-- [! psi -> []psi -> P !]) as H3.
+  apply modal_ax5 in H2; auto with modal.
 \end{minted}
@@
 generates a new proof hypothesis named \mintinline{coq}{H3}. The
-deduction of this fact is immediate from K4 axioms.
+deduction of this fact is immediate from K4 axioms, of which A is a superset.
+We use some basic automation to move on.
 
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! [](psi -> []psi -> P) !])
+  assert (A; nil |-- [! [](psi -> []psi -> P) !])
     as H4.
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! []psi -> []([]psi -> P) !])
+  assert (A; nil |-- [! []psi -> []([]psi -> P) !])
     as H5.
-  apply modal_axK in H4; try repeat constructor ;
-    try eassumption.
+  apply modal_axK in H4; auto with modal.
 \end{minted}
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! []([]psi -> P) ->
+  assert (A; nil |-- [! []([]psi -> P) ->
                           [][]psi -> []P !]) as H6.
-  eapply Ax with (a := axK ?[X] ?[Y]);
-     try repeat constructor.
+  eapply Ax with (a := axK ?[X] ?[Y]); auto with modal.
+  reflexivity.
 \end{minted}
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! []psi -> [][]psi -> []P !])
+  assert (A; nil |-- [! []psi -> [][]psi -> []P !])
     as H7.
-  eapply modal_compose; try repeat constructor ;
-    try eassumption.
+  eapply modal_compose; eauto with modal.
 \end{minted}
@@
 \noindent Now, we can show that $\vdash\Box\Psi\to\Box\Box\Psi$ using
-the axioms of K4. The next snippet of tactics perform this
+the axiom K4 of A. The next snippet of tactics perform this
 reasoning step.
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! []psi -> [][]psi !]) as H8.
-  apply modal_axK4; try constructor 2.
-\end{minted}
-
-\noindent From $\vdash \Box\Psi \to \Box P$ and $\vdash \Box P \to P$, we can
+  assert (A; nil |-- [! []psi -> [][]psi !]) as H8.
+  apply modal_axK4; auto with modal.
+\end{minted}
+
+\noindent From the previou two hypotheses we gather necessary info to derive
+$\vdash \Box \Psi \to P$, using a propositional axiom.
+
+\begin{minted}{coq}
+  assert (A; nil |-- [! []psi -> []P !]) as H9.
+  eapply modal_ax2; eauto with modal.
+\end{minted}
+
+\noindent Then, from $\vdash \Box\Psi \to \Box P$ and $\vdash \Box P \to P$, we can
 conclude $\vdash \Box \Psi \to P$, by composition of implication as
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! []psi -> P !]) as H10.
-    eapply modal_compose; try repeat constructor ;
-      try eassumption.
+  assert (A; nil |-- [! []psi -> P !]) as H10.
+  eapply modal_compose; eauto with modal.
 \end{minted}
@@
 \noindent Next, we can conclude $\vdash \Box (\Psi \to P) \to \Psi$ from
-$\Psi \leftrightarrow (\Box \Psi \to P)$ and the axioms of K4 using
+$\Psi \leftrightarrow (\Box \Psi \to P)$ using
 the following tactics:
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! ([]psi -> P) -> psi !]) as H11.
-  apply modal_ax6 in H2; try repeat constructor ;
-     try eassumption.
-\end{minted}
-
-\noindent From $\vdash (\Box\Psi \to P) \to \Psi$ and $\vdash \Box \Psi \to P$
+  assert (A; nil |-- [! ([]psi -> P) -> psi !]) as H11.
+  apply modal_ax6 in H2; auto with modal.
+\end{minted}
+
+\noindent Now, from $\vdash (\Box\Psi \to P) \to \Psi$ and $\vdash \Box \Psi \to P$
 we can conclude $\vdash \Psi$ by modus ponens.
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- psi) as H12.
-  eapply Mp ; try eassumption.
+  assert (A; nil |-- psi) as H12.
+  eapply Mp; try eassumption.
 \end{minted}
@@
 \begin{minted}{coq}
-  assert (K4; nil |-- [! []psi !]) as H13.
-  apply Nec ; try eassumption.
+  assert (A; nil |-- [! []psi !]) as H13.
+  apply Nec; try eassumption.
 \end{minted}
@@
 \begin{minted}{coq}
-  eapply Mp ; try eassumption.
+  eapply Mp; try eassumption.
 \end{minted}
@@
 \noindent This concludes the formalization of L\"ob's theorem using
-the developed library.
+the developed library, in a Hilbert style proof.
 
@@
 
-The application of a multimodal separation logic for the semantics of low-level
+The application of a multimodal separation logic for the semantics of low-level languages
 was the subject of Dockins et. al. work~\cite{Dockins08}. In that work, authors
@@
 semantics of the G\"odel-L\"ob logic of provability. All their definitions were
-mechanized through a shallow embedding in Coq.
+also mechanized through a shallow embedding in Coq.
 
@@
 reader might refer to~\cite{Kavvos16c}, which presents a survey on the computational
-interpretation of several modal logic systems. Whether or not the current implementation
+interpretation of several modal logic systems. Whether or not our current implementation
 is suitable for extraction is left for future work.

### 2022-09-04 paper/paper.tex
Fix some typos

@@
     define its Kripke semantics.
-  \item We model the Hilbert deductive systems as a Coq inductive type that is
+  \item We model the Hilbert deductive system as a Coq inductive type that is
     parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL,
     and prove its soundness with respect to Kripke semantics.
-    We also formalize several frame correspondence results.
-  \item We show a proof of L\"ob's theorem using the Hilbert deductive system
+    In such, we also formalize several frame correspondence results.
+  \item As an example of usage, we show a proof of L\"ob's theorem using the deductive system
     developed in our library.
@@
 proving in Coq follows the ideas of the so-called
-``BHK-cor\-res\-pon\-dence''\footnote{Abbreviation of Brouwer, Heyting,
+``BHK-in\-ter\-pre\-ta\-tion''\footnote{Abbreviation of Brouwer, Heyting,
   Kolmogorov, de Bruijn and Martin-L\"of Correspondence. This is also
@@
 However, writing a proof term whose type is that of a logical formula
-can be a hard task, even for very simple propositions.  In order to
+can be challenging, even for very simple propositions.  In order to
 make the writing of complex proofs easier, Coq provides
@@
 about truth across various --- abstract --- worlds. In such logics, a
-proposition may be true in some world, but false in another one. The versions of
-modal logics considered in this work extends traditional propositional logics
+proposition may be true in some world, but false in another one. The systems of
+modal logic considered in this work extend traditional propositional logic
 with two operators on propositions: $\Box$ (box) and $\Diamond$ (diamond).
-The syntax of modal logic formulas are defined by the following context free
+The syntax for modal logic formulas are defined by the following context free
 grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$ and
@@
   \begin{array}{lcl}
-    \varphi & ::= & A\,|\,\neg\varphi\,|\,\varphi\circ\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
+    \varphi & ::= & \mathcal{P}\,|\,\neg\varphi\,|\,\varphi\circ\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
   \end{array}  
@@
 Intuitively, the semantics of the operator $\Box$ is similar to an universal
-quantifier, since it express \emph{necessity}: $\Box\,\varphi$ is true at the
-current world iff $\varphi$ is true at all worlds accessible from the current
-world. Similarly, the semantics of operator
-$\Diamond$ is that a formula $\Diamond\varphi$ is true at the current
+quantifier, since it expresses \emph{necessity}: $\Box\,\varphi$ is true in the
+current world iff $\varphi$ is true in every world accessible from the current
+one. Similarly, the semantics of operator
+$\Diamond$ is that a formula $\Diamond\varphi$ is true in the current
 world iff $\varphi$ is \emph{possible}, i.e. true in at least some world

### 2022-09-05 paper/paper.tex
Fix syntax and def. of entailment

@@
 The syntax for modal logic formulas are defined by the following context free
-grammar (where $\circ \in\{\land,\lor,\to,\leftrightarrow\}$ and
+grammar (where
 $\mathcal{P} = \{A,B,C,...\}$ is the enumerable set of propositional variables):
@@
   \begin{array}{lcl}
-    \varphi & ::= & \mathcal{P}\,|\,\neg\varphi\,|\,\varphi\circ\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
+    \varphi & ::= & \mathcal{P}\,|\,\neg\varphi\,|\,\varphi\land\varphi\,|\,\varphi\lor\varphi\,|\,\varphi\to\varphi\,|\,\Box\varphi\,|\,\Diamond\varphi
   \end{array}  
@@
 world iff $\varphi$ is \emph{possible}, i.e. true in at least some world
-accessible from the current world.
+accessible from the current world. We also write
+$\phi_1\leftrightarrow\phi_2$ to mean $\phi_1\to\phi_2\land\phi_2\to\phi_1$.
 
@@
 \begin{enumerate}
-  \item $\mathcal{M}\,;\,w\not\models \bot$ and $\mathcal{M}\,;\,w\models \top$;
+  % \item $\mathcal{M}\,;\,w\not\models \bot$ and $\mathcal{M}\,;\,w\models \top$;
   \item $\mathcal{M}\,;\,w\models A$ iff $\mathcal{V}(A,w) = T$;
@@
 
-Using the semantics, we can define the concept of \emph{entailment}: Given a set
+Using the semantics, we can define the concept of \emph{entailment}: given a set
 of formulas $\Gamma$ and a formula $\varphi$, we say that $\Gamma$ entails
-$\varphi$, written $\Gamma\models\varphi$, iff $\forall \varphi' \in\Gamma .
-\mathcal{M} \models \varphi' \to \mathcal{M} \models \varphi$, for some model $\mathcal{M}$.
-
-In this work, we consider Hilbert deductive systems for modal logics.
+$\varphi$, written $\Gamma\models\varphi$, iff $\forall \mathcal{M} . (\forall \varphi' \in\Gamma .
+\mathcal{M} \models \varphi') \to \mathcal{M} \models \varphi$.
+
+In this work, we consider Hilbert deductive systems for modal logic.
 Essentially, the considered system adds axioms for the modalities

### 2022-09-05 paper/paper.tex
Fix notation (plus typos)

@@
 is formed by a frame $\mathcal{F}$ and a total labeling function
-$\mathcal{V} : \mathcal{P}\times W\to \{F,T\}$, which assigns
+$\mathcal{V} : \mathcal{P}\to 2^W$, which assigns
 a truth value to a propositional variable at a given world, i.e. when $w \in \mathcal{V}(A)$
@@
   % \item $\mathcal{M}\,;\,w\not\models \bot$ and $\mathcal{M}\,;\,w\models \top$;
-  \item $\mathcal{M}\,;\,w\models A$ iff $\mathcal{V}(A,w) = T$;
+  \item $\mathcal{M}\,;\,w\models A$ iff $w \in \mathcal{V}(A)$;
   \item $\mathcal{M}\,;\,w\models \neg\varphi$ iff $\mathcal{M}\,;\,w\not\models\varphi$;
@@
     \textbf{Possibility} & \Diamond (\varphi_1 \lor \varphi_2) \to
-                           (\Diamond \varphi_2 \lor \Diamond \varphi_2)\\
+                           (\Diamond \varphi_1 \lor \Diamond \varphi_2)\\
   \end{array}
@@
 \end{figure}
-The set of axioms and the two previous rules forms a sound and complete Hilbert-style
+The set of axioms and the two previous rules form a sound and complete Hilbert-style
 deductive system for modal logics~\cite{blackburn}.
@@
 \rangle$ in which the relation $R$ is reflexive: $\forall w \in W. wRw$, i.e.
-every world is accessible from itself. In such frame when $\Box A$ holds at
-a world $w$, $w$ itself is among the accessible worlds at which $A$ must be true.
+every world is accessible from itself. In such frame when $\Box A$ holds in
+a world $w$, $w$ itself is among the accessible worlds in which $A$ must be true.
 In this way, if the accessibility relation $R$ is reflexive, then whatever world

### 2022-09-05 paper/paper.tex
Fix a few more typos

@@
 is formed by a frame $\mathcal{F}$ and a total labeling function
-$\mathcal{V} : \mathcal{P}\to 2^W$, which assigns
-a truth value to a propositional variable at a given world, i.e. when $w \in \mathcal{V}(A)$
+$\mathcal{V} : \mathcal{P}\to 2^W$, which indicates the set of worlds where each propositional variable holds, i.e. when $w \in \mathcal{V}(A)$
 for some $A \in\mathcal{P}$, we say that $A$ holds in $w$. The satisfiability of a formula
@@
 Most constructors of the type have an immediate meaning, since they represent
-connectives of propositional logics. The constructor \mintinline{coq}{Lit}
+connectives of propositional logic. The constructor \mintinline{coq}{Lit}
 is used to construct propositional variables; constructor \mintinline{coq}{Box}
@@
 A model is then denoted by another record, formed by a frame and a labeling function,
-which assign the truth value of a variable at a given world.
+which assigns the truth value of a variable at a given world.
 \begin{minted}{coq}
@@
 \mintinline{coq}{v} from the model \mintinline{coq}{M}. The meaning of
-modalities $\Box$ and $\Diamond$ are represented by quantifying over worlds
+modalities $\Box$ and $\Diamond$ is represented by quantifying over worlds
 using the accessibility relation $R$. 
@@
 In order to ease the task of writing formulas and its semantics properties, we
-define Coq notations that mimics the standard way of writing them.
+define Coq notations that mimic the standard way of writing them.
 
@@
                (f g: formula)
-               (d1: deduction A t (f -> g))
+               (d1: deduction A t (Implies f g))
                (d2: deduction A t f),
@@
 represent rules for hypothesis, modus ponens and generalization, respectively.
-Before discussing the \mintinline{coq}{Ax}, we need to discuss how axioms are
+Before discussing the \mintinline{coq}{Ax} constructor, we need to discuss how axioms are
 represented in our library. The inductive type \mintinline{coq}{axiom} specifies
@@
 \end{minted}
-Each constructor of type \mintinline{coq}{axiom} represents an axiom and
+Each constructor of type \mintinline{coq}{axiom} represents an axiom scheme and
 its type stores the subformulas which compose the axiom. The type has
-17 constructors: the first 10 encode axioms for propositional logics
+17 constructors: the first 10 encode axioms for propositional logic
 and, the last 7 are for different formulas which are taken as axioms by common
@@
 included in the system. Our approach is to define predicates
-which denote each system. The simplest set of axioms is usually refereed as
+which denote each system. The simplest set of axioms is usually referred as
 \textbf{K} and it includes the axiom $\Box(A \to B) \to (\Box A \to \Box B)$
 and all the axioms for propositional logic. We represent system \textbf{K}
-by the following inductive type:
+by the following inductive predicate:
 \begin{minted}{coq}
@@
 \end{minted}
-Note that the first parameter of constructor \mintinline{coq}{Ax} is a value
+Note that the fourth parameter of constructor \mintinline{coq}{Ax} is a value
 of type \mintinline{coq}{A a}, which is only inhabited if the predicate
@@
 using function \mintinline{coq}{instantiate}. In this way, we are able to
-represent all systems of axioms for modal logics using a single constructor
+represent all systems of axioms for modal logics using a single type
 for its axioms.
@@
 implies the correspondent axiom is usually  direct, the inverse often needs
-more effort. On the Coq proofs developed for this work,  the former case is
+more effort. In the Coq proofs developed for this work,  the former case is
 mostly unfoldings, simplifications and applications of premises
@@
 requires the construction of a witness valuation function. 
-Nevertheless, all Coq's proofs are fairly similar to the respective
+Nevertheless, all Coq proofs are fairly similar to the respective
 ``pen and paper'' proofs found in the literature. 
@@
 
-\paragraph{Formalizations of modal logics on proof assistants}
-
-The formalization of epistemic modal logic in Coq was the subject of Wind's
+\paragraph{Formalizations of modal logics in proof assistants}
+
+The formalization of epistemic modal logic in Coq was the subject of de Wind's
 work~\cite{dewind}.
@@
 modeling of two well known puzzles.
-Wind's formalization was separated into some components. The first consists in
+De Wind's formalization was separated into some components. The first consists in
 defining the syntax of formulas and its Kripke semantics. Next, the author uses
@@
 follows the approach due to Henkin for S5 completeness and it depends on
-the \texttt{mathlib}, the Lean's library of mathematical results~\cite{mathlib20}.
-
-Maggesi et. al. developed a formalization of the completeness proof for GL using
+\texttt{mathlib}, the Lean's library of mathematical results~\cite{mathlib20}.
+
+Maggesi et al. developed a formalization of the completeness proof for GL using
 the HOL light theorem prover~\cite{Maggesi21}. The authors argue that using the
 standard Henkin strategy does not work in GL without modifications. The core idea of
-Maggesi et. al. strategy was building maximal consistent lists in terms of
+Maggesi et al. strategy was building maximal consistent lists in terms of
 components of the given formula non-proven in the calculus.
@@
                                & K & D & B & T & K4 & K5 & S4 & S5 & GL\\ \hline
-      Wind et. al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X & \\ \hline
-      Doczkal et. al.~\cite{saarland}   & X &   &   & X & X  &    &    &  &  \\ \hline
-      Benzmüller et. al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X &  \\ \hline
+      Wind et al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X & \\ \hline
+      Doczkal et al.~\cite{saarland}   & X &   &   & X & X  &    &    &  &  \\ \hline
+      Benzmüller et al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X &  \\ \hline
       Bentzen~\cite{Bentzen2021} & X & & & X & & X & X & X & \\ \hline
-      Maggesi et. al.~\cite{Maggesi21} & X & & & & & & & & X \\ \hline
+      Maggesi et al.~\cite{Maggesi21} & X & & & & & & & & X \\ \hline
       This work         & X & X & X & X & X  &  X &  X & X  & X \\ \hline
@@
 Dynamic logic was introduced by Pratt in late 70's as modal logic for reasoning
-about programs~\cite{Pratt76,Pratt80}. The key idea on dynamic logic is to express the
+about programs~\cite{Pratt76,Pratt80}. The key idea of dynamic logic is to express the
 Hoare triples by a modality meaning ``after a program $p$ executes, formula
@@
 In that work, modalities are used to represent types for mobile code
-and remote addresses. Murphy et. al defined the language by interpreting
+and remote addresses. Murphy et al defined the language by interpreting
 the possible worlds of the intutionistic modal logic IS5 to be nodes on a
@@
 The application of a multimodal separation logic for the semantics of low-level languages
-was the subject of Dockins et. al. work~\cite{Dockins08}. In that work, authors
+was the subject of Dockins et al.'s work~\cite{Dockins08}. In that work, authors
 combine several modal operators that do not need to satisfy the same frame
@@
 A model for recursive, impredicatively quantified types with mutable references
-was studied by Appel et. al.~\cite{Appel07}. Their interest in formalizing such
+was studied by Appel et al.~\cite{Appel07}. Their interest in formalizing such
 advanced typing concepts was in the context of typed assembly languages.
@@
 as a predicate parameter in the deduction inductive type. Using this generic
-representation of deduction we are able to prove the soundness theorem and
-apply it on proving L\"ob's theorem by assuming fixed points. % All source code in this article has
+representation of deduction we are able to prove a soundness theorem, and, in
+order to demonstrate the applicability of the proposed work, prove L\"ob's theorem
+by assuming fixed points. % All source code in this article has
 % been formalized in Coq version 8.15.0 and it is available at~\cite{modal-coq-lib}.

### 2022-09-05 paper/paper.tex
Remove grant number (but why?)

@@
   Gerais (FAPEMIG) under grant APQ-01683-21 and Fundação de Amparo à Pesquisa de
-  Santa Catarina (FAPESC) under grant 2021TR961.
+  Santa Catarina (FAPESC). %under grant 2021TR961.
 \end{acks}

### 2022-09-05 paper/paper.tex
Add repository location and fix example layout

@@
 on Section~\ref{sec:related} and Section~\ref{sec:conclusion} draws some conclusions and
-provide some pointers for future works.
-
-All source code in this article has been formalized in Coq version 8.15.0 and it
-is available at~\cite{modal-coq-lib}.
+provide some pointers for future work.
+
+All source code in this article has been formalized in Coq version 8.15 and
+is available at \url{https://github.com/funcao/ModalLibrary/} under a permissive free software license.
 
@@
    Proof.
-       intros H H' HA. apply H'. apply H. assumption. 
+       intros H H' HA.
+       apply H'.
+       apply H.
+       assumption. 
    Qed.
@@
 \begin{minted}{coq}
-Definition example :(A -> B) -> (B -> C) -> A -> C :=
- fun (H: A -> B)
-     (H': B -> C)
-     (HA: A) => H' (H HA).
+Definition example: (A -> B) -> (B -> C) -> A -> C :=
+  fun (H: A -> B) (H': B -> C) (HA: A) => H' (H HA).
 \end{minted}
@@
 with two operators on propositions: $\Box$ (box) and $\Diamond$ (diamond).
-The syntax for modal logic formulas are defined by the following context free
+The syntax for modal logic formulas is defined by the following context free
 grammar (where
-$\mathcal{P} = \{A,B,C,...\}$ is the enumerable set of propositional variables):
+$\mathcal{P} = \{A,B,C,...\}$ is the countable set of propositional variables):
 \[
@@
 world iff $\varphi$ is \emph{possible}, i.e. true in at least some world
-accessible from the current world. We also write
+accessible from the current world. We the remaining of the paper, we also write
 $\phi_1\leftrightarrow\phi_2$ to mean $\phi_1\to\phi_2\land\phi_2\to\phi_1$.

### 2024-01-10 paper/COLA/paper.tex
Adding paper to COLA

--- 2022-09-05 paper/paper.tex
@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf]{acmart}
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    Bib\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\copyrightyear{2022}
-\acmYear{2022}
-\setcopyright{acmcopyright}\acmConference[SBLP 2022]{XXVI Brazilian Symposium
-on Programming Languages}{October 6--7, 2022}{Virtual Event, Brazil}
-\acmBooktitle{XXVI Brazilian Symposium on Programming Languages (SBLP 2022),
-October 6--7, 2022, Virtual Event, Brazil}
-\acmPrice{15.00}
-\acmDOI{10.1145/3561320.3561329}
-\acmISBN{978-1-4503-9744-5/22/10}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% For managing citations, it is recommended to use bibliography
-%% files in BibTeX format.
-%%
-%% You can then either use BibTeX with the ACM-Reference-Format style,
-%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
-%% support for advanced citation of software artefact from the
-%% biblatex-software package, also separately available on CTAN.
-%%
-%% Look at the sample-*-biblatex.tex files for templates showcasing
-%% the biblatex styles.
-%%
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
+% Template for Elsevier CRC journal article
+% version 1.2 dated 09 May 2011
+
+% This file (c) 2009-2011 Elsevier Ltd.  Modifications may be freely made,
+% provided the edited file is saved under a different name
+
+% This file contains modifications for Procedia Computer Science
+% but may easily be adapted to other journals
+
+% Changes since version 1.1
+% - added "procedia" option compliant with ecrc.sty version 1.2a
+%   (makes the layout approximately the same as the Word CRC template)
+% - added example for generating copyright line in abstract
+
+%-----------------------------------------------------------------------------------
+
+%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
+%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
+%% Further resources available at http://www.elsevier.com/latex
+
+%-----------------------------------------------------------------------------------
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                          %%
+%% Important note on usage                                  %%
+%% -----------------------                                  %%
+%% This file should normally be compiled with PDFLaTeX      %%
+%% Using standard LaTeX should work but may produce clashes %%
+%%                                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
+%% Add the 'procedia' option to approximate to the Word template
+%\documentclass[3p,times,procedia]{elsarticle}
+\documentclass[3p,times]{elsarticle}
+
+%% The `ecrc' package must be called to make the CRC functionality available
+\usepackage{ecrc}
+
+%% The ecrc package defines commands needed for running heads and logos.
+%% For running heads, you can set the journal name, the volume, the starting page and the authors
+
+%% set the volume if you know. Otherwise `00'
+\volume{00}
+
+%% set the starting page if not 1
+\firstpage{1}
+
+%% Give the name of the journal:
+\journalname{Journal of Computer Languages}
+
+%% Give the author list to appear in the running head
+%% Example \runauth{C.V. Radhakrishnan et al.}
+\runauth{A.A. da Silveira et. al.}
+
+%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
+%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
+%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
+%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo
+
+%% Give the abbreviation of the Journal.  Contact the journal editorial office if in any doubt
+\jid{COLA}
+
+%% Give a short journal name for the dummy logo (if needed)
+\jnltitlelogo{COLA}
+
+%% Provide the copyright line to appear in the abstract
+%% Usage:
+%   \CopyrightLine[<text-before-year>]{<year>}{<restt-of-the-copyright-text>}
+%   \CopyrightLine[Crown copyright]{2011}{Published by Elsevier Ltd.}
+%   \CopyrightLine{2011}{Elsevier Ltd. All rights reserved}
+\CopyrightLine{2011}{Published by Elsevier Ltd.}
+
+%% Hereafter the template follows `elsarticle'.
+%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.
+
+%% Elsevier CRC generally uses a numbered reference style
+%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
+%% If using BibTeX, use the style file elsarticle-num.bst
+
+%% End of ecrc-specific commands
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amssymb}
 \usepackage{minted}
@@
 \usepackage{multirow}
-
+\usepackage{hyperref}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers after \end{frontmatter}.
+%% \usepackage{lineno}
+
+%% natbib.sty is loaded by default. However, natbib options can be
+%% provided with \biboptions{...} command. Following options are
+%% valid:
+
+%%   round  -  round parentheses are used (default)
+%%   square -  square brackets are used   [option]
+%%   curly  -  curly braces are used      {option}
+%%   angle  -  angle brackets are used    <option>
+%%   semicolon  -  multiple citations separated by semi-colon
+%%   colon  - same as semicolon, an earlier confusion
+%%   comma  -  separated by comma
+%%   numbers-  selects numerical citations
+%%   super  -  numerical citations as superscripts
+%%   sort   -  sorts multiple citations according to order in ref. list
+%%   sort&compress   -  like sort, but also compresses numerical citations
+%%   compress - compresses without sorting
 %%
-%% end of the preamble, start of the body of the document source.
+%% \biboptions{comma,round}
+
+% \biboptions{}
+
+% if you have landscape tables
+\usepackage[figuresright]{rotating}
+
+% put your own definitions here:
+%   \newcommand{\cZ}{\cal{Z}}
+%   \newtheorem{def}{Definition}[section]
+%   ...
+
+% add words to TeX's hyphenation exception list
+%\hyphenation{author another created financial paper re-commend-ed Post-Script}
+
+% declarations for front matter
+
 \begin{document}
+
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for the associated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for the associated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for the associated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
 %%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-\title{A Sound Deep Embedding of Arbitrary Normal Modal Logics in Coq}
-
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
+\dochead{}
+%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
+%% \dochead can also be used to include a conference title, if directed by the editors
+%% e.g. \dochead{17th International Conference on Dynamical Processes in Excited States of Solids}
+
+\title{A Sound and Complete Deep Embedding of Arbitrary Normal Modal Logics in Coq}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{<author name>}
+%% \address[label1]{<address>}
+%% \address[label2]{<address>}
 
 \author{Ariel Agne da Silveira}
-\affiliation{\institution{Federal University of Ouro Preto}\country{}}
-\email{ariel.silveira@aluno.ufop.edu.br}
+\ead{ariel.silveira@aluno.ufop.edu.br}
+\address{Federal University of Ouro Preto}
 
 \author{Miguel Nunes}
-\affiliation{\institution{Santa Catarina State University}\country{}}
-\email{miguel.alfredo.nunes@gmail.com}
+\ead{miguel.alfredo.nunes@gmail.com}
+\address{Santa Catarina State University}
 
 \author{Karina Roggia}
-\affiliation{\institution{Santa Catarina State University}\country{}}
-\email{karina.roggia@udesc.br}
+\ead{karina.roggia@udesc.br}
+\address{Santa Catarina State University}
 
 \author{Paulo Torrens}
-\affiliation{\institution{Santa Catarina State University}\country{}}
-\email{paulotorrens@gnu.org}
+\ead{paulotorrens@gnu.org}
+\address{Santa Catarina State University}
 
 \author{Rodrigo Ribeiro}
-\affiliation{\institution{Federal University of Ouro Preto}\country{}}
-\email{rodrigo.ribeiro@ufop.edu.br}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-\renewcommand{\shortauthors}{da Silveira et al.}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
+\ead{rodrigo.ribeiro@ufop.edu.br}
+\address{Federal University of Ouro Preto}
+
+%
 \begin{abstract}
@@
 
+\begin{keyword}
+Modal logic \sep soundness \sep completeness \sep Coq proof assistant
+\end{keyword}
+
+\end{frontmatter}
+
 %%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{Modal logics, soundness, Coq proof assistant}
-
-\maketitle
+%% Start line numbering here if you want
+%%
+% \linenumbers
 
@@
   \item We model the Hilbert deductive system as a Coq inductive type that is
-    parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL,
-    and prove its soundness with respect to Kripke semantics.
-    In such, we also formalize several frame correspondence results.
+    parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL.
+  \item We formally verify several frame correspondence results.
+  \item We prove soundness and completeness for the Hilbert deductive system. For the
+        completeness proof we follow the so-called Henkin-style and it closely follows
+        the proofs presented in classical modal logics textbooks~\cite{Blackburn-ModalLogics}.
   \item As an example of usage, we show a proof of L\"ob's theorem using the deductive system
@@
 
+This article is an extended version of our SBLP 2022 work~\cite{Silveira22}. In this paper we include
+a detailed discussion about our formalization of a completeness proof for the Hilbert deductive
+system following the strategies found in traditional modal logic books~\cite{Blackburn-ModalLogics}.
 The rest of this work is organized as follows: Section~\ref{sec:background} provides
@@
 
-\paragraph{An overview of Coq proof assistant.}
+\subsection{An overview of Coq proof assistant.}
 
@@
 
-However, writing a proof term whose type is that of a logical formula
-can be challenging, even for very simple propositions.  In order to
-make the writing of complex proofs easier, Coq provides
+In order to make writing proofs easier, Coq provides
 \emph{tactics}, which are commands that can be used to construct proof
-terms in a more user-friendly way.
-
-As a tiny example, consider the task of proving the following 
+terms in a more user-friendly way. As an example, consider the task of proving the following 
 formula of the propositional logic:
@@
 \end{minted}
-In the previous source code piece, we have defined a Coq section named
+In the above, we have defined a Coq section named
 \mintinline{coq}{EXAMPLE}\footnote{In Coq, we can use sections to delimit the
@@
 \mintinline{coq}{Prop}). Tactic \mintinline{coq}{intros} introduces variables
-\mintinline{coq}{H}, \mintinline{coq}{H'} and \mintinline{coq}{HA} into the (typing) context,
-respectively with types \mintinline{coq}{A -> B}, \mintinline{coq}{B -> C} and
-\mintinline{coq}{A} and leaves goal \mintinline{coq}{C} to be proved. Tactic
+\mintinline{coq}{H}, \mintinline{coq}{H'} and \mintinline{coq}{HA} into the typing context,
+respectively with types \mintinline{coq}{A -> B}, \mintinline{coq}{B -> C}, and
+\mintinline{coq}{A}, and leaves goal \mintinline{coq}{C} to be proved. Tactic
 \mintinline{coq}{apply}, used with a term \mintinline{coq}{t}, generates goal
@@
 changes the goal to \mintinline{coq}{A}. Tactic \mintinline{coq}{assumption}
-traverses the typing context to find a hypothesis that matches the goal.
+traverses the typing context to find a hypothesis that matches the goal,
+concluding the proof.
 
 We define next a proof of the previous propositional logical formula
-that, in contrast to the previous proof, that was built using tactics
-(\mintinline{coq}{intros}, \mintinline{coq}{apply} and \mintinline{coq}{assumption}), is coded
+that, in contrast to the previous proof, that was built using tactics, is coded
 directly as a function:
@@
 
-\paragraph{An overview of modal logic.} Modal logics are designed for reasoning
+\subsection{An overview of modal logic.} Modal logics are designed for reasoning
 about truth across various --- abstract --- worlds. In such logics, a
@@
 \end{tabular}
+\centering
 \end{figure}
 The set of axioms and the two previous rules form a sound and complete Hilbert-style
-deductive system for modal logics~\cite{blackburn}.
+deductive system for modal logics~\cite{blackburn}. Given a formula $\varphi$ and a
+set of hypothesis $\Gamma$, the soundness property ensures that
+whenever $\varphi$ is deduced from $\Gamma$, then $\Gamma\models \varphi$. On the
+other hand, completeness guarantee that if $\Gamma\models\varphi$, then $\varphi$ is
+provable from $\Gamma$ using the deductive system. While the soundness argument can
+be proved by a immediate inductive argument over derivations of $\Gamma\vdash\varphi$,
+the same strategy does not apply to the completeness property: the key step in proving
+completeness is to produce a \emph{cannonical model} using maximal consistent sets
+of formulas.
 
@@
 \end{tabular}
+\centering
 \end{figure}
@@
 
-\paragraph{Syntax and semantics of modal logics}
+\subsection{Syntax and semantics of modal logics}
+
 To implement our library, we use a \emph{deep embedding} of
@@
 denotes the modal operator $\Box$ and \mintinline{coq}{Dia}, the operator
-$\Diamond$. Furthermore, the type \mintinline{coq}{theory} is defined as a list
-of formulas.
-
-\begin{minted}{coq}
-Definition theory := list formula.
+$\Diamond$. Furthermore, the type \mintinline{coq}{theory} is defined as a function
+between \mintinline{coq}{formula} and Coq's \mintinline{coq}{Prop}s.
+
+\begin{minted}{coq}
+Definition theory := formula -> Prop.
 \end{minted}
@@
 
-\paragraph{The Hilbert deductive system}
+\subsection{The Hilbert deductive system}
+
 Basically, a Hilbert-style deductive system allows the use of four
@@
   | Prem: forall (t: theory)
-                 (f: formula)
-                 (i: nat),
-          (nth_error t i = Some f) -> deduction A t f
+                 (f: formula),
+        t f -> deduction A t f
   (* Axiom. *)
@@
                 (d1: deduction A t f),
-         deduction A t (Box f).
+         deduction A t ([] f).
 \end{minted}
@@
 
-\paragraph{Systems of modal logics} In order to use the deductive system to
+\subsubsection{Systems of modal logics}
+
+In order to use the deductive system to
 check the provability of formulas, we need to specify the axioms that are
@@
 
-\paragraph{Properties} A key property of the Hilbert deductive system as formalized
+\subsection{Properties}
+
+In this section we describe the main properties of the Hilbert deductive systems formalized
+in our library. We start with the soundness property (Section~\ref{sec:soundness}).
+Section~\ref{sec:completeness} discuss the formalization of the completeness result. Finally,
+Section~\ref{sec:correspondence} deals with frame correspondence proofs.
+
+\subsubsection{Soundness}\label{sec:soundness}
+
+A key property of the Hilbert deductive system as formalized
 in our library is its soundness (through system K). If a formula $\varphi$ is derived from a
@@
 \begin{minted}{coq}
-Theorem soundness:
-  forall G p, (K; G |-- p) -> (G ||= p).
+Theorem soundness : forall G p, (K; G |-- p) -> (G ||= p).
 \end{minted}
@@
 Lemma Modus_Ponens_soundness:
-  forall M w p q,
-  (M ' w ||- p) /\ (M ' w ||- [! p -> q !]) ->
-  (M ' w ||- q).
+  forall M w p q, (M ' w ||- p) /\ (M ' w ||- [! p -> q !]) -> (M ' w ||- q).
 \end{minted}
@@
 
-\paragraph{Frame Correspondence} 
+\subsubsection{Completeness}\label{sec:completeness}
+
+In this section, we describe our formalization of the completeness result for the
+Hilbert proof system with respect to the Kripke semantics. The proof consists of two
+basic steps: 1) the proof of the existence of maximal consistent sets and; 2) the
+construction of a cannonical model from the maximal consistent set. We describe these
+steps separately.
+
+\paragraph{Maximal consistent sets} The first step on the completeness proof is the construction of
+a \emph{maximal consistent sets}, $\Delta$, of formulas for $\Gamma$ and $\varphi$.  The $\Delta$ set
+is defined as follows:
+
+\[
+  \begin{array}{lcl}
+    \Delta_{0}    & = & \Gamma \cup \{\neg \varphi\}\\
+    \Delta_{n + 1} & = & \left \{
+                         \begin{array}{ll}
+                           \Delta_{n} \cup \{ \varphi_{n + 1}\}      & \emph{if }\Delta_{n} \cup \{\varphi_{n + 1}\} \emph{ is consistent.}\\
+                           \Delta_{n} \cup \{\neg \varphi_{n  + 1}\} & \emph{otherwise}\\
+                         \end{array}
+                         \right. \\ & & \\
+    \Delta        & = & \bigcup_{n \in \mathbb{N}} \Delta_{n}
+  \end{array}
+\]
+In order to define $\Delta$, we first define an iterative version $\Delta_{i}$ by letting $\Delta_{0}$ be the
+set formed by $\Gamma \cup \{\neg\varphi\}$. Further steps in the construction of $\Delta_{i}$ are made by
+the inclusion of formulas $\varphi_{i}$ which keep the set consistent.
+
+We start with some basic definitions. We say that a theory is consistent if it is not possible to deduce
+$\varphi$ and $\neg \varphi$ for any formula $\varphi$. A set of formulas is maximal if it holds $\varphi$ or $\neg \varphi$
+for an arbitrary formula $\varphi$. We define such concepts using the following Coq predicates:
+
+\begin{minted}{coq}
+Definition Consistent: Prop := forall p, ~(A; G |-- [! p /\ ~p !]).
+Definition Maximal: Prop := forall p, G p \/ G [! ~p !].
+\end{minted}
+
+Which states that it is not possible to deduce \verb|p| and \verb|~ p|, for any formula \verb|p|.
+In order to define maximal consistent sets, we use the following representation of a theory
+subset predicate:
+
+\begin{minted}{coq}
+Definition Subset (A B : theory): Prop := forall t, A t -> B t.
+\end{minted}
+
+The completeness proof rely on the fact that the set of modal logic formulas is countable. We model
+this fact by two functions \mintinline{coq}{encode :: formula -> nat} and \mintinline{coq}{decode :: nat -> option formula}
+which satisfies the following property: $\forall \varphi . \texttt{decode(encode }\varphi\texttt{)} = \varphi$.
+
+Using the previously defined predicates, we encode the action of including a formula or its
+negation in a set using the type \mintinline{coq}{Insert}:
+
+\begin{minted}{coq}
+Inductive Insert: option formula -> theory -> theory :=
+| Insert_skip: forall D, Subset D (Insert None D)
+| Insert_valid:
+  forall D p, Consistent A (Extend p D) -> Subset (Extend p D) (Insert (Some p) D)
+| Insert_invalid:
+  forall D p, ~ Consistent A (Extend p D) -> Subset (Extend [! ~p !] D) (Insert (Some p) D).
+\end{minted}
+
+The first constructor states that when we do not insert a formula in a theory,
+we do not change it. The second constructor includes a formula in a theory if
+adding that formula keeps the theory consistent and the final constructor considers
+the situation when the inclusion of a formula \verb|p| makes the theory inconsistent.
+In this case, we add the negation of \verb|p| to the theory.
+
+Using the type \mintinline{coq}{Insert}, we can define the $\Delta_{i}$ sets using the
+following function:
+
+\begin{minted}{coq}
+ Fixpoint Delta (n: nat): theory :=
+   match n with
+   | 0   => G
+   | S n' => Insert (decode n') (Delta n')
+   end.
+\end{minted}
+
+In order to produce the $\Delta_{i}$, we need to include the $i$-th formula (built using the function \mintinline{coq}{decode})
+in the theory $\Delta_{i - 1}$. Finally, the maximal set is defined as:
+\begin{minted}{coq}
+Definition Max : theory := UnionOf Delta.
+Definition UnionOf (A : nat -> theory) : theory := fun t => exists i, A i t.
+\end{minted}
+Function \mintinline{coq}{UnionOf} is a immediate encoding the indexed set union operator using Coq. An important property
+of this construction of the maximal set is that it only increases the number of elements. This fact is proved by
+the followed lemma, which has an immediate proof using the previous definitions.
+\begin{minted}{coq}
+Lemma subset_of_insert: forall D p, Subset D (Insert p D).
+\end{minted}
+Next, we show that the set \mintinline{coq}{Max} is such that either \verb|p| or \verb|~ p| is an element of \verb|Max|.
+\begin{minted}{coq}
+Lemma max_is_maximal: forall p, Max [! p !] \/ Max [! ~p !].
+\end{minted}
+The next step in the proof is showing that \mintinline{coq}{Max} is consistent. In order to prove this fact,
+we use some lemmas. First, we show, by case analysis on formulas, that inserting an formula into a theory
+preserves consistency:
+\begin{minted}{coq}
+Lemma insert_preserves_consistency: forall D p, Consistent A D -> Consistent A (Insert p D).
+\end{minted}
+and using the previous lemma, we prove that operation \mintinline{coq}{Delta n} preserves consistency, by induction on \verb|n|.
+\begin{minted}{coq}
+Lemma delta_is_consistent: Consistent A G -> forall n, Consistent A (Delta n).
+\end{minted}
+Next, the proof that \mintinline{coq}{Max} is consistent is a corollary of the previously stated results.
+\begin{minted}{coq}
+Corollary max_is_consistent: Consistent A G -> Consistent A Max.
+\end{minted}
+The final piece is the proof of the Lindenbaum's theorem, which states that any consistent theory has
+a maximal consistent set. The proof of this fact is immediate from the previous lemmas.
+\begin{minted}{coq}
+Theorem lindebaum : Consistent A G -> exists D, Consistent A D /\ Maximal D /\ Subset G D.
+\end{minted}
+This concludes our formalization of the existence of the maximal consistent sets for modal logics formula.
+
+\paragraph{Canonical model construction} The second part of a Henkin style completeness proof for modal logic
+is the construction of the cannonical model. The cannonical model
+$\mathcal{M} = \langle\mathcal{F},\mathcal{V}\rangle$ is such that:
+\begin{itemize}
+  \item The frame $mathcal{F} = \langle W, \mathcal{R}\rangle$ is
+        \begin{itemize}
+          \item $\mathcal{W}$ is the set of all maximal consistent sets of formulas;
+          \item $\mathcal{R}(w_{1},w_{2})$ iff $\Box \varphi \in w_{1}$ implies $\varphi \in w_{2}$;
+        \end{itemize}
+  \item The labelling function $\mathcal{V}$ is such that $\mathcal{V}(A) \in \mathcal{W}$ and $p\in \mathcal{V}(A)$,
+        for all propositional variables $A$.
+\end{itemize}
+
+The canonical model is formalized in Coq as follows. First, we define the type \mintinline{coq}{W} that pack
+the construction of the maximal consistent sets for a theory \mintinline{coq}{G}.
+\begin{minted}{coq}
+Inductive W: Type :=
+| W_mk D: Consistent A D -> Maximal D -> Subset G D -> W.
+\end{minted}
+Next, we encode the accessibility relation, $\mathcal{R}$, by
+the definition \mintinline{coq}{canonical_accessibility} and use
+notation \mintinline{coq}{R} for it. Observe that it is just the
+translation of the mathematical notation into Gallina's code.
+\begin{minted}{coq}
+Definition canonical_accessibility: relation W :=
+  fun w v => forall p, w [! []p !] -> v p.
+
+Local Notation R := canonical_accessibility.
+\end{minted}
+Using the defined accessibility relation and the set of maximal consistent sets,
+we can build the canonical frame and define notation \mintinline{coq}{F} for it.
+\begin{minted}{coq}
+Definition canonical_frame := Build_Frame W R.
+
+Local Notation F := canonical_frame.
+\end{minted}
+The canonical labelling function is just a defined as a simple function which maps variables (represented in
+our formalization as natural numbers) to its membership in a world. From these components, the canonical model
+is built as a pair formed by the frame and the labelling function. Notation \verb|[F -- V]| denotes the
+construction of a model from the frame \verb|F| and the labelling \verb|V|.
+\begin{minted}{coq}
+Definition canonical_valuation: nat -> W -> Prop := fun p w => w [! #p !].
+
+Local Notation V := canonical_valuation.
+
+Definition canonical_model := [F -- V].
+\end{minted}
+
+\paragraph{Truth as membership in the canonical model}
+The key insight to prove the completeness result is to show that truth is membership in the canonical model.
+In other words: a formula is true at a world in the canonical model iff it is a member of that world.
+This fact is formalized by the truth theorem, which is proved by induction over formula \verb|p|:
+\begin{minted}{coq}
+Lemma truth: forall p w, (M ' w ||- p) <-> (w p).
+\end{minted}
+Notation \mintinline{coq}{M ' w ||- p} denotes the valuation of a fórmula at a given world \verb|w| of the
+canonical model. Note that a world in the canonical model is an (infinite) consistent set of formulas.
+Since we are modelling sets by their characteristic functions (functions of type \mintinline{coq}{formula -> Prop}),
+set membership is modelled by function application, \verb|w p|.
+
+\paragraph{The completeness proof} We almost have all the pieces to prove the completeness theorem. The next lemma
+shows that a formula holds in the canonical model if, and only if, it is provable in the deductive system.
+\begin{minted}{coq}
+Lemma determination : forall p, (M |= p) <-> (A; G |-- p).
+\end{minted}
+The proof is divided in two more lemmas for each implication direction using the truth lemma. Finally, the completeness theorem is stated as follows:
+\begin{minted}{coq}
+Theorem completeness : forall p, (G ||= p) -> (A; G |-- p).
+Proof.
+  intros p.
+  apply contrapositive; intros.
+  - apply classic.
+  - assert (Consistent A G).
+      + apply nonderivation_implies_consistency with p.
+        assumption.
+      + assert ((M |= p) -> False); intros.
+          * apply determination with p in H2; auto.
+          * apply H2, H0.
+            intros q ? w; apply truth.
+            destruct w as (w, ?H, ?H, ?H); simpl.
+            apply H6; assumption.
+Qed.
+\end{minted}
+The proof proceeds by the contrapositive.
+Suppose that \mintinline{coq}{A; G |-- p} does not hold.
+If the theory \verb|G| is inconsistent, the desired result follows.
+So, assume that \verb|G| is consistent. Since \verb|G| is consistent, it
+can be extended into the canonical model \verb|M|. By the fact that
+\mintinline{coq}{A; G |-- p} is false, by the determination lemma, we know
+that \verb|p| is not true on the canonical model \verb|M|. Conclusion follows
+by the truth lemma.
+
+\subsubsection{Frame Correspondence}\label{sec:correspondence}
+
 As already mentioned, there are some modal formula schemas that correspond
@@
 requires the construction of a witness valuation function.
+
+As an example, let's consider frame correspondence proofs for the reflexivity property.
+First, we consider that reflexive frames implies the axiom T. The proof script
+uses basic Coq tactics and the crucial step is the unfolding of the definition
+reflexive frame.
+\begin{minted}{coq}
+Theorem reflexive_frame_implies_axiomT :
+   forall f p, reflexivity_frame f -> forall v, [f -- v] |= [! []p -> p !].
+Proof.
+  intros f p HR v w1 H1.
+  simpl in H1.
+  unfold reflexivity_frame in HR.
+  apply H1 in HR.
+  assumption.
+Qed.
+\end{minted}
+However, the proof that axiom T implies the reflexive frame is a bit more involved. The proof
+proceeds by the contrapositive (\mintinline{coq}{apply contrapositive}). We finish the proof
+using some classical predicate logic identities involving quantifiers and negation, following
+the proofs found in classical modal logic textbooks.
+\begin{minted}{coq}
+Theorem axiomT_implies_reflexive_frame:
+   forall f, (forall v p, [f -- v] |= [! []p -> p !]) -> reflexivity_frame f.
+Proof.
+  intros f.
+  apply contrapositive.
+  - apply classic.
+  - intros H; unfold reflexivity_frame in H.
+    apply not_all_ex_not in H; destruct H as [w1].
+    apply ex_not_not_all.
+    exists (fun _ x => R f w1 x).
+    apply ex_not_not_all.
+    exists [! #0 !].
+    intros H1; unfold validate_model in H1; simpl in H1.
+    destruct H.
+    apply H1.
+    intros w2 H'; assumption.
+Qed.
+\end{minted}
 Nevertheless, all Coq proofs are fairly similar to the respective
 ``pen and paper'' proofs found in the literature. 
-Our work proves correspondence for the following properties: reflexivity,
+Our work proves correspondence for the following properties:
 transitivity, symmetry,  euclidean, seriality, functionality, density, convergency,
@@
 
-In this work, we presented a sound deep embedding of arbitrary
+In this work, we presented a sound and complete deep embedding of arbitrary
 normal modal logics in Coq. Our approach is able to represent deduction
@@
 
-As future work we intend to prove the completeness result for the Hilbert
-deduction system and continue the development of our library by including
-support for multimodalities (fusion of modal logics) and non-normal modal
-logic systems, along with improving usability through proof automation.
-
-\begin{acks}
-  The present work was supported by Fundação de Amparo à Pesquisa de Minas
-  Gerais (FAPEMIG) under grant APQ-01683-21 and Fundação de Amparo à Pesquisa de
-  Santa Catarina (FAPESC). %under grant 2021TR961.
-\end{acks}
-
-\bibliographystyle{ACM-Reference-Format}
+As future work we intend to continue the development of our library by
+including support for multimodalities (fusion of modal logics) and
+non-normal modal logic systems, along with improving usability through proof automation.
+
+%\begin{acks}
+%  The present work was supported by Fundação de Amparo à Pesquisa de Minas
+%  Gerais (FAPEMIG) under grant APQ-01683-21 and Fundação de Amparo à Pesquisa de
+%  Santa Catarina (FAPESC). %under grant 2021TR961.
+%\end{acks}
+%% appendix sections are then done as normal sections
+%% \appendix
+
+%% \section{}
+%% \label{}
+
+%% References
+%%
+%% Following citation commands can be used in the body text:
+%% Usage of \cite is as follows:
+%%   \cite{key}         ==>>  [#]
+%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
+%%
+
+%% References with BibTeX database:
+
+\bibliographystyle{elsarticle-num}
 \bibliography{references}
 
+%% Authors are advised to use a BibTeX database file for their reference list.
+%% The provided style file elsarticle-num.bst formats references in the required Procedia style
+
+%% For references without a BibTeX database:
+
+% \begin{thebibliography}{00}
+
+%% \bibitem must have the following form:
+%%   \bibitem{key}...
+%%
+
+% \bibitem{}
+
+% \end{thebibliography}
+
 \end{document}
+
+%%
+%% End of file `ecrc-template.tex'.

### 2024-01-10 paper/COLA/paper.tex
Minor fixes

@@
 \begin{itemize}
-  \item The frame $mathcal{F} = \langle W, \mathcal{R}\rangle$ is
+  \item The frame $\mathcal{F} = \langle W, \mathcal{R}\rangle$ is
         \begin{itemize}
@@
         \end{itemize}
-  \item The labelling function $\mathcal{V}$ is such that $\mathcal{V}(A) \in \mathcal{W}$ and $p\in \mathcal{V}(A)$,
-        for all propositional variables $A$.
+  \item The labelling function $\mathcal{V}$ is such that $\mathcal{V}(A) = w$, $A\in w$,
+        for some propositional variable $A$ and $w \in \mathcal{W}$.
 \end{itemize}

### 2024-01-27 paper/COLA/paper.tex
Little changes.

@@
 world iff $\varphi$ is \emph{possible}, i.e. true in at least some world
-accessible from the current world. We the remaining of the paper, we also write
+accessible from the current world. In the remaining of the paper, we also write
 $\phi_1\leftrightarrow\phi_2$ to mean $\phi_1\to\phi_2\land\phi_2\to\phi_1$.
@@
     $\mathcal{M}\,;\,w\not\models\varphi_1$ or $\mathcal{M}\,;\,w\models\varphi_2$;
-  \item $\mathcal{M}\,;\,w\models\Box\varphi$ iff $\forall y \in W. w R y \to \mathcal{M}\,;\,y \models\varphi$;
-  \item $\mathcal{M}\,;\,w\models\Diamond\varphi$ iff $\exists y \in W. w R y \land \mathcal{M}\,;\,y \models\varphi$;
+  \item $\mathcal{M}\,;\,w\models\Box\varphi$ iff $\forall y \in W. (w R y \to \mathcal{M}\,;\,y \models\varphi)$;
+  \item $\mathcal{M}\,;\,w\models\Diamond\varphi$ iff $\exists y \in W. (w R y \land \mathcal{M}\,;\,y \models\varphi)$;
   \end{enumerate}
@@
 We say that a formula $\varphi$ is satisfiable in a model $\mathcal{M}$,
-denoted by $\mathcal{M}\models \varphi$, if $\forall w \in W. M\,;w\,\models \varphi$ holds.
+denoted by $\mathcal{M}\models \varphi$, if $\forall w \in W. (M\,;w\,\models \varphi)$ holds.
 Similarly, $\varphi$ is valid in a frame $\mathcal{F}$, written $\mathcal{F}\models\varphi$, if
@@
 of formulas $\Gamma$ and a formula $\varphi$, we say that $\Gamma$ entails
-$\varphi$, written $\Gamma\models\varphi$, iff $\forall \mathcal{M} . (\forall \varphi' \in\Gamma .
-\mathcal{M} \models \varphi') \to \mathcal{M} \models \varphi$.
+$\varphi$, written $\Gamma\models\varphi$, iff $\forall \mathcal{M} . ((\forall \varphi' \in\Gamma .
+\mathcal{M} \models \varphi') \to \mathcal{M} \models \varphi)$.
 
@@
 to the set of axioms for classical propositional logic.
-As examples of modalities axioms we mention \textbf{K}, which express that $\Box$ distributes
+As examples of modalities axioms we mention K, which express that $\Box$ distributes
 over implication, and possibility which states that $\Diamond$ distributes
@@
   \begin{array}{ll}
-    \textbf{K} & \Box(\varphi_1 \to \varphi_2) \to (\Box \varphi_1 \to \Box \varphi_2)\\
-    \textbf{Possibility} & \Diamond (\varphi_1 \lor \varphi_2) \to
+    K & \Box(\varphi_1 \to \varphi_2) \to (\Box \varphi_1 \to \Box \varphi_2)\\
+    Possibility & \Diamond (\varphi_1 \lor \varphi_2) \to
                            (\Diamond \varphi_1 \lor \Diamond \varphi_2)\\
@@
 it. As an example, consider a frame $\mathcal{F} = \langle W,R
-\rangle$ in which the relation $R$ is reflexive: $\forall w \in W. wRw$, i.e.
+\rangle$ in which the relation $R$ is reflexive: $\forall w \in W. (wRw)$, i.e.
 every world is accessible from itself. In such frame when $\Box A$ holds in
@@
   Property of $R$                          & Formula schema \\ \hline
-  reflexive: $\forall w.wRw$               & $\Box A \to A$ \\ 
-  serial: $\forall w\exists w'. wRw'$      & $\Box A \to \Diamond A$ \\
-  symmetric: $\forall w\,w'.wRw' \to w'Rw$ & $A \to \Box\Diamond A$ \\
+  reflexive: $\forall w.(wRw)$               & $\Box A \to A$ \\
+  serial: $\forall w\exists w'. (wRw')$      & $\Box A \to \Diamond A$ \\
+  symmetric: $\forall w\,w'.(wRw' \to w'Rw)$ & $A \to \Box\Diamond A$ \\
   \hline
@@
 Basically, a Hilbert-style deductive system allows the use of four
-possibilities in a proof: 1) the use of a hypothesis; 2) the use of an axiom;
+possibilities in a proof: 1) the use of a hypothesis; 2) the use of an instantiation of an axiom;
 3) the use of the modus ponens rule or 4) the use of the generalization (necessity)
@@
 \begin{minted}{coq}
-Inductive deduction (A: axiom -> Prop):
-                          theory -> formula -> Prop :=
+Inductive deduction (A: axiom -> Prop) : theory -> formula -> Prop :=
   (* Premise. *)
@@
 which denote each system. The simplest set of axioms is usually referred as
-\textbf{K} and it includes the axiom $\Box(A \to B) \to (\Box A \to \Box B)$
-and all the axioms for propositional logic. We represent system \textbf{K}
+K and it includes the axiom $\Box(A \to B) \to (\Box A \to \Box B)$
+and all the axioms for propositional logic. We represent system K
 by the following inductive predicate:
@@
 of \mintinline{coq}{axiom}, a value of type \mintinline{coq}{K a} denotes a
-proof that a given axiom is part of the system \textbf{K}. This is the main
+proof that a given axiom is part of the system K. This is the main
 trick used by the constructor \mintinline{coq}{Ax}, of type
@@
 The proof proceeds by induction on the derivation of $\Gamma \vdash\varphi$ and
-it has a case for each of the 11 axioms of the Hilbert system for propositional
-logics, one for the $K$ axiom and cases for using hypothesis and the rules modus
-ponens. In order to modularize the proof, we defined a lemma for each case.
+it has a case for each of the axioms of the Hilbert system for propositional
+logics, one for the $K$ axiom and cases for using hypothesis, the rules modus
+ponens and necessitation. In order to modularize the proof, we defined a lemma for each case.
 As example, the following is the statement for modus ponens case.
@@
 In this section, we describe our formalization of the completeness result for the
-Hilbert proof system with respect to the Kripke semantics. The proof consists of two
+Hilbert proof system with respect to the Kripke semantics.
+It is worth mentioning that there are different notions of completeness and
+the literature is not always clear about the difference of those.
+
+{\color{red} TORRENS: Escrever definições de completude}
+
+The proof consists of two
 basic steps: 1) the proof of the existence of maximal consistent sets and; 2) the
-construction of a cannonical model from the maximal consistent set. We describe these
-steps separately.
-
-\paragraph{Maximal consistent sets} The first step on the completeness proof is the construction of
-a \emph{maximal consistent sets}, $\Delta$, of formulas for $\Gamma$ and $\varphi$.  The $\Delta$ set
+construction of a cannonical model from the maximal consistent set. We describe these steps separately.
+
+\paragraph{Maximal consistent sets} The first step on the completeness proof is the construction of a \emph{maximal consistent set}, $\Delta$, of formulas for $\Gamma$ and $\varphi$.  The $\Delta$ set
 is defined as follows:
@@
 
-Which states that it is not possible to deduce \verb|p| and \verb|~ p|, for any formula \verb|p|.
+Which states that it is not possible to deduce \mintinline{coq}{p} and \mintinline{coq}{~p}, for any formula \mintinline{coq}{p}.
 In order to define maximal consistent sets, we use the following representation of a theory
@@
 this fact by two functions \mintinline{coq}{encode :: formula -> nat} and \mintinline{coq}{decode :: nat -> option formula}
-which satisfies the following property: $\forall \varphi . \texttt{decode(encode }\varphi\texttt{)} = \varphi$.
+which satisfies the following property: $\forall \varphi . (\texttt{decode(encode }\varphi\texttt{)} = \varphi)$.
 
@@
 the situation when the inclusion of a formula \verb|p| makes the theory inconsistent.
-In this case, we add the negation of \verb|p| to the theory.
+In this case, we add the negation of \verb|p| to the theory. Notation \mintinline{coq}{Extend p D} denotes the union of theory \mintinline{coq}{D} with the
+singleton set containing \mintinline{coq}{p}.
 
@@
 \end{minted}
-Next, we show that the set \mintinline{coq}{Max} is such that either \verb|p| or \verb|~ p| is an element of \verb|Max|.
+Next, we show that the set \mintinline{coq}{Max} is such that either \verb|p| or \verb|~p| is an element of \verb|Max|.
 \begin{minted}{coq}
@@
 \begin{minted}{coq}
-Theorem lindebaum : Consistent A G -> exists D, Consistent A D /\ Maximal D /\ Subset G D.
-\end{minted}
-This concludes our formalization of the existence of the maximal consistent sets for modal logics formula.
+Theorem lindenbaum : Consistent A G -> exists D, Consistent A D /\ Maximal D /\ Subset G D.
+\end{minted}
+This concludes our formalization of the existence of the maximal consistent set for modal logic theories.
 
@@
 fixed points. Formally, let $\varphi(X)$ be a formula with only one propositional variable
-$X$. The modal fixed point of $\varphi$ is then a formula $\Psi$ such that such that
+$X$. The modal fixed point of $\varphi$ is then a formula $\Psi$ such that
 $\Psi \leftrightarrow \varphi(\Box \Psi)$.
@@
 which is parameterized by a system of modal logic, \mintinline{coq}{S}, and a
-theory, \mintinline{coq}{g}, and encodes the logical equivalence demonstrating
+theory, \mintinline{coq}{G}, and encodes the logical equivalence demonstrating
 that \mintinline{coq}{p} is a modal fixed point for \mintinline{coq}{f}.
 \begin{minted}{coq}
-  Definition fixed_point S g: Prop :=
+  Definition fixed_point A G: Prop :=
     forall f, exists p,
-    (S; g |-- [! (p <-> f ([]p)) !]).
+    (A; G |-- [! (p <-> f ([]p)) !]).
 \end{minted}
@@
   Theorem Lob:
-    forall A, subset K4 A /\ fixed_point A nil ->
-    forall P, (A; nil |-- [! []P -> P !]) ->
-              (A; nil |-- [! P !]).
-\end{minted}
-
-\noindent The proof starts by assuming an arbitrary formula $P$, the existence of modal fixed
-points and that $\Box P \to P$ is provable with axioms A. These deduction steps
+    forall A G, subset K4 A /\ fixed_point A G ->
+    forall p, (A; G |-- [! []p -> p !]) ->
+              (A; G |-- [! p !]).
+\end{minted}
+
+\noindent The proof starts by assuming an arbitrary formula $p$, the existence of modal fixed
+points and that $\Box p \to p$ is provable with axioms A. These deduction steps
 are performed by the following tactic:
@@
 \begin{minted}{coq}
-  intros A [I FP] P H1.
+  intros A G [I FP] P H1.
 \end{minted}
@@
 we know that there is a formula $\Psi$ such that $\Psi \leftrightarrow (\Box
-\Psi \to P)$. The following tactic performs this step of reasoning.
+\Psi \to p)$. The following tactic performs this step of reasoning.
 
@@
 
-\noindent Now, we are able to deduce that $\Psi \to (\Box \Psi\to P)$ from the previous
+\noindent Now, we are able to deduce that $\Psi \to (\Box \Psi\to p)$ from the previous
 step result as follows:
@@
 \begin{minted}{coq}
-  assert (A; nil |-- [! psi -> []psi -> P !]) as H3.
+  assert (A; G |-- [! psi -> []psi -> P !]) as H3.
   apply modal_ax5 in H2; auto with modal.
@@
 auxiliary proof goals. In the previous code piece, we specified
-$\vdash \Psi \to (\Box \Psi \to P)$ as a new goal which
+$\vdash \Psi \to (\Box \Psi \to p)$ as a new goal which
 generates a new proof hypothesis named \mintinline{coq}{H3}. The
@@
 
-From $\vdash \Psi \to (\Box \Psi \to P)$ and the necessitation
-rule, we can deduce $\vdash \Box (\Psi \to (\Box \Psi \to P))$
+From $\vdash \Psi \to (\Box \Psi \to p)$ and the necessitation
+rule, we can deduce $\vdash \Box (\Psi \to (\Box \Psi \to p))$
 using the following tactics:
 \begin{minted}{coq}
-  assert (A; nil |-- [! [](psi -> []psi -> P) !])
+  assert (A; G |-- [! [](psi -> []psi -> p) !])
     as H4.
@@
 
-\noindent Now, using the axiom \textbf{K} we can deduce $\vdash \Box \Psi \to
-\Box (\Box \Psi \to P)$. Again, we deduce this fact using the
+\noindent Now, using the axiom K we can deduce $\vdash \Box \Psi \to
+\Box (\Box \Psi \to p)$. Again, we deduce this fact using the
 \mintinline{coq}{assert} tactic.
@@
 \begin{minted}{coq}
-  assert (A; nil |-- [! []psi -> []([]psi -> P) !])
+  assert (A; G |-- [! []psi -> []([]psi -> p) !])
     as H5.
@@
 \noindent The next proof step involves showing that
-$\vdash \Box (\Box \Psi \to P)\to (\Box \Box \Psi \to \Box P)$ holds
-by axiom \textbf{K}.
-
-\begin{minted}{coq}
-  assert (A; nil |-- [! []([]psi -> P) ->
-                          [][]psi -> []P !]) as H6.
+$\vdash \Box (\Box \Psi \to p)\to (\Box \Box \Psi \to \Box p)$ holds
+by axiom K.
+
+\begin{minted}{coq}
+  assert (A; G |-- [! []([]psi -> p) ->
+                          [][]psi -> []p !]) as H6.
   eapply Ax with (a := axK ?[X] ?[Y]); auto with modal.
@@
 
-\noindent Using $\vdash \Box (\Box \Psi \to P)\to (\Box \Box \Psi \to \Box P)$ and
-$\vdash \Box \Psi \to\Box (\Box \Psi \to P)$ we can deduce
+\noindent Using $\vdash \Box (\Box \Psi \to p)\to (\Box \Box \Psi \to \Box p)$ and
+$\vdash \Box \Psi \to\Box (\Box \Psi \to p)$ we can deduce
 $\vdash \Box \Psi \to \Box \Box \Psi \to \Box P$ using transitivity of
@@
 \begin{minted}{coq}
-  assert (A; nil |-- [! []psi -> [][]psi -> []P !])
+  assert (A; G |-- [! []psi -> [][]psi -> []p !])
     as H7.
@@
 \begin{minted}{coq}
-  assert (A; nil |-- [! []psi -> [][]psi !]) as H8.
+  assert (A; G |-- [! []psi -> [][]psi !]) as H8.
   apply modal_axK4; auto with modal.
@@
 \noindent From the previou two hypotheses we gather necessary info to derive
-$\vdash \Box \Psi \to P$, using a propositional axiom.
-
-\begin{minted}{coq}
-  assert (A; nil |-- [! []psi -> []P !]) as H9.
+$\vdash \Box \Psi \to p$, using a propositional axiom.
+
+\begin{minted}{coq}
+  assert (A; G |-- [! []psi -> []p !]) as H9.
   eapply modal_ax2; eauto with modal.
@@
 
-\noindent Then, from $\vdash \Box\Psi \to \Box P$ and $\vdash \Box P \to P$, we can
-conclude $\vdash \Box \Psi \to P$, by composition of implication as
+\noindent Then, from $\vdash \Box\Psi \to \Box p$ and $\vdash \Box p \to p$, we can
+conclude $\vdash \Box \Psi \to p$, by composition of implication as
 follows.
@@
 \begin{minted}{coq}
-  assert (A; nil |-- [! []psi -> P !]) as H10.
+  assert (A; G |-- [! []psi -> p !]) as H10.
   eapply modal_compose; eauto with modal.
@@
 
-\noindent Next, we can conclude $\vdash \Box (\Psi \to P) \to \Psi$ from
-$\Psi \leftrightarrow (\Box \Psi \to P)$ using
+\noindent Next, we can conclude $\vdash \Box (\Psi \to p) \to \Psi$ from
+$\Psi \leftrightarrow (\Box \Psi \to p)$ using
 the following tactics:
@@
 \begin{minted}{coq}
-  assert (A; nil |-- [! ([]psi -> P) -> psi !]) as H11.
+  assert (A; G |-- [! ([]psi -> p) -> psi !]) as H11.
   apply modal_ax6 in H2; auto with modal.
@@
 
-\noindent Now, from $\vdash (\Box\Psi \to P) \to \Psi$ and $\vdash \Box \Psi \to P$
+\noindent Now, from $\vdash (\Box\Psi \to p) \to \Psi$ and $\vdash \Box \Psi \to p$
 we can conclude $\vdash \Psi$ by modus ponens.
@@
 \begin{minted}{coq}
-  assert (A; nil |-- psi) as H12.
+  assert (A; G |-- psi) as H12.
   eapply Mp; try eassumption.
@@
 \begin{minted}{coq}
-  assert (A; nil |-- [! []psi !]) as H13.
+  assert (A; G |-- [! []psi !]) as H13.
   apply Nec; try eassumption.
@@
 as a predicate parameter in the deduction inductive type. Using this generic
-representation of deduction we are able to prove a soundness theorem, and, in
-order to demonstrate the applicability of the proposed work, prove L\"ob's theorem
+representation of deduction we are able to prove soundness and completeness theorems, and, in order to demonstrate the applicability of the proposed work, prove L\"ob's theorem
 by assuming fixed points. % All source code in this article has
@@
 non-normal modal logic systems, along with improving usability through proof automation.
+{\color{red} TORRENS: Dedução natural e táticas.}
 

### 2024-01-30 paper/COLA/paper.tex
Some fixes and missing pieces

@@
 \ead{paulotorrens@gnu.org}
-\address{Santa Catarina State University}
+\address{University of Kent}
 
@@
 \begin{abstract}
-  This work describes an encoding of modal logics using the Coq proof assistant.
+  This work describes a deep-embedding formalization of modal logic systems using the Coq proof assistant.
   Our formalization differs from previous attempts by providing a uniform
   representation of several systems for modal logic using Coq's rich type
-  structure. We illustrate the usefulness of our library in a formalization of
+  structure by quantifying over the set of axioms, for which we prove soundness and completeness. Finally, we motivate the use of our library in a formalization of
   L\"ob's theorem which closely follows a classical proof of this result.
@@
 
-The start of modal logics can be traced back to Lewis and his definition of 
+The origins of modal logic can be traced back to Lewis and his definition of
 logical systems with strict implication instead of material implication \cite{Lewis1918}.
-Based in this work, independently of classical propositional logic,
+Based on this work, and independently of classical propositional logic, he
 defined the systems $\mathsf{S1}$--$\mathsf{S5}$~\cite{Lewis1932}, where $\mathsf{S4}$ and $\mathsf{S5}$
-are well known normal modal systems. It was, in fact, G{\"o}del that had the idea 
+are now well known normal modal systems. It was, in fact, G{\"o}del that had the idea
 to build a modal logic system as an extension of the classical propositional calculus,
@@
 Despite that the original motivation of modal systems was to deal with strict implication,
-for many years modal logic was seen as ``the logic of necessity and possibility''. 
+for many years modal logic has been seen as ``the logic of necessity and possibility''.
 Only after the introduction of Kripke semantics they were also seen as ``simple yet
@@
 formalisms for talking about time, space, knowledge, beliefs, actions,
-obligations, among others.
-
-The interpretation of modal logic simply as ``the logic of necessity'' restricts 
-not only its applications but also the range of modal systems. In fact, modal logic 
+obligations, among others, by axiomatizing different structures.
+
+We note that the interpretation of modal logic simply as ``the logic of necessity'' restricts
+not only its applications but also the possible range of modal systems. In fact, modal logic
 can also be seen as ``the study of \emph{modalities} -- logical operations that qualify
 assertions about the truth of statements''~\cite{Goldblatt-MathofModality}. In this sense,
-the necessitation rule (from $\varphi$ infer $\Box\varphi$) may be relaxed. This brings 
+the standard necessitation rule (i.e., that from $\varphi$ we may infer $\Box\varphi$) may be actually relaxed. This brings
 a number of modal systems known as non-normal, whose semantics were defined only in the
@@
 Such wide range of highly expressive logical languages is the ideal tool to
-formalize the many complexities of modern advanced software systems, 
+formalize the many non-trivial properties of modern advanced software systems,
 such as those found in cars, airplanes, medical devices and banking applications.
-In combining said modal logics with interactive proof assistants,
+By combining said modal logics with interactive proof assistants,
 such as Coq~\cite{manual_coq} and Isabelle/HOL~\cite{manual_isabelle}, both of
 which have been applied with success in non-trivial projects like a certified compiler
-for C programming language~\cite{Leroy09} and the kernel of seL4 operating
-system~\cite{Klein10}, it is possible to formalize and verify highly complex computer
-systems.
+for C programming language~\cite{Leroy09} and the kernel of an operating
+system~\cite{Klein10},
+% it is possible to formalize and verify highly complex computer systems.
+we argue we may achieve betters ways of formally describing and verifying complex programs.
 
 Although there are some dedicated provers for modal logics~\cite{Gleissner17,Mora11},
-we believe that using mature tools, like a proof assistant, is a better option for
+we believe that using mature tools, like a proof assistant, would be a better option for
 verifying facts using modal logics. We justify this choice by: 1) the recent
 successful histories on verification of software and
-mathematics~\cite{Leroy09,Klein10,Gonthier13,Gonthier08}; 2) there is a growing
+mathematics~\cite{Leroy09,Klein10,Gonthier13,Gonthier08}; 2) the growing
 community of developers using such tools thanks to
@@
     define its Kripke semantics.
-  \item We model the Hilbert deductive system as a Coq inductive type that is
-    parameterized by different axiom systems like K, D, B, T, K4, K5, S4, S5 and GL.
+  \item We model the Hilbert-style deductive system as a Coq inductive type that is
+    parameterized by the set of axioms, allowing to derive systems like K, D, B, T, K4, K5, S4, S5 and GL.
   \item We formally verify several frame correspondence results.
-  \item We prove soundness and completeness for the Hilbert deductive system. For the
-        completeness proof we follow the so-called Henkin-style and it closely follows
-        the proofs presented in classical modal logics textbooks~\cite{Blackburn-ModalLogics}.
-  \item As an example of usage, we show a proof of L\"ob's theorem using the deductive system
-    developed in our library.
+  \item We prove both soundness and completeness for the deductive system with regard to its embedding within Coq's own logic (the calculus of inductive constructions).
+
+  % For the
+  %       completeness proof we follow the so-called Henkin-style and it closely follows
+  %      the proofs presented in classical modal logics textbooks~\cite{Blackburn-ModalLogics}.
+  \item As an example of usage, and in order to motivate the use of our library, we show a proof of L\"ob's theorem using the deductive system
+    developed.
 \end{itemize}
 
-This article is an extended version of our SBLP 2022 work~\cite{Silveira22}. In this paper we include
-a detailed discussion about our formalization of a completeness proof for the Hilbert deductive
+This article is an extended version of our SBLP 2022 work~\cite{Silveira22}. In this paper we also include
+a detailed discussion about our formalization of a completeness proof for the Hilbert-style deductive
 system following the strategies found in traditional modal logic books~\cite{Blackburn-ModalLogics}.
@@
 provide some pointers for future work.
-
+%
 All source code in this article has been formalized in Coq version 8.15 and
-is available at \url{https://github.com/funcao/ModalLibrary/} under a permissive free software license.
+is available at \url{https://github.com/funcao/LML/} under a permissive free software license.
 
@@
 
-\subsection{An overview of Coq proof assistant.}
+\subsection{A brief overview of Coq proof assistant.}
 
 Coq is a proof assistant based on the calculus of inductive
-constructions (CIC)~\cite{manual_cic}, a higher-order typed
+constructions ~\cite{manual_cic}, a higher-order dependently typed
 $\lambda$-calculus extended with inductive definitions.  Theorem
@@
   scope of local variables.} which declares variables \mintinline{coq}{A},
-\mintinline{coq}{B} and \mintinline{coq}{C} as being propositions (i.e. with type
+\mintinline{coq}{B} and \mintinline{coq}{C} as being propositions (i.e., with type
 \mintinline{coq}{Prop}). Tactic \mintinline{coq}{intros} introduces variables
@@
 \]
-Intuitively, the semantics of the operator $\Box$ is similar to an universal
+Intuitively, the semantics of the operator $\Box$ is similar to a universal
 quantifier, since it expresses \emph{necessity}: $\Box\,\varphi$ is true in the
@@
 $\Diamond$ is that a formula $\Diamond\varphi$ is true in the current
-world iff $\varphi$ is \emph{possible}, i.e. true in at least some world
+world iff $\varphi$ is \emph{possible}, i.e., true in at least some world
 accessible from the current world. In the remaining of the paper, we also write
@@
 defined as follows.
-A \emph{frame} $\mathcal{F} = \langle W, R \rangle$ consists of a pair formed by
+A \emph{frame} $\mathcal{F} = \langle W, R \rangle$ \linebreak consists of a pair formed by
 a non-empty set of worlds, $W$, and a binary accessibility relation between
-worlds, $R \subseteq W \times W$. A pair $w_1Rw_2$ denotes that world $w_2$ is
+worlds,
+\linebreak
+$R \subseteq W \times W$. A pair $w_1Rw_2$ denotes that world $w_2$ is
 accessible from $w_1$. A model $\mathcal{M} = \langle \mathcal{F}, \mathcal{V} \rangle$
 is formed by a frame $\mathcal{F}$ and a total labeling function
-$\mathcal{V} : \mathcal{P}\to 2^W$, which indicates the set of worlds where each propositional variable holds, i.e. when $w \in \mathcal{V}(A)$
+$\mathcal{V} : \mathcal{P}\to 2^W$, which indicates the set of worlds where each propositional variable holds, i.e., when $w \in \mathcal{V}(A)$
 for some $A \in\mathcal{P}$, we say that $A$ holds in $w$. The satisfiability of a formula
@@
   \item $\mathcal{M}\,;\,w\models\Box\varphi$ iff $\forall y \in W. (w R y \to \mathcal{M}\,;\,y \models\varphi)$;
-  \item $\mathcal{M}\,;\,w\models\Diamond\varphi$ iff $\exists y \in W. (w R y \land \mathcal{M}\,;\,y \models\varphi)$;
+  \item $\mathcal{M}\,;\,w\models\Diamond\varphi$ iff $\exists y \in W. (w R y \land \mathcal{M}\,;\,y \models\varphi)$.
   \end{enumerate}
@@
 Similarly, $\varphi$ is valid in a frame $\mathcal{F}$, written $\mathcal{F}\models\varphi$, if
-it holds on all models $\mathcal{M} = \langle \mathcal{F},\mathcal{V}\rangle$, i.e.
+it holds on all models $\mathcal{M} = \langle \mathcal{F},\mathcal{V}\rangle$, i.e.,
 for any $\mathcal{M}$, we have that $\mathcal{M}\models\varphi$. A formula
-is valid, written $\models \varphi$, if it is valid in any frame $\mathcal{F}$.
-
-Using the semantics, we can define the concept of \emph{entailment}: given a set
+is valid, written $\models \varphi$, if it is valid in any frame $\mathcal{F}$. Using the above semantics, we can define the concept of \emph{entailment}: given a set
 of formulas $\Gamma$ and a formula $\varphi$, we say that $\Gamma$ entails
@@
 \mathcal{M} \models \varphi') \to \mathcal{M} \models \varphi)$.
+In this definition, if $\Gamma$ entails $\varphi$ it means that $\varphi$ is a global semantic consequence of $\Gamma$, which will be further elaborated on Section~\ref{sec:completeness}.
 
 In this work, we consider Hilbert deductive systems for modal logic.
-Essentially, the considered system adds axioms for the modalities
-to the set of axioms for classical propositional logic.
-As examples of modalities axioms we mention K, which express that $\Box$ distributes
-over implication, and possibility which states that $\Diamond$ distributes
+Essentially, the considered system add axioms for the modalities
+to the axioms for classical propositional logic.
+As examples of modalities axioms we mention $K$, which express that $\Box$ distributes
+over implication, and possibility, which states that $\Diamond$ distributes
 over disjunction.
@@
 
-A central aspect in the study of modal logics is the so-called \emph{frame
+%% -----------------------------------------------------------------------------
+
+It's worth to mention about the choice of Hilbert-style systems as the deductive system in our library.
+The main reason is itself the soundness and completeness proofs: deductive systems such as natural deduction~\cite{Russo_PMLD, basin.ea:labelled:1997, SVRS02-recipe}, sequent calculi~\cite{Governatori00labelledmodal, 03-MSSV-fiblog19}, or tableaux~\cite{BrodaRusso05}, in modal logics, usually need the aid of labels or provisos.
+This makes, for instance, the metatheorem of soundness harder to proof: while Hilbert systems need only a simple structural induction on the size of the proof, labeled systems need to deal with several corner cases.
+The disadvantage of the Hilbert system is to use it as a way to establish proofs \emph{in} the system itself, which is not ``direct'' or ``natural''.
+However, it is still possible to use the $\mathcal{L}$tac language in Coq to automatize parts of this work.
+
+%% -----------------------------------------------------------------------------
+
+Finally, a central aspect in the study of modal logics is the so-called \emph{frame
   correspondence}, which is a relationship between the
@@
 it. As an example, consider a frame $\mathcal{F} = \langle W,R
-\rangle$ in which the relation $R$ is reflexive: $\forall w \in W. (wRw)$, i.e.
+\rangle$ in which the relation $R$ is reflexive: $\forall w \in W. (wRw)$, i.e.,
 every world is accessible from itself. In such frame when $\Box A$ holds in
@@
 generalize results about derivations to several deductive systems, independently
-of the axioms used. We conclude by discussing the soundness and the frame
+of the axioms used. We conclude by discussing the soundness, completeness, and the frame
 correspondence proofs.
@@
 To implement our library, we use a \emph{deep embedding} of
-formulas, i.e. we represent formulas as the following Coq inductive datatype.
+formulas, i.e., we represent formulas as the following Coq inductive datatype.
 \begin{minted}{coq}
@@
 $\Diamond$. Furthermore, the type \mintinline{coq}{theory} is defined as a function
-between \mintinline{coq}{formula} and Coq's \mintinline{coq}{Prop}s.
+between \mintinline{coq}{formula} and Coq's \mintinline{coq}{Prop}s, which is used to represent a set of formulas.
 
@@
 interpret a modal logic formula. We encode a frame by a record formed by
-a type \mintinline{coq}{W: Set} to represent the set of worlds and an
+a type \mintinline{coq}{W: Type} to represent the set of worlds and an
 accessibility relation denoted by a type \mintinline{coq}{R: W -> W -> Prop}.
@@
 Record Frame: Type := {
-  W: Set;
-  R: W -> W -> Prop;
+  W: Type;
+  R: W -> W -> Prop
 }.
@@
   F: Frame;
-  v: nat -> (W F) -> Prop; 
+  v: nat -> (W F) -> Prop
 }.
@@
 accessing field \mintinline{coq}{W} from the record \mintinline{coq}{F}.}. The encoding of the Kripke semantics is done by a Coq function
-that implements the relation of Figure~\ref{fig:kripke}. We omit the code
+that implements the relation of Figure~\ref{fig:kripke}; we omit the code
 for propositional logics connectives for brevity. The interesting parts are for
@@
 \begin{minted}{coq}
-Fixpoint formula_valuation (M: Model) (w: W (F M))
-                           (phi: formula): Prop :=
+Fixpoint formula_valuation (M: Model) (w: W (F M)) (phi: formula): Prop :=
   match phi with
@@
   | Box psi
-     => forall w': W (F M), R (F M) w w' ->
-                            formula_valuation M w' psi
+     => forall w': W (F M), R (F M) w w' -> formula_valuation M w' psi
   | Dia psi
-     => exists w': W (F M), R (F M) w w' /\
-                            formula_valuation M w' psi
+     => exists2 w': W (F M), R (F M) w w' & formula_valuation M w' psi
   (* some code omitted *)
@@
 In order to ease the task of writing formulas and its semantics properties, we
-define Coq notations that mimic the standard way of writing them.
+define Coq notations that mimic the standard way of writing them
+by using Coq's custom notation features.
 
@@
 \begin{minted}{coq}
-Inductive deduction (A: axiom -> Prop) : theory -> formula -> Prop :=
+Inductive deduction (A: axiom -> Prop): theory -> formula -> Prop :=
   (* Premise. *)
@@
   | axD  : formula -> axiom
-  | axK5 : formula -> axiom.
+  | axK5 : formula -> axiom
+  | axGL : formula -> axiom.
 \end{minted}
@@
 its type stores the subformulas which compose the axiom. The type has
-17 constructors: the first 10 encode axioms for propositional logic
+18 constructors: the first 10 encode axioms for propositional logic
 and, the last 7 are for different formulas which are taken as axioms by common
@@
      => [! (p -> (q -> r)) -> ((p -> q) -> (p -> r)) !]
-  | ax3    p   q
-     => [! (~ q -> ~ p) -> (p -> q) !]
+  (* some code omitted *)
   | axK p q
@@
 Notation \mintinline{coq}{[! p !]} is used to write modal logic formulas using a
-syntax closer to mathematical practice using Coq. This is necessary to avoid
+syntax closer to mathematical practice using Coq. This is useful to use a short notation while still avoiding
 parsing ambiguities, since some symbols are also used by Coq's syntax.
@@
 and \mintinline{coq}{<>}, respectively.
-
-The \mintinline{coq}{Ax} constructor uses the \mintinline{coq}{instantiate}
+%
+With such definitions at hand,
+the \mintinline{coq}{Ax} constructor uses the \mintinline{coq}{instantiate}
 function to create a formula instance from a given system,
@@
 axiom \mintinline{coq}{a} is valid within a system \mintinline{coq}{A} whenever
-the type \mintinline{coq}{A a} is inhabited, i.e. it is not empty. Next, we
+the type \mintinline{coq}{A a} is inhabited, i.e., it is not empty. Next, we
+proceed to
 describe how to represent different systems of modal logic.
@@
 which denote each system. The simplest set of axioms is usually referred as
-K and it includes the axiom $\Box(A \to B) \to (\Box A \to \Box B)$
+K and it includes the axioms $K$, possibility,
 and all the axioms for propositional logic. We represent system K
@@
   | K_axK: forall p q, K (axK p q)
+  | K_axPos: forall p q, K (axPos p q).
 \end{minted}
@@
 in our library is its soundness (through system K). If a formula $\varphi$ is derived from a
-context $\Gamma$, i.e. $\Gamma\vdash\varphi$, then $\varphi$ is entailed within the deep embedding
-from the set of hypotheses $\Gamma$, i.e. $\Gamma\models\varphi$. The Coq statement for the soundness
+context $\Gamma$, i.e., $\Gamma\vdash\varphi$, then $\varphi$ is entailed within the deep embedding
+from the set of hypotheses $\Gamma$, i.e., $\Gamma\models\varphi$. The Coq statement for the soundness
 result is given as follow:
@@
 \begin{minted}{coq}
-Theorem soundness : forall G p, (K; G |-- p) -> (G ||= p).
+Theorem soundness: forall G p, (K; G |-- p) -> (G ||= p).
 \end{minted}
@@
 it has a case for each of the axioms of the Hilbert system for propositional
-logics, one for the $K$ axiom and cases for using hypothesis, the rules modus
+logics, one for each axioms $K$ and possibility, and the rules for modus
 ponens and necessitation. In order to modularize the proof, we defined a lemma for each case.
@@
 semantics and entailment. We note that system K is used as it doesn't impose
-any restrictions on the frame.
+any restrictions on the frame:
+i.e., the relation on the frame may be an arbitrary relation.
 
@@
 Hilbert proof system with respect to the Kripke semantics.
-It is worth mentioning that there are different notions of completeness and
-the literature is not always clear about the difference of those.
-
-{\color{red} TORRENS: Escrever definições de completude}
+Before we start,
+it is worth mentioning that there are different notions of completeness, and
+the literature is not always clear about the difference of those~\cite{Popkorn_1994,modal_lean}.
+%-------------------------------------------------------------------------------
+The semantic entailment, $\Gamma\models\varphi$, represents a global consequence since, by our definition, we expect that the hypotheses are valid in every world in a model instead of only on those worlds where $\varphi$ hold \cite{de1999global}.
+Also, the completeness result may be weak, which is only valid without hypotheses, or strong, allowing it to be stated with any set of hypotheses \cite{blackburn}.
+We follow the definitions as stabilished by Popkorn \cite{Popkorn_1994} and note that we prove a strong completeness result using a global consequence: as this subsumes both weak completeness and a local one (as local consequence implies global consequence \cite{de1999global}), our result allows us to derive other possible definitions of completeness, while still being a weak result enough that it will work for any of the possible sets of axioms \cite{Popkorn_1994}.
+
+%-------------------------------------------------------------------------------
 
@@
   \begin{array}{lcl}
-    \Delta_{0}    & = & \Gamma \cup \{\neg \varphi\}\\
+    \Delta_{0}    & = & \Gamma \cup \{\neg \varphi\}\\ \\
     \Delta_{n + 1} & = & \left \{
@@
 The completeness proof rely on the fact that the set of modal logic formulas is countable. We model
-this fact by two functions \mintinline{coq}{encode :: formula -> nat} and \mintinline{coq}{decode :: nat -> option formula}
-which satisfies the following property: $\forall \varphi . (\texttt{decode(encode }\varphi\texttt{)} = \varphi)$.
+this fact by two functions \mintinline{coq}{encode: formula -> nat} and \mintinline{coq}{decode: nat -> option formula}
+which satisfies the following property: $\forall \varphi . (\texttt{decode(encode }\varphi\texttt{)} = \varphi)$, stated in the \mintinline{coq}{countable} axiom.
 
@@
 Inductive Insert: option formula -> theory -> theory :=
-| Insert_skip: forall D, Subset D (Insert None D)
+| Insert_skip:
+  forall D,
+  Subset D (Insert None D)
 | Insert_valid:
-  forall D p, Consistent A (Extend p D) -> Subset (Extend p D) (Insert (Some p) D)
+  forall D p,
+  Consistent A (Extend p D) -> Subset (Extend p D) (Insert (Some p) D)
 | Insert_invalid:
-  forall D p, ~ Consistent A (Extend p D) -> Subset (Extend [! ~p !] D) (Insert (Some p) D).
+  forall D p,
+  ~Consistent A (Extend p D) -> Subset (Extend [! ~p !] D) (Insert (Some p) D).
 \end{minted}
@@
 \begin{minted}{coq}
-Definition Max : theory := UnionOf Delta.
-Definition UnionOf (A : nat -> theory) : theory := fun t => exists i, A i t.
+Definition Max: theory := UnionOf Delta.
+Definition UnionOf (A: nat -> theory): theory := fun t => exists i, A i t.
 \end{minted}
@@
 \end{minted}
-and using the previous lemma, we prove that operation \mintinline{coq}{Delta n} preserves consistency, by induction on \verb|n|.
+And, using the previous lemma, we prove that operation \mintinline{coq}{Delta n} preserves consistency, by induction on \verb|n|.
 \begin{minted}{coq}
@@
 \begin{minted}{coq}
-Theorem lindenbaum : Consistent A G -> exists D, Consistent A D /\ Maximal D /\ Subset G D.
+Theorem lindenbaum: Consistent A G -> exists D, Consistent A D /\ Maximal D /\ Subset G D.
 \end{minted}
@@
 is the construction of the cannonical model. The cannonical model
-$\mathcal{M} = \langle\mathcal{F},\mathcal{V}\rangle$ is such that:
+$\mathcal{M}_\Gamma = \langle\mathcal{F}_\Gamma,\mathcal{V}\rangle$ is such that:
 \begin{itemize}
-  \item The frame $\mathcal{F} = \langle W, \mathcal{R}\rangle$ is
+  \item The frame $\mathcal{F}_\Gamma = \langle W, \mathcal{R}\rangle$ is
         \begin{itemize}
-          \item $\mathcal{W}$ is the set of all maximal consistent sets of formulas;
+          \item $\mathcal{W}$ is the set of all maximal consistent sets of formulas extending some context $\Gamma$;
           \item $\mathcal{R}(w_{1},w_{2})$ iff $\Box \varphi \in w_{1}$ implies $\varphi \in w_{2}$;
         \end{itemize}
-  \item The labelling function $\mathcal{V}$ is such that $\mathcal{V}(A) = w$, $A\in w$,
+  \item The labelling function $\mathcal{V}$ is such that $\mathcal{V}(A) = \{\ w\ |\ A\in w\ \}$,
         for some propositional variable $A$ and $w \in \mathcal{W}$.
@@
 
-The canonical model is formalized in Coq as follows. First, we define the type \mintinline{coq}{W} that pack
+In the folllowing, we omit the $\Gamma$ subscripts. The canonical model is formalized in Coq as follows: first, we define the type \mintinline{coq}{W} that pack
 the construction of the maximal consistent sets for a theory \mintinline{coq}{G}.
@@
 Definition canonical_model := [F -- V].
+
+Local Notation M := canonical_model.
 \end{minted}
@@
 \begin{minted}{coq}
-Lemma determination : forall p, (M |= p) <-> (A; G |-- p).
-\end{minted}
-The proof is divided in two more lemmas for each implication direction using the truth lemma. Finally, the completeness theorem is stated as follows:
-\begin{minted}{coq}
-Theorem completeness : forall p, (G ||= p) -> (A; G |-- p).
+Lemma determination: forall p, (M |= p) <-> (A; G |-- p).
+\end{minted}
+The proof is divided in two more lemmas for each implication direction using the truth lemma. Finally, the desired completeness theorem is stated simply as follows:
+\begin{minted}{coq}
+Theorem completeness: forall p, (G ||= p) -> (A; G |-- p).
 Proof.
@@
 Suppose that \mintinline{coq}{A; G |-- p} does not hold.
-If the theory \verb|G| is inconsistent, the desired result follows.
-So, assume that \verb|G| is consistent. Since \verb|G| is consistent, it
+As this system can't derive every formula, it has to be consistent, which we can safely assume.
+Since \verb|G| is consistent, it
 can be extended into the canonical model \verb|M|. By the fact that
@@
 \begin{minted}{coq}
-  destruct FP with (fun X => [! X -> P !]) as (psi, H2).
+  destruct FP with (fun X => [! X -> p !]) as (psi, H2).
 \end{minted}
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! psi -> []psi -> P !]) as H3.
+  assert (A; G |-- [! psi -> []psi -> p !]) as H3.
   apply modal_ax5 in H2; auto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! [](psi -> []psi -> p) !])
-    as H4.
+  assert (A; G |-- [! [](psi -> []psi -> p) !]) as H4.
   apply Nec; auto.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []psi -> []([]psi -> p) !])
-    as H5.
+  assert (A; G |-- [! []psi -> []([]psi -> p) !]) as H5.
   apply modal_axK in H4; auto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []([]psi -> p) ->
-                          [][]psi -> []p !]) as H6.
+  assert (A; G |-- [! []([]psi -> p) -> [][]psi -> []p !]) as H6.
   eapply Ax with (a := axK ?[X] ?[Y]); auto with modal.
@@
 $\vdash \Box \Psi \to \Box \Box \Psi \to \Box P$ using transitivity of
-implication.
-
-\begin{minted}{coq}
-  assert (A; G |-- [! []psi -> [][]psi -> []p !])
-    as H7.
+implication,
+which we defined in an auxiliary lemma.
+
+\begin{minted}{coq}
+  assert (A; G |-- [! []psi -> [][]psi -> []p !]) as H7.
   eapply modal_compose; eauto with modal.
@@
 modeling of two well known puzzles.
-De Wind's formalization was separated into some components. The first consists in
-defining the syntax of formulas and its Kripke semantics. Next, the author uses
+De Wind's formalization was separated into some components: the first consists in
+defining the syntax of formulas and its Kripke semantics, and next, the author uses
 their definitions to give some examples of modal operators, worlds and how they
@@
 introduction and elimination rules for modal logic operators and use them to
-model the wize man's and muddy children puzzles.
+model the wise man's and muddy children puzzles.
+
+\pagebreak
 
@@
 system of normal modal logics, unlike any other formalizations found in the literature.
+In fact, our completeness result makes no assumption over the set of axioms other than it extends system K.
 
@@
 as a predicate parameter in the deduction inductive type. Using this generic
-representation of deduction we are able to prove soundness and completeness theorems, and, in order to demonstrate the applicability of the proposed work, prove L\"ob's theorem
+representation of deduction we are able to prove soundness and completeness theorems along with a demonstration of frame correctness, and, in order to demonstrate the applicability of the proposed work, prove L\"ob's theorem
 by assuming fixed points. % All source code in this article has
@@
 
-As future work we intend to continue the development of our library by
-including support for multimodalities (fusion of modal logics) and
+As future work, we intend to continue the development of our library by
+including support for different systems such as for multimodalities (fusion of modal logics) and
 non-normal modal logic systems, along with improving usability through proof automation.
-{\color{red} TORRENS: Dedução natural e táticas.}
+The development also makes it clear that there is a necessity of better implementing tactics to deal directly with derivations in the embedded modal logic.
+As of now, several auxiliary tactics have been developed which were used in the completeness result, but which we intend to extend by adding custom tactics and further taking advantage of Coq's proof automation facilities.
 

### 2025-08-19 paper/COLA/paper.tex
Fix a few typos

@@
 \ead{miguel.alfredo.nunes@gmail.com}
-\address{Santa Catarina State University}
+\address{State University of Campinas}
 
@@
 % it is possible to formalize and verify highly complex computer systems.
-we argue we may achieve betters ways of formally describing and verifying complex programs.
+we argue we may achieve better ways of formally describing and verifying complex programs.
 
@@
 \end{minted}
-However, even for very simple theorems, coding a definition directly
+However, even for very simple theorems, coding a proof directly
 as a Coq term can be a hard task. Because of this, the use of tactics
@@
 it holds on all models $\mathcal{M} = \langle \mathcal{F},\mathcal{V}\rangle$, i.e.,
-for any $\mathcal{M}$, we have that $\mathcal{M}\models\varphi$. A formula
+for any such $\mathcal{M}$, we have that $\mathcal{M}\models\varphi$. A formula
 is valid, written $\models \varphi$, if it is valid in any frame $\mathcal{F}$. Using the above semantics, we can define the concept of \emph{entailment}: given a set
@@
 Apart from the axioms, we also need \emph{modus ponens} and necessitation rules.
+The complete deduction system is as follows:
 
 \begin{figure}[H]
-\begin{tabular}{cc}
+\begin{tabular}{cccc}
+    \AxiomC{$\varphi \in \Gamma$}
+    \RightLabel{Prem}
+    \UnaryInfC{$\Gamma\vdash \varphi$}
+    \DisplayProof
+  &
+    \AxiomC{$\varphi\:\textit{is an axiom}$}
+    \RightLabel{Ax}
+    \UnaryInfC{$\Gamma \vdash \varphi$}
+    \DisplayProof
+  &
     \AxiomC{$\Gamma\vdash\varphi_1\to\varphi_2$}
@@
   &
-    \AxiomC{$\Gamma\vdash\varphi$}
+    \AxiomC{$\vdash\varphi$}
     \RightLabel{Nec}
@@
 \end{figure}
-The set of axioms and the two previous rules form a sound and complete Hilbert-style
+\noindent In the SBLP 2022 version of this work, we use a different deductive system in which
+the necessitation rule was as follows:
+\[
+    \AxiomC{$\Gamma\vdash\varphi$}
+    \RightLabel{Nec}
+    \UnaryInfC{$\Gamma\vdash\Box \varphi$}
+    \DisplayProof
+\]
+However, this system does not admit the deduction theorem. Since we aim for a
+completeness result and it implies the deduction theorem, the system which we used in
+the previous version can't be complete and hence requires this changes.
+For a clarification on the change and the role of the deduction theorem in modal logic,
+we refer the reader to~\cite{HakliN12}.
+The set of axioms and the two last rules form a sound and complete Hilbert-style
 deductive system for modal logics~\cite{blackburn}. Given a formula $\varphi$ and a
@@
 It's worth to mention about the choice of Hilbert-style systems as the deductive system in our library.
-The main reason is itself the soundness and completeness proofs: deductive systems such as natural deduction~\cite{Russo_PMLD, basin.ea:labelled:1997, SVRS02-recipe}, sequent calculi~\cite{Governatori00labelledmodal, 03-MSSV-fiblog19}, or tableaux~\cite{BrodaRusso05}, in modal logics, usually need the aid of labels or provisos.
-This makes, for instance, the metatheorem of soundness harder to proof: while Hilbert systems need only a simple structural induction on the size of the proof, labeled systems need to deal with several corner cases.
-The disadvantage of the Hilbert system is to use it as a way to establish proofs \emph{in} the system itself, which is not ``direct'' or ``natural''.
-However, it is still possible to use the $\mathcal{L}$tac language in Coq to automatize parts of this work.
+The main reason is itself the soundness and completeness proofs: deductive systems such as natural
+deduction~\cite{Russo_PMLD, basin.ea:labelled:1997, SVRS02-recipe}, sequent
+calculi~\cite{Governatori00labelledmodal, 03-MSSV-fiblog19}, or tableaux~\cite{BrodaRusso05},
+in modal logics, usually need the aid of labels or provisos. This makes, for instance, the
+metatheorem of soundness harder to proof: while Hilbert systems need only a simple structural induction
+on the size of the proof, labeled systems need to deal with several corner cases. The disadvantage of the
+Hilbert system is to use it as a way to establish proofs \emph{in} the system itself, which is not
+``direct'' or ``natural''. However, it is still possible to use the $\mathcal{L}$tac language in Coq
+to automatize parts of this work.
 
@@
                 (f: formula)
-                (d1: deduction A t f),
+                (d1: deduction A Empty i),
          deduction A t ([] f).
@@
 %-------------------------------------------------------------------------------
-The semantic entailment, $\Gamma\models\varphi$, represents a global consequence since, by our definition, we expect that the hypotheses are valid in every world in a model instead of only on those worlds where $\varphi$ hold \cite{de1999global}.
-Also, the completeness result may be weak, which is only valid without hypotheses, or strong, allowing it to be stated with any set of hypotheses \cite{blackburn}.
-We follow the definitions as stabilished by Popkorn \cite{Popkorn_1994} and note that we prove a strong completeness result using a global consequence: as this subsumes both weak completeness and a local one (as local consequence implies global consequence \cite{de1999global}), our result allows us to derive other possible definitions of completeness, while still being a weak result enough that it will work for any of the possible sets of axioms \cite{Popkorn_1994}.
+The semantic entailment, $\Gamma\models\varphi$, represents a global consequence since, by our definition,
+we expect that the hypotheses are valid in every world in a model instead
+of having $\varphi$ hold only on those worlds where the hypotheses are valid \cite{de1999global}.
+Also, the completeness result may be
+weak, which is only valid without hypotheses, or strong, allowing it to be stated with any set of
+hypotheses \cite{blackburn}. We follow the definitions as stabilished by Popkorn \cite{Popkorn_1994}
+and note that we prove a strong completeness result using a global consequence: as this subsumes
+both weak completeness and a local one (as local consequence implies global consequence
+\cite{de1999global}), our result allows us to derive other possible definitions of completeness,
+while still being a weak result enough that it will work for any of the possible sets of
+axioms \cite{Popkorn_1994}.
 
@@
 \end{minted}
-Notation \mintinline{coq}{M ' w ||- p} denotes the valuation of a fórmula at a given world \verb|w| of the
+Notation \mintinline{coq}{M ' w ||- p} denotes the valuation of a formula at a given world \verb|w| of the
 canonical model. Note that a world in the canonical model is an (infinite) consistent set of formulas.
@@
 We represent the existence of modal fixed points for all one-variable formulas by a definition \mintinline{coq}{fixed_point},
-which is parameterized by a system of modal logic, \mintinline{coq}{S}, and a
+which is parameterized by a system of modal logic, \mintinline{coq}{A}, and a
 theory, \mintinline{coq}{G}, and encodes the logical equivalence demonstrating
@@
   Theorem Lob:
-    forall A G, subset K4 A /\ fixed_point A G ->
-    forall p, (A; G |-- [! []p -> p !]) ->
-              (A; G |-- [! p !]).
+    forall A, subset K4 A /\ fixed_point A Empty ->
+    forall p, (A; Empty |-- [! []p -> p !]) ->
+              (A; Empty |-- [! p !]).
 \end{minted}
@@
 \begin{minted}{coq}
-  intros A G [I FP] P H1.
+  intros A [I FP] P H1.
 \end{minted}
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! psi -> []psi -> p !]) as H3.
+  assert (A; Empty |-- [! psi -> []psi -> p !]) as H3.
   apply modal_ax5 in H2; auto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! [](psi -> []psi -> p) !]) as H4.
+  assert (A; Empty |-- [! [](psi -> []psi -> p) !]) as H4.
   apply Nec; auto.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []psi -> []([]psi -> p) !]) as H5.
+  assert (A; Empty |-- [! []psi -> []([]psi -> p) !]) as H5.
   apply modal_axK in H4; auto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []([]psi -> p) -> [][]psi -> []p !]) as H6.
+  assert (A; Empty |-- [! []([]psi -> p) -> [][]psi -> []p !]) as H6.
   eapply Ax with (a := axK ?[X] ?[Y]); auto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []psi -> [][]psi -> []p !]) as H7.
+  assert (A; Empty |-- [! []psi -> [][]psi -> []p !]) as H7.
   eapply modal_compose; eauto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []psi -> [][]psi !]) as H8.
+  assert (A; Empty |-- [! []psi -> [][]psi !]) as H8.
   apply modal_axK4; auto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []psi -> []p !]) as H9.
+  assert (A; Empty |-- [! []psi -> []p !]) as H9.
   eapply modal_ax2; eauto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []psi -> p !]) as H10.
+  assert (A; Empty |-- [! []psi -> p !]) as H10.
   eapply modal_compose; eauto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! ([]psi -> p) -> psi !]) as H11.
+  assert (A; Empty |-- [! ([]psi -> p) -> psi !]) as H11.
   apply modal_ax6 in H2; auto with modal.
@@
 \begin{minted}{coq}
-  assert (A; G |-- psi) as H12.
+  assert (A; Empty |-- psi) as H12.
   eapply Mp; try eassumption.
@@
 \begin{minted}{coq}
-  assert (A; G |-- [! []psi !]) as H13.
+  assert (A; Empty |-- [! []psi !]) as H13.
   apply Nec; try eassumption.
@@
 model the wise man's and muddy children puzzles.
-
-\pagebreak
 

### 2025-08-20 paper/COLA/paper.tex
Future work: soundness for other systems
Clarify completeness result

@@
 i.e., the relation on the frame may be an arbitrary relation.
+We intend to cover soundness for larger systems as future work, which requires
+checking the additional axioms individually.
 
@@
 weak, which is only valid without hypotheses, or strong, allowing it to be stated with any set of
-hypotheses \cite{blackburn}. We follow the definitions as stabilished by Popkorn \cite{Popkorn_1994}
-and note that we prove a strong completeness result using a global consequence: as this subsumes
-both weak completeness and a local one (as local consequence implies global consequence
-\cite{de1999global}), our result allows us to derive other possible definitions of completeness,
-while still being a weak result enough that it will work for any of the possible sets of
-axioms \cite{Popkorn_1994}.
+hypotheses \cite{blackburn}. We follow the definitions as established by Popkorn \cite{Popkorn_1994}
+and note that we prove a weak completeness result using a global consequence, which subsumes a local one (as local consequence implies global consequence
+\cite{de1999global}). A strong completeness proof, as of now, is left as future work.
+It's worth noting that our completeness theorem is, in the words of
+Popkorn, \emph{rather weak} \cite[p. 119-120]{Popkorn_1994}; the result is applicable to any
+possible sets of axioms extending System K. Other
+notions of completeness, argueable more expressive, will relate sets of axioms
+to respective sets of model frames (cf. Section \ref{sec:correspondence}). Still, the present result provides a basis
+for most such stronger completeness theorems.
 

### 2025-08-20 paper/COLA/paper.tex
Some rephrasing; clarifications

@@
 of having $\varphi$ hold only on those worlds where the hypotheses are valid \cite{de1999global}.
+
 Also, the completeness result may be
@@
 and note that we prove a weak completeness result using a global consequence, which subsumes a local one (as local consequence implies global consequence
-\cite{de1999global}). A strong completeness proof, as of now, is left as future work.
+\cite{de1999global}). A strong completeness proof, as of now, is left as future work, as such result may be carried as an extension of the present one \cite{Popkorn_1994}.
 It's worth noting that our completeness theorem is, in the words of
@@
 possible sets of axioms extending System K. Other
-notions of completeness, argueable more expressive, will relate sets of axioms
+notions of completeness, considered more expressive, will relate sets of axioms
 to respective sets of model frames (cf. Section \ref{sec:correspondence}). Still, the present result provides a basis

### 2025-08-20 paper/COLA/paper.tex
Change definition ordering
Minor changes

@@
 this fact by two functions \mintinline{coq}{encode: formula -> nat} and \mintinline{coq}{decode: nat -> option formula}
-which satisfies the following property: $\forall \varphi . (\texttt{decode(encode }\varphi\texttt{)} = \varphi)$, stated in the \mintinline{coq}{countable} axiom.
+which satisfies the following property: $\forall \varphi . (\texttt{decode (encode }\varphi\texttt{)} = \varphi)$, stated in the \mintinline{coq}{countable} axiom.
 
@@
    | 0   => G
-   | S n' => Insert (decode n') (Delta n')
+   | S m => Insert (decode m) (Delta m)
    end.
@@
 \begin{minted}{coq}
+Definition UnionOf (A: nat -> theory): theory := fun t => exists i, A i t.
 Definition Max: theory := UnionOf Delta.
-Definition UnionOf (A: nat -> theory): theory := fun t => exists i, A i t.
 \end{minted}

### 2025-08-20 paper/COLA/paper.tex
Change some snippets

@@
 In the folllowing, we omit the $\Gamma$ subscripts. The canonical model is formalized in Coq as follows: first, we define the type \mintinline{coq}{W} that pack
-the construction of the maximal consistent sets for a theory \mintinline{coq}{G}.
+the construction of the maximal consistent sets for a theory.
 \begin{minted}{coq}
 Inductive W: Type :=
-| W_mk D: Consistent A D -> Maximal D -> Subset G D -> W.
+| W_mk D: Consistent A D -> Maximal D -> W.
 \end{minted}
@@
 \begin{minted}{coq}
-Lemma determination: forall p, (M |= p) <-> (A; G |-- p).
+Lemma determination: forall p, (M |= p) <-> (A; Empty |-- p).
 \end{minted}

### 2025-08-20 paper/COLA/paper.tex
Fix completeness theorem

@@
 \begin{minted}{coq}
-Theorem completeness: forall p, (G ||= p) -> (A; G |-- p).
+Theorem completeness: forall p, (Empty ||= p) -> (A; Empty |-- p).
 Proof.
+  (* Comments present in the source code omitted. *)
   intros p.
@@
   - apply classic.
-  - assert (Consistent A G).
-    + apply nonderivation_implies_consistency with p.
-      assumption.
+  - assert (Consistent A Empty).
+    + now apply nonderivation_implies_consistency with p.
     + assert ((M |= p) -> False); intros.
-      * apply determination with p in H2; auto.
-      * apply H2, H0.
-        intros q ? w; apply truth.
-        destruct w as (w, ?H, ?H, ?H); simpl.
-        apply H6; assumption.
+      * apply determination with p in H2.
+        contradiction.
+      * apply H2, H0; easy.
 Qed.
@@
 The proof proceeds by the contrapositive.
-Suppose that \mintinline{coq}{A; G |-- p} does not hold.
+Suppose that \mintinline{coq}{A; Empty |-- p} does not hold.
 As this system can't derive every formula, it has to be consistent, which we can safely assume.
-Since \verb|G| is consistent, it
+Since the empty context is clearly consistent, it
 can be extended into the canonical model \verb|M|. By the fact that
-\mintinline{coq}{A; G |-- p} is false, by the determination lemma, we know
-that \verb|p| is not true on the canonical model \verb|M|. Conclusion follows
-by the truth lemma.
+\mintinline{coq}{A; Empty |-- p} is false, by the determination lemma, we know
+that \verb|p| is not true on the canonical model \verb|M|. Conclusion then follows directly
+by our current hypotheses, namely that \mintinline{coq}{Empty ||= p} and \mintinline{coq}{(A; Empty |-- p) -> False} (which includes showing that all formulas in the empty set are valid).
 

### 2025-08-20 paper/COLA/paper.tex
Some rephrasing

@@
 that \verb|p| is not true on the canonical model \verb|M|. Conclusion then follows directly
-by our current hypotheses, namely that \mintinline{coq}{Empty ||= p} and \mintinline{coq}{(A; Empty |-- p) -> False} (which includes showing that all formulas in the empty set are valid).
+by our hypotheses, \mintinline{coq}{Empty ||= p} and \mintinline{coq}{(A; Empty |-- p) -> False} (which includes showing that all formulas in the empty set are valid).
 

### 2025-08-20 paper/COLA/paper.tex
Justify Löb's theorem

@@
 incompleteness results.
-
-In order to state L\"ob's theorem we need a way to encode modal
+%
+The purpose of this exercise is to formally prove that the existence of fixed
+points (defined below) in a modal logic system implies the GL axiom: for any formula $\Psi$, it
+holds that $\Box(\Box\Psi \to \Psi) \to \Box\Psi$. Among the systems defined in
+this paper, System GL possesses the fixed point property, though this proof is
+non-trivial \cite{fixpoint}; but, in this case, our conclusion becomes rather
+uninteresting: the GL axiom is already an axiom in System GL, which trivializes
+the result. Still, we justify this case study as other systems that do not have
+GL as an axiom but do possess the fixed point property (see \cite{fixpoint}) are
+definable, and the main purpose for this code is to formalize a piece of modal
+logic proof within Coq by using our framework.
+
+Proceeding,
+in order to state L\"ob's theorem we need a way to encode modal
 fixed points. Formally, let $\varphi(X)$ be a formula with only one propositional variable

### 2025-08-20 paper/COLA/paper.tex
Add minor comment
Clarify change from SBLP

@@
 a detailed discussion about our formalization of a completeness proof for the Hilbert-style deductive
-system following the strategies found in traditional modal logic books~\cite{Blackburn-ModalLogics}.
+system following the strategies found in traditional modal logic books~\cite{Blackburn-ModalLogics,Popkorn_1994},
+which was not present in the earlier version of the paper.
 The rest of this work is organized as follows: Section~\ref{sec:background} provides
@@
 
-\noindent We can now write down L\"ob's theorem as follows:
+\noindent We can now write down L\"ob's theorem as follows (representing the axiom through derivations):
 

### 2025-08-20 paper/COLA/paper.tex
Clarifications

@@
 Section~\ref{sec:implementation}. A case study in proving L\"ob's theorem is
-the subject of Section~\ref{sec:case}. Related works are discussed
+the subject of Section~\ref{sec:case}, showcasing an use of our library. Related works are discussed
 on Section~\ref{sec:related} and Section~\ref{sec:conclusion} draws some conclusions and
@@
 %
-All source code in this article has been formalized in Coq version 8.15 and
+All source code in this article has been formalized in Coq version 8.18 and
 is available at \url{https://github.com/funcao/LML/} under a permissive free software license.

### 2025-08-20 paper/COLA/paper.tex
List our axioms

@@
 
+Every result stated has been fully checked. We note, though, that the
+completeness result (Section \ref{sec:completeness}) relies on the excluded
+middle as an additional axiom, given that our completeness proof is fully classical;
+we also axiomatize that formulas (defined below) are countable, and leave as
+future work to properly demonstrate this.
+
 \subsection{Syntax and semantics of modal logics}

### 2025-08-20 paper/COLA/paper.tex
s/isomorphism/correspondence/

@@
 ``BHK-in\-ter\-pre\-ta\-tion''\footnote{Abbreviation of Brouwer, Heyting,
-  Kolmogorov, de Bruijn and Martin-L\"of Correspondence. This is also
-  known as the Curry-Howard isomorphism.}, where types represent
+  Kolmogorov, de Bruijn and Martin-L\"of correspondence. This is also
+  known as the Curry-Howard correspondence.}, where types represent
 logical formulas, $\lambda$-terms represent proofs

### 2025-08-21 paper/COLA/paper-sblp.tex
changing authors.
Signed-off-by: Rodrigo Ribeiro <rodrigogribeiro@gmail.com>

--- 2022-09-05 paper/paper.tex
@@
 %% used to denote shared contribution to the research.
+\author{Miguel Nunes}
+\affiliation{\institution{Santa Catarina State University}\country{}}
+\email{miguel.alfredo.nunes@gmail.com}
 
@@
 
-\author{Miguel Nunes}
+\author{Paulo Torrens}
 \affiliation{\institution{Santa Catarina State University}\country{}}
-\email{miguel.alfredo.nunes@gmail.com}
+\email{paulotorrens@gnu.org}
 
@@
 \email{karina.roggia@udesc.br}
-
-\author{Paulo Torrens}
-\affiliation{\institution{Santa Catarina State University}\country{}}
-\email{paulotorrens@gnu.org}
 
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{da Silveira et al.}
+\renewcommand{\shortauthors}{Nunes et al.}
 
@@
 
+\begin{CCSXML}
+  <ccs2012>
+    <concept>
+        <concept_id>10003752.10003790.10003793</concept_id>
+        <concept_desc>Theory of computation~Modal and temporal logics</concept_desc>
+        <concept_significance>500</concept_significance>
+        </concept>
+    <concept>
+        <concept_id>10003752.10003790.10002990</concept_id>
+        <concept_desc>Theory of computation~Logic and verification</concept_desc>
+        <concept_significance>500</concept_significance>
+        </concept>
+  </ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Theory of computation~Modal and temporal logics}
+\ccsdesc[500]{Theory of computation~Logic and verification}
+
 %%
@@
 %% the work being presented. Separate the keywords with commas.
-\keywords{Modal logics, soundness, Coq proof assistant}
+\keywords{Modal logic, soundness, Coq proof assistant}
 
@@
 
-However, writing a proof term whose type is that of a logical formula
-can be challenging, even for very simple propositions.  In order to
-make the writing of complex proofs easier, Coq provides
+In order to make writing proofs easier, Coq provides
 \emph{tactics}, which are commands that can be used to construct proof
-terms in a more user-friendly way.
-
-As a tiny example, consider the task of proving the following 
+terms in a more user-friendly way. As an example, consider the task of proving the following
 formula of the propositional logic:
